### 2022-09-15 max_dp endogen stød.Rmd
commit

---
title: "Model LF"
author: "Simon"
date: "11/7/2022"
output:
  pdf_document: default
  html_document: default
---

# Packages

```{r}
rm(list = ls())

library(tstools)
library(readxl)
library(tidyverse)
```

# Import data

```{r, message=FALSE}
AAMK <- read_excel("AAMK.xlsx")
globaldemand1 <- read_excel("globaldemand1.xlsx")
w_h_r1 <- read_excel("w_h_r1.xlsx")
emp1= ts(w_h_r1$emp1, start = c(2003,1), frequency = 4)
w_h_r1= ts(w_h_r1$w_h_r1, start = c(2003,1), frequency = 4)

names(AAMK) <- tolower(names(AAMK))
names(globaldemand1) <- tolower(names(globaldemand1))

AAMK=lapply(AAMK, function(t) ts(t, start=c(2005, 2), end=c(2020, 1),frequency=4))
globaldemand1=lapply(globaldemand1, function(t) ts(t, start=c(2005, 2), end=c(2020, 1),frequency=4))

attach(AAMK)
attach(globaldemand1)
```

```{r}
## Change name in globaldemand1

g_20= globaldemand1$`g-20`

```

```{r}
eq_nf_l_nf = res_eq_nf_l
eq_f_l_f = res_eq_f_l
```

# Create dummies

```{r}
d_2006q1=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2006,1), dummy_end = c(2006,1),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2006q2=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2006,2), dummy_end = c(2006,2),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2006q3=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2006,3), dummy_end = c(2006,3),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2006q4=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2006,4), dummy_end = c(2006,4),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2007q34=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2007,3), dummy_end = c(2007,4),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2007q3=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2007,3), dummy_end = c(2007,3),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2017q23=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2017,2), dummy_end = c(2017,3),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2008q2=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2008,2), dummy_end = c(2008,2),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2008q3=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2008,3), dummy_end = c(2008,3),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2008q4=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2008,4), dummy_end = c(2008,4),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2009q4=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2009,4), dummy_end = c(2009,4),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2018q2=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2018,2), dummy_end = c(2018,2),
 basic_value = 0, dummy_value = 1, frequency = 4)

##### Er der noget specielt med dem her?????

d_20123 =  ts(c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), start = c(2005, 2), end =  c(2020,1), frequency = 4)

d_20124 =  ts(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1), start = c(2005, 2), end =  c(2020,1), frequency = 4)

#### Videre igen

d_2009q1=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2009,1), dummy_end = c(2009,1),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2009q2=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2009,2), dummy_end = c(2009,2),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2011q1=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2011,1), dummy_end = c(2011,1),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2014q3=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2014,3), dummy_end = c(2014,3),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2014q4=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2014,4), dummy_end = c(2014,4),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2013q4=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2013,4), dummy_end = c(2013,4),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2018q1=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2018,1), dummy_end = c(2018,1),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2016q3=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2016,3), dummy_end = c(2016,3),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2019q3=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2019,3), dummy_end = c(2019,3),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2019q4=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2019,4), dummy_end = c(2019,4),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2020q1=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2020,1), dummy_end = c(2020,1),
 basic_value = 0, dummy_value = 1, frequency = 4)

# DUMMY 4
dummy_2018q34 =create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2018,3), dummy_end = c(2018,4),
 basic_value = 0, dummy_value = 1, frequency = 4)

dummy_2011q1 =create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2011,1), dummy_end = c(2011,1),
 basic_value = 0, dummy_value = 1, frequency = 4)

dummy_2009q2 =create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2009,2), dummy_end = c(2009,2),
 basic_value = 0, dummy_value = 1, frequency = 4)

time= ts(1:length(dummy_2009q2), start = c(2005,2), frequency = 4)

DUMMY_4= ts(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,0,0,0,0,0), start = c(2005, 2), end =  c(2020,1), frequency = 4)

DUMMY_10= ts(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0), start = c(2005, 2), end =  c(2020,1), frequency = 4)

DUMMY_11= ts(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,0,0,0,0,0), start = c(2005, 2), end =  c(2020,1), frequency = 4)

#### Variable der kun har en værdi:

wage_2010q3 = ts(c(86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268,86.36268), start = c(2005, 2), end =  c(2020,1), frequency = 4)

tax_rate2= ts(c(0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767,0.152767), start = c(2005, 2), end =  c(2020,1), frequency = 4)

tax_rate1= ts(c(0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369,0.389369), start = c(2005, 2), end =  c(2020,1), frequency = 4)

zz= ts(c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1), start = c(2005, 2), end =  c(2020,1), frequency = 4)

urs= ts(c(0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04,0.04), start = c(2005, 2), end =  c(2020,1), frequency = 4)

shock= ts(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), start = c(2005, 2), end =  c(2020,1), frequency = 4)

```

Dummies til stød

```{r}
d_2013q4=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2013,4), dummy_end = c(2013,4),
 basic_value = 0, dummy_value = 1, frequency = 4)

d_2017q1=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2017,1), dummy_end = c(2017,1),
 basic_value = 0, dummy_value = 1, frequency = 4)
```

# Real Variables

```{r}
pc = cpi1/100
px = px/100
pi_1 = pi/100
pg= pg/100
pm = pm/100
py = py/100
pf = pc*xr/rer
rer = pc*xr/pf
pconk = pcon/pc
gk = g/pg
xk = x/px
mk = m/pm
```

```{r}

i_adj_h = i_h -  (i_equip_h + i_bd_h)
i_adj_nfc = i_nf - (i_equip_nfc + i_bd_nfc)
i_adj_fc = i_f - (i_equip_fc + i_bd_fc)
i_adj_g = i_g - (i_equip_g +i_bd_g)
```

```{r}
i_equip_h_k = i_equip_h/p_equip
i_equip_nfc_k =  i_equip_nfc/p_equip
i_equip_g_k = i_equip_g/p_equip
i_equip_fc_k =  i_equip_fc/p_equip
```

```{r}
i_bd_h_k =  i_bd_h/p_bd
i_bd_nfc_k =  i_bd_nfc/p_bd
i_bd_g_k =  i_bd_g/p_bd
i_bd_fc_k =  i_bd_fc/p_bd
```

I just deflate the adjustment terms using p_equip as well (assuming most of it is changes in inventories)

```{r}
i_adj_h_k = i_adj_h /p_equip
i_adj_nfc_k = i_adj_nfc /p_equip
i_adj_fc_k = i_adj_fc /p_equip
i_adj_g_k = i_adj_g /p_equip
```

we can compare ik with ikk. just to make sure we have all components of investment.

```{r}
ik =  i_equip_h_k + i_equip_nfc_k  +  i_equip_g_k + i_equip_fc_k +  i_bd_h_k + i_bd_nfc_k  +  i_bd_g_k  + i_bd_fc_k + i_adj_h_k + i_adj_nfc_k + i_adj_fc_k + i_adj_g_k

ikk = i/pi_1
```

```{r}
yk = ik + pconk + gk + xk - mk
```

```{r}

npropinc_h = propinc_r_h - propinc_p_h
nsben_h = sben_h_r - scon_h_p
```

```{r}
iba_f = - (iba_h + iba_nf + iba_g + iba_row) # The data from the database is not consistent (the sum of the stocks of IBA does not add up to zero). I fix it by assuming that iba_f is entirely demand-led.

iba_f_tr = - (iba_h_tr + iba_nf_tr + iba_g_tr + iba_row_tr) #The data from the database is not consistent (the sum of IBA transactions does not add up to zero). I fix it by assuming that iba_f_tr is entirely demand-led.
```

```{r}
### Jeg tager bare og fjerner første observation fra dem der er minuse så den først starer i 2005q3

iba_h_rvx=diff(iba_h)-iba_h_tr[-1]
iba_f_rvx=diff(iba_f)-iba_f_tr[-1]
iba_nf_rvx=diff(iba_nf)-iba_nf_tr[-1]
iba_g_rvx=diff(iba_g)-iba_g_tr[-1]
iba_row_rvx=diff(iba_row)-iba_row_tr[-1]
```

```{r}
eq_h = - ((eq_f_a - eq_f_l) + (eq_nf_a - eq_nf_l) + eq_g + (eq_row_a - eq_row_l)) #The data from the database is not consistent (the sum of the stocks of equities does not add up to zero). I fix it by assuming that eq_h absorbbs the small discrepancies that come from the database.
```

```{r}
### Der ages diff så de sarer alle i 2005Q3

eq_h_rvx=diff(eq_h)-eq_h_tr[-1]
eq_f_a_rvx=diff(eq_f_a)-eq_f_a_tr[-1]
eq_f_l_rvx=diff(eq_f_l)-eq_f_l_tr[-1]
eq_nf_a_rvx=diff(eq_nf_a)-eq_nf_a_tr[-1]
eq_nf_l_rvx=diff(eq_nf_l)-eq_nf_l_tr[-1]
eq_g_rvx=diff(eq_g)-eq_g_tr[-1]
eq_row_a_rvx=diff(eq_row_a)-eq_row_a_tr[-1]
eq_row_l_rvx=diff(eq_row_l)-eq_row_l_tr[-1]
```

```{r}

### Tror også alle disse starer i 2005Q3

Neq_row = eq_row_a - eq_row_l 
Neq_nf = eq_nf_a - eq_nf_l 
Neq_f = eq_f_a - eq_f_l 
Neq_nf_rvx = eq_nf_a_rvx-eq_nf_l_rvx
Neq_f_rvx = eq_f_a_rvx-eq_f_l_rvx
Neq_row_rvx = eq_row_a_rvx-eq_row_l_rvx
Neq_nf_tr = diff(Neq_nf)-(Neq_nf_rvx)
Neq_f_tr = diff(Neq_f)-(Neq_f_rvx)
alpha_neq_nf = -Neq_nf/i_nf
Neq_row_tr = diff(Neq_row) - Neq_row_rvx
```

```{r}
eq_h_tr = - (Neq_nf_tr  + Neq_f_tr + eq_g_tr[-1] + Neq_row_tr) # The data from the database is not consistent (the sum of the stocks of equities does not add up to zero). I fix it by assuming that eq_h_tr absorbbs the small discrepancies that come from the database.
```

```{r}
eq_f_l_rvx = diff(eq_f_l)-eq_f_l_tr[-1]
eq_nf_l_rvx = diff(eq_nf_l)-eq_nf_l_tr[-1]
eq_row_l_rvx = diff(eq_row_l)-eq_row_l_tr[-1]
```

```{r}
ins_f = - (ins_row + ins_h + ins_nf + ins_g) # The database is almost consistent, but I redifine ins_f to make it fully consistent and avoid carrying discrepancies along when running the simulations.
ins_f_tr = - (ins_row_tr + ins_h_tr + ins_nf_tr + ins_g_tr) # The database is almost consistent, but I redifine ins_f_tr to make it fully consistent and avoid carrying discrepancies along when running the simulations.
```

```{r}
ins_h_rvx=diff(ins_h)-ins_h_tr[-1]
ins_f_rvx=diff(ins_f)-ins_f_tr[-1]
ins_nf_rvx=diff(ins_nf)-ins_nf_tr[-1]
ins_g_rvx=diff(ins_g)-ins_g_tr[-1]
ins_row_rvx=diff(ins_row)-ins_row_tr[-1]
```

```{r}
sec_g = - (sec_row + sec_h + sec_f_a + sec_f_d + sec_nf) # The database is almost consistent, but I redifine sec_g to make it fully consistent and avoid carrying discrepancies along when running the simulations.
sec_g_tr = - (sec_row_tr + sec_h_tr + sec_f_a_tr + sec_f_d_tr + sec_nf_tr) # The database is almost consistent, but I redifine sec_g to make it fully consistent and avoid carrying discrepancies along when running the simulations.
```

```{r}
sec_h_rvx=diff(sec_h)-sec_h_tr[-1]
sec_f_a_rvx=diff(sec_f_a)-sec_f_a_tr[-1]
sec_f_d_rvx=diff(sec_f_d)-sec_f_d_tr[-1]
sec_nf_rvx=diff(sec_nf)-sec_nf_tr[-1]
sec_g_rvx=diff(sec_g)-sec_g_tr[-1]
sec_row_rvx=diff(sec_row)-sec_row_tr[-1]
```

```{r}
l_f = - (l_h + l_nf + l_g + l_row) # The database is almost consistent, but I redifine l_f to make it fully consistent and avoid carrying discrepancies along when running the simulations.
l_f_tr = - (l_h_tr + l_nf_tr + l_g_tr + l_row_tr) # The database is almost consistent, but I redifine l_f to make it fully consistent and avoid carrying discrepancies along when running the simulations.
```

```{r}
l_h_rvx=diff(l_h)-l_h_tr[-1]
l_f_rvx=diff(l_f)-l_f_tr[-1]
l_nf_rvx=diff(l_nf)-l_nf_tr[-1]
l_g_rvx=diff(l_g)-l_g_tr[-1]
l_row_rvx=diff(l_row)-l_row_tr[-1]
```

```{r}
int_nf_adj = (int_r_nf - int_p_nf) - (stats::lag(idep,-1)*stats::lag(iba_nf,-1) + stats::lag(ibd,-1)*stats::lag(sec_nf,-1) + stats::lag(iloan,-1)*stats::lag(l_nf,-1))

int_h_adj = (int_r_h - int_p_h) - (stats::lag(idep,-1)*stats::lag(iba_h,-1) +stats::lag(iloan,-1)*stats::lag(l_h,-1)+ stats::lag(ibd,-1)*stats::lag(sec_h,-1) ) # this calculation calculates the discrepancy for interest payments as net flow.
```

```{r}
#Below I (hamid) compute discrepancy between interest payments and reciepts separately to create a measure of propert income (paid and recieved) instead of just using net property income (npropinc) in the model

int_r_h_adj = int_r_h  - (stats::lag(idep,-1)*stats::lag(iba_h,-1)  + stats::lag(ibd,-1)*stats::lag(sec_h,-1))

int_p_h_adj = int_p_h  + stats::lag(iloan,-1)*stats::lag(l_h,-1)
```

```{r}
#For the rest of the sectors, we just calculate the discrepancies based on net asset holding
int_g_adj = (int_r_g - int_p_g) - (stats::lag(idep,-1)*stats::lag(iba_g,-1) + stats::lag(iloan,-1)*stats::lag(l_g,-1) + stats::lag(ibd,-1)*stats::lag(sec_g,-1))

int_row_adj = (int_r_row - int_p_row) - (stats::lag(idep,-1)*stats::lag(iba_row,-1) + stats::lag(iboa,-1)*stats::lag(sec_row,-1) + stats::lag(iloan,-1)*stats::lag(l_row,-1))

int_f_adj = (int_r_f - int_p_f) - (stats::lag(idep,-1)*stats::lag(iba_f,-1) + stats::lag(iloan,-1)*stats::lag(l_f,-1) + stats::lag(ibd,-1)*stats::lag(sec_f_d,-1) + stats::lag(iboa,-1)*stats::lag(sec_f_a,-1))
```

```{r}
test = int_nf_adj  + int_h_adj  + int_g_adj  + int_f_adj  + int_row_adj
```

Below we create series for equities viz-a-viz counterparties. The data is only available from 2012 onwards so we use the porportion w.r.t to the aggregate payments to create data back in time (because we have aggregate data)

```{r}
alpha_nf_rv = eq_h_nf_rv[31:60]/(eq_h_nf_rv[31:60] + eq_h_f_rv[31:60] + eq_h_row_rv[31:60]) ## mean of alpha is 0.57 (excluding the spike in the start of the sample somwhere 2013q3 (not sure about the exact point))
alpha_f_rv = eq_h_f_rv[31:60]/(eq_h_nf_rv[31:60] + eq_h_f_rv[31:60] + eq_h_row_rv[31:60]) #  mean of alpha is 0.37 (excluding the spike in the start of the sample somwhere 2013q3 (not sure about the exact point))
alpha_row_rv = eq_h_row_rv[31:60]/(eq_h_nf_rv[31:60] + eq_h_f_rv[31:60] + eq_h_row_rv[31:60]) # mean of alpha is 0.06 (excluding the spike in the start of the sample somwhere 2013q3 (not sure about the exact point))
```

dummies for equity revaluation equations

```{r}
dummy_1=ts(c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0), start = c(2005, 2), end = c(2020,1), frequency = 4)

dummy_2=create_dummy_ts(start_basic = c(2005, 2), end_basic=c(2020,1),
 dummy_start=c(2013,1), dummy_end = c(2020,1),
 basic_value = 0, dummy_value = 1, frequency = 4)
```

```{r}
eq_h_row_rvx = dummy_1*0.06*eq_h_rvx + dummy_2*eq_h_row_rv
eq_h_nf_rvx = dummy_1*0.57*eq_h_rvx + dummy_2*eq_h_nf_rv
eq_h_f_rvx = dummy_1*0.37*eq_h_rvx + dummy_2*eq_h_f_rv
eq_h_rvx1 = eq_h_nf_rvx + eq_h_f_rvx + eq_h_row_rvx
```

We now create stock of equities data based on counter parties.

```{r}
alpha_nf = eq_h_nf[31:60]/eq_h[31:60] #mean of the series 0.275
alpha_f = eq_h_f[31:60]/eq_h[31:60] #mean of the series 0.663
#alpha_row = eq_h_row/eq_h
alpha_row = 1 - alpha_nf - alpha_f

### I create these to keep the same time dimension

alpha_nf= ts(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,alpha_nf),start = c(2005,2), frequency = 4)

alpha_f= ts(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,alpha_f),start = c(2005,2), frequency = 4)

alpha_row= ts(c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,alpha_row),start = c(2005,2), frequency = 4)

```

```{r}
#eq_h_nf = dummy_1*0.275*eq_h + dummy_2*eq_h_nf
#eq_h_f = dummy_1*0.663*eq_h + dummy_2*eq_h_f
#eq_h_row = dummy_1*0.062*eq_h  + dummy_2*eq_h_row 
```

```{r}
gamma_1 = eq_h_nf / eq_h
gamma_2 = eq_h_f / eq_h
```

```{r}
#series  iba_h_tr = fn
```

```{r}
library(stats)

div_nf2_adj = (div_r_nf  - div_p_nf) - ((stats::lag(eq_nf_a_nf,-1) + stats::lag(eq_nf_a_f,-1) - stats::lag(eq_nf_l_nf,-1) + stats::lag(eq_h_nf,-1)+ stats::lag(eq_row_a_nf,-1) + stats::lag(eq_f_a_nf,-1) + stats::lag(eq_g_nf,-1)))* stats::lag(divd,-1) + stats::lag(eq_nf_a_row,-1)*stats::lag(diva,-1)
```

```{r}
div_nf1_adj =  (div_r_nf  - div_p_nf) - (stats::lag(Neq_nf,-1)*stats::lag(divd,-1))
```

```{r}
div_nf_adj = d_20124*(div_nf2_adj) + (1-d_20124)*(div_nf1_adj)
```

```{r}
div_nf =d_20124*(((stats::lag(eq_nf_a_nf,-1) + stats::lag(eq_nf_a_f,-1) - stats::lag(eq_nf_l_nf,-1) + stats::lag(eq_h_nf,-1) + stats::lag(eq_row_a_nf,-1) + stats::lag(eq_f_a_nf,-1) + stats::lag(eq_g_nf,-1)))* stats::lag(divd,-1) + stats::lag(eq_nf_a_row,-1)*stats::lag(diva,-1)) + (1-d_20124)*(stats::lag(Neq_nf,-1)*stats::lag(divd,-1)) + div_nf_adj

div_nf1 =  div_r_nf - div_p_nf
```

```{r}
div_f2_adj = (div_r_f  - div_p_f) - ((stats::lag(eq_f_a_nf,-1) + stats::lag(eq_f_a_f,-1) - (stats::lag(eq_f_l_f,-1) + stats::lag(eq_h_f,-1)+ stats::lag(eq_row_a_f,-1) + stats::lag(eq_row_a_f,-1) + stats::lag(eq_g_f,-1)))* stats::lag(divd,-1) + stats::lag(eq_f_a_row,-1)*stats::lag(diva,-1))
```

```{r}
div_f1_adj =  (div_r_f  - div_p_f) - (stats::lag(Neq_f,-1)*stats::lag(divd,-1))

div_f_adj = d_20124*div_f2_adj + (1-d_20124)*(div_f1_adj)

```

```{r}
div_f = d_20124 *(((stats::lag(eq_f_a_f,-1) + stats::lag(eq_nf_a_nf,-1)) - (stats::lag(eq_f_l_f,-1) + stats::lag(eq_h_f,-1)+ stats::lag(eq_row_a_f,-1) + stats::lag(eq_nf_a_f,-1) + stats::lag(eq_g_f,-1)))* stats::lag(divd,-1) + stats::lag(eq_f_a_row,-1)*stats::lag(diva,-1)) + (1-d_20124)*((stats::lag(Neq_f,-1)*stats::lag(divd,-1))) +  div_f_adj

div_f1 =  div_r_f - div_p_f
```

```{r}
div_h2_adj = div_r_h - ((stats::lag(eq_h_nf,-1) + stats::lag(eq_h_f,-1))*stats::lag(divd,-1) + stats::lag(eq_h_row,-1)*stats::lag(diva,-1))

div_h1_adj =  div_r_h - (stats::lag(eq_h,-1)*stats::lag(divd,-1))

div_h_adj  = d_20124*div_h2_adj + (1-d_20124)*(div_h1_adj)
```

```{r}
div_h = d_20124*((stats::lag(eq_h_nf,-1) + stats::lag(eq_h_f,-1))*stats::lag(divd,-1) + stats::lag(eq_h_row,-1)*stats::lag(diva,-1)) + (1-d_20124)*(stats::lag(eq_h,-1)*stats::lag(divd,-1)) +  div_h_adj

div_h1 = div_r_h - div_p_h
```

```{r}
div_g2_adj = div_r_g - ((stats::lag(eq_g_nf,-1) + stats::lag(eq_g_f,-1))*stats::lag(divd,-1) + stats::lag(eq_g_row,-1)*stats::lag(diva,-1))

div_g1_adj =  div_r_g - (stats::lag(eq_g,-1)*stats::lag(divd,-1))

div_g_adj = d_20124*div_g2_adj + (1-d_20124)*(div_g1_adj)

div_g = d_20124 * ( (stats::lag(eq_g_nf,-1) + stats::lag(eq_g_f,-1))*stats::lag(divd,-1) + stats::lag(eq_g_row,-1)*stats::lag(diva,-1) ) + (1-d_20124) * ((stats::lag(eq_g,-1)*stats::lag(divd,-1))) + div_g_adj

div_g1 =  div_r_g - div_p_g
```

```{r}
div_row2_adj = (div_r_row - div_p_row) - ((stats::lag(eq_row_a_nf,-1) + stats::lag(eq_row_a_f,-1))*stats::lag(divd,-1) - ((stats::lag(eq_nf_a_row,-1)+stats::lag(eq_f_a_row,-1)+stats::lag(eq_g_row,-1)+stats::lag(eq_h_row,-1))*stats::lag(diva,-1)))

div_row1_adj = (div_r_row - div_p_row) - (stats::lag(Neq_row,-1)*stats::lag(divd,-1))

div_row_adj = d_20124*div_row2_adj + (1-d_20124)*(div_row1_adj)
```

```{r}
div_row = d_20124 * (  (stats::lag(eq_row_a_nf,-1) + stats::lag(eq_row_a_f,-1))*stats::lag(divd,-1) - ((stats::lag(eq_nf_a_row,-1)+stats::lag(eq_f_a_row,-1)+stats::lag(eq_g_row,-1)+stats::lag(eq_h_row,-1))*stats::lag(diva,-1))) + (1-d_20124) * ((stats::lag(Neq_row,-1)*stats::lag(divd,-1))) + div_row_adj
```

```{r}
div_row1 =  div_r_row - div_p_row
```

```{r}
check_div1 =  (div_r_row - div_p_row) +  (div_r_g - div_p_g) +  (div_r_h - div_p_h) +  (div_r_nf - div_p_nf) +  (div_r_f - div_p_f)

check_div = div_h + div_f + div_g + div_nf + div_row

check_div_adj1 = div_h1_adj  + div_f1_adj  +div_nf1_adj  +div_g1_adj  +div_row1_adj

check_div_adj2 = div_h2_adj  + div_f2_adj  +div_nf2_adj  +div_g2_adj  +div_row2_adj   
```

```{r}
ins_nf_adj = (ins_r_nf - ins_p_nf) - (stats::lag(ins_nf,-1)*stats::lag(insu,-1))
ins_f_adj = (ins_r_f - ins_p_f) - (stats::lag(ins_f,-1)*stats::lag(insu,-1))
ins_g_adj = (ins_r_g - ins_p_g) - (stats::lag(ins_g,-1)*stats::lag(insu,-1))
ins_h_adj = (ins_r_h - ins_p_h) - (stats::lag(ins_h,-1)*stats::lag(insu,-1))
ins_row_adj = (ins_r_row - ins_p_row) - (stats::lag(ins_row,-1)*stats::lag(insu,-1))
```

```{r}
ins_h_tr_excl_d8 = ins_h_tr - d8_h
```

Below I compute the adjustment variables that correct the discrepancy between net lending and financial net lending

```{r}
nl_h = s_h - i_h - np_h + ctr_h
fnl_h = iba_h_tr + sec_h_tr + l_h_tr + eq_h_tr + ins_h_tr
nl_h_adj = fnl_h -nl_h
```

```{r}
nl_nf = s_nf - i_nf - np_nf + ctr_nf
fnl_nf = iba_nf_tr + sec_nf_tr + l_nf_tr + Neq_nf_tr + ins_nf_tr
nl_nf_adj = fnl_nf - nl_nf
```

```{r}
nl_f = s_f - i_f - np_f + ctr_f
fnl_f = iba_f_tr + sec_f_a_tr + sec_f_d_tr + l_f_tr + eq_f_a_tr - eq_f_l_tr + ins_f_tr
nl_f_adj = fnl_f - nl_f
```

```{r}
nl_g = s_g - i_g - np_g + ctr_g
fnl_g = iba_g_tr + eq_g_tr + sec_g_tr + l_g_tr + ins_g_tr
nl_g_adj = fnl_g - nl_g 
```

```{r}
int_row = (stats::lag(idep,-1)*stats::lag(iba_row,-1) + stats::lag(iboa,-1)*stats::lag(sec_row,-1) + stats::lag(iloan,-1)*stats::lag(l_row,-1))

div_row = d_20124*(  (stats::lag(eq_row_a_nf,-1) + stats::lag(eq_row_a_f,-1))*stats::lag(divd,-1) - ((stats::lag(eq_nf_a_row,-1)+stats::lag(eq_f_a_row,-1)+stats::lag(eq_g_row,-1)+stats::lag(eq_h_row,-1))*stats::lag(diva,-1))) + (1-d_20124) * ((stats::lag(Neq_row,-1)*stats::lag(divd,-1))) + div_row_adj

insu_row = stats::lag(ins_row,-1)*stats::lag(insu,-1)

npropinc_row = int_row + int_row_adj + div_row + insu_row + ins_row_adj + (res_r_row) - (res_p_row)

nl_row = m - x + p_tax_row - p_sub_row + (w_row_r - w_row_p) +  (npropinc_row) + (tax_row) + scon_row_r - scon_row_p + sben_row_r - sben_row_p + oth_row + ctr_row - np_row

#nl_row1 = s_row + ctr_row - np_row

fnl_row = iba_row_tr + Neq_row_tr  + sec_row_tr + l_row_tr + ins_row_tr

nl_row_adj = fnl_row - nl_row 
```

```{r}
sumfnl = fnl_h + fnl_f + fnl_nf + fnl_g + fnl_row
sumnl = nl_h + nl_f + nl_nf + nl_g + nl_row
sumnladj = nl_h_adj + nl_f_adj + nl_nf_adj + nl_g_adj + nl_row_adj
```

tax rates for wages and property income (the two tax rates are determined throgh an exogenous regression without an intercept

```{r}
tax_h_hat =  tax_rate1*(w_h_r +nsben_h + oth_h) +  tax_rate2*(propinc_r_h - propinc_p_h + b2_h_r) 
tax_h_adj = (-tax_h -  tax_h_hat)
tax_h =  -(tax_h_hat + tax_h_adj)
yd2a_h =(1 - tax_rate2)*(propinc_r_h + b2_h_r) # property income part of the disposable income (capitalists)
yd2b_h = (1 - tax_rate2)*(-propinc_p_h) 
yd1_h =(1 - tax_rate1)*(w_h_r + nsben_h + oth_h) - tax_h_adj # wage and benefits part of the disposable income (workers)
tax_rate_nf = -tax_nf/y
```

Creating Real variables

```{r}
yd_hk=yd_h/pc # real disposable income
 w_h_rk = w_h_r/pc #real wages recieved by households
 nsben_hk  = nsben_h /pc #real social benefits recieved by the households
 npropinc_hk =  npropinc_h/pc # real property income (net) 
 b2_hk = b2_h_r/pc
 yd1_hk = yd1_h/pc  
 yd2a_hk = yd2a_h/pc
 yd2b_hk = yd2b_h/pc
```

```{r}
fnw_hk = fnw_h/pc #real net wealth of households
```

Variables for the estimation of the investment equation

```{r}
u=yk/(bd_nfc_k+equip_nfc_k) #Rate of capacity utilization
prate=s_nf/(stats::lag(bd_nfc,-1)+stats::lag(equip_nfc,-1)) # Profit rate
```

Labour market

```{r}
#empl = N/1000
#empl_ds = Nadj/1000
 wage = w_h_r/emp
wage1= w_h_r1/emp1
# wage = w_h_rk_ds /(empl_ds)
 Nu=(w_row_r - w_row_p)/wage
 Nf = Nu + emp
 prod = y/Nf
 LF = unemp + emp
 lfadj = unempadj + empadj
 ur = unemp/LF
 ur_ds = unempadj/lfadj
# yk = y/(py/100)
 prodk = yk/Nf
 yf = w + b2
 ws = w/yf
 ps = 1 - ws
# w = w_h_r + (W_row_r - W_row_p)
 fdis = yf/y
 urterm = ur - urs
 wageindex = wage/wage[22] #element from 2010Q3
 priv = pconk + ik + xk
 part = LF*1000/pop
 old_age_ratio = retired_pop/pop
 part1= LF*1000/(pop*(1-old_age_ratio))
 zaland_jesper = pop - LF*1000  
 nben_h = nsben_h + d8_h
```

```{r}
nl_check= (nl_h + nl_f + nl_nf + nl_g + nl_row) # needs to be fixed:
 fnl_check= (fnl_h + fnl_f + fnl_nf + fnl_g + fnl_row)  # this one is fine: because there is almost no connection between the real side and financial side at the moment
 check_np = np_row + np_h + np_f + np_nf + np_g # this one is fine:
 check_ctr = ctr_row + ctr_h + ctr_f + ctr_nf + ctr_g # this one is fine:
 check_invest = i - i_f - i_nf - i_g - i_h # this one is fine:
 check_iba = iba_h + iba_f + iba_nf + iba_g + iba_row  #needs to be fixed:
 check_tax = tax_g + tax_nf + tax_f + tax_h + tax_row # this one is fine:
  check_iba_rv = iba_h_rv + iba_f_rv + iba_nf_rv + iba_g_rv + iba_row_rv 
  check_iba_tr = iba_h_tr + iba_f_tr + iba_nf_tr + iba_g_tr + iba_row_tr
 check_eq = eq_h + (eq_f_a - eq_f_l) + (eq_nf_a - eq_nf_l) + eq_g + (eq_row_a - eq_row_l) # This one is now fixed
 check_eq_tr = eq_h_tr + Neq_nf_tr  + Neq_f_tr + eq_g_tr + Neq_row_tr + eq_g_tr
  check_l = l_h + l_f + l_nf + l_g + l_row # problematic after scenario 2
 check_l_tr = l_f_tr + l_h_tr + l_nf_tr + l_g_tr + l_row_tr
  check_ins = ins_h + ins_f + ins_nf + ins_g + ins_row # problematic after scenario 2
 check_ins_tr =  ins_f_tr  + ins_h_tr + ins_nf_tr + ins_g_tr + ins_row_tr
check_sec = sec_h + sec_f_a + sec_f_d + sec_nf + sec_g + sec_row ## needs to be fixed
```

Ved ik helt om nedenstående skal med

```{r}
#smpl 2005q2 2019q1

#equation eq41.ls emp = P(251) + P(252)*SD1 + P(253)*SD2 + P(254)*SD3
#series  emp_ds = emp - (P(251) + P(252)*SD1 + P(253)*SD2 + P(254)*SD3) + @mean(emp)

tobinq = (eq_nf_l) / (equip_nfc + bd_nfc)
```

```{r}
library(mFilter)
library(bimets)
library(knitr)
```

# Deseasonalyse

```{r}

## I_BD_H_K
reg_1 = lm(i_bd_h_k~ sd1+ sd2 + sd3)

i_bd_h_k_ds= reg_1$residuals + mean(i_bd_h_k)

i_bd_h_k_ds=ts(i_bd_h_k_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_10 =reg_1$coefficients[1]
alpha_10=ts(alpha_10,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_11=reg_1$coefficients[2]
alpha_11=ts(alpha_11,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_12=reg_1$coefficients[3]
alpha_12=ts(alpha_12,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_13=reg_1$coefficients[4]
alpha_13=ts(alpha_13,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_i_bd_h_k= mean(i_bd_h_k)
mean_i_bd_h_k=ts(mean_i_bd_h_k,start = c(2005,2), end = c(2020,1), frequency = 4 )

## I_EQUIP_H_K

reg_3 = lm(i_equip_h_k~ sd1+ sd2 + sd3)

i_equip_h_k_ds= reg_3$residuals + mean(i_equip_h_k)

i_equip_h_k_ds=ts(i_equip_h_k_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_20 =reg_3$coefficients[1]
alpha_20=ts(alpha_20,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_21=reg_3$coefficients[2]
alpha_21=ts(alpha_21,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_22=reg_3$coefficients[3]
alpha_22=ts(alpha_22,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_23=reg_3$coefficients[4]
alpha_23=ts(alpha_23,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_i_equip_h_k= mean(i_equip_h_k)
mean_i_equip_h_k=ts(mean_i_equip_h_k,start = c(2005,2), end = c(2020,1), frequency = 4 )

# i_bd_nfc_k_ds

reg_4 = lm(i_bd_nfc_k~ sd1+ sd2 + sd3)

i_bd_nfc_k_ds= reg_4$residuals + mean(i_bd_nfc_k)

i_bd_nfc_k_ds=ts(i_bd_nfc_k_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_30 =reg_4$coefficients[1]
alpha_30=ts(alpha_30,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_31=reg_4$coefficients[2]
alpha_31=ts(alpha_31,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_32=reg_4$coefficients[3]
alpha_32=ts(alpha_32,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_33=reg_4$coefficients[4]
alpha_33=ts(alpha_33,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_i_bd_nfc_k= mean(i_bd_nfc_k)
mean_i_bd_nfc_k=ts(mean_i_bd_nfc_k,start = c(2005,2), end = c(2020,1), frequency = 4 )

## u_ds
reg_5 = lm(u~ sd1+ sd2 + sd3)

u_ds= reg_5$residuals + mean(u)

u_ds=ts(u_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## ps_ds
reg_6 = lm(ps~ sd1+ sd2 + sd3)

ps_ds= reg_6$residuals + mean(ps)

ps_ds=ts(ps_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## i_equip_nfc_k_ds
reg_7 = lm(i_equip_nfc_k~ sd1+ sd2 + sd3)

i_equip_nfc_k_ds= reg_7$residuals + mean(i_equip_nfc_k)

i_equip_nfc_k_ds=ts(i_equip_nfc_k_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_40 =reg_7$coefficients[1]
alpha_40=ts(alpha_40,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_41=reg_7$coefficients[2]
alpha_41=ts(alpha_41,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_42=reg_7$coefficients[3]
alpha_42=ts(alpha_42,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_43=reg_7$coefficients[4]
alpha_43=ts(alpha_43,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_i_equip_nfc_k= mean(i_equip_nfc_k)
mean_i_equip_nfc_k=ts(mean_i_equip_nfc_k,start = c(2005,2), end = c(2020,1), frequency = 4 )

## pconk_ds

reg_8 = lm(pconk~ sd1+ sd2 + sd3)

pconk_ds= reg_8$residuals + mean(pconk)

pconk_ds=ts(pconk_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_00 =reg_8$coefficients[1]
alpha_00=ts(alpha_00,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_01=reg_8$coefficients[2]
alpha_01=ts(alpha_01,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_02=reg_8$coefficients[3]
alpha_02=ts(alpha_02,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_03=reg_8$coefficients[4]
alpha_03=ts(alpha_03,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_pconk= mean(pconk)
mean_pconk=ts(mean_pconk,start = c(2005,2), end = c(2020,1), frequency = 4 )

## B2_HK_ds

reg_9 = lm(b2_hk~ sd1+ sd2 + sd3)

b2_hk_ds= reg_9$residuals + mean(b2_hk)

b2_hk_ds=ts(b2_hk_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## yd1_hk

reg_10 = lm(yd1_hk~ sd1+ sd2 + sd3)

yd1_hk_ds= reg_10$residuals + mean(yd1_hk)

yd1_hk_ds=ts(yd1_hk_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_100 =reg_10$coefficients[1]
alpha_100=ts(alpha_100,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_101=reg_10$coefficients[2]
alpha_101=ts(alpha_101,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_102=reg_10$coefficients[3]
alpha_102=ts(alpha_102,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_103=reg_10$coefficients[4]
alpha_103=ts(alpha_103,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_yd1_hk= mean(yd1_hk)
mean_yd1_hk=ts(mean_yd1_hk,start = c(2005,2), end = c(2020,1), frequency = 4)

## yd2a_hk

reg_11 = lm(yd2a_hk~ sd1+ sd2 + sd3)

yd2a_hk_ds= reg_11$residuals + mean(yd2a_hk)

yd2a_hk_ds=ts(yd2a_hk_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_110 =reg_11$coefficients[1]
alpha_110=ts(alpha_110,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_111=reg_11$coefficients[2]
alpha_111=ts(alpha_111,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_112=reg_11$coefficients[3]
alpha_112=ts(alpha_112,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_113=reg_11$coefficients[4]
alpha_113=ts(alpha_113,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_yd2a_hk= mean(yd2a_hk)
mean_yd2a_hk=ts(mean_yd2a_hk,start = c(2005,2), end = c(2020,1), frequency = 4)

##w_h_rk

reg_12 = lm(w_h_rk~ sd1+ sd2 + sd3)

w_h_rk_ds= reg_12$residuals + mean(w_h_rk)

w_h_rk_ds=ts(w_h_rk_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## wage

reg_13 = lm(wage~ sd1+ sd2 + sd3)

wage_ds= reg_13$residuals + mean(wage)

wage_ds=ts(wage_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_50 =reg_13$coefficients[1]
alpha_50=ts(alpha_50,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_51=reg_13$coefficients[2]
alpha_51=ts(alpha_51,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_52=reg_13$coefficients[3]
alpha_52=ts(alpha_52,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_53=reg_13$coefficients[4]
alpha_53=ts(alpha_53,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_wage= mean(wage)
mean_wage=ts(mean_wage,start = c(2005,2), end = c(2020,1), frequency = 4)

## pc

reg_14 = lm(pc~ sd1+ sd2 + sd3)

pc_ds= reg_14$residuals + mean(pc)

pc_ds=ts(pc_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_60 =reg_14$coefficients[1]
alpha_60=ts(alpha_60,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_61=reg_14$coefficients[2]
alpha_61=ts(alpha_61,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_62=reg_14$coefficients[3]
alpha_62=ts(alpha_62,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_63=reg_14$coefficients[4]
alpha_63=ts(alpha_63,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_pc= mean(pc)
mean_pc=ts(mean_pc,start = c(2005,2), end = c(2020,1), frequency = 4)

## pm

reg_15 = lm(pm~ sd1+ sd2 + sd3)

pm_ds= reg_15$residuals + mean(pm)

pm_ds=ts(pm_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

### ur

reg_16 = lm(ur~ sd1+ sd2 + sd3)

ur_ds= reg_16$residuals + mean(ur)

ur_ds=ts(ur_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_140 =reg_16$coefficients[1]
alpha_140=ts(alpha_140,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_141=reg_16$coefficients[2]
alpha_141=ts(alpha_141,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_142=reg_16$coefficients[3]
alpha_142=ts(alpha_142,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_143=reg_16$coefficients[4]
alpha_143=ts(alpha_143,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_ur= mean(ur)
mean_ur=ts(mean_ur,start = c(2005,2), end = c(2020,1), frequency = 4)

## yk

reg_17 = lm(yk~ sd1+ sd2 + sd3)

yk_ds= reg_17$residuals + mean(yk)

yk_ds=ts(yk_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_90 =reg_17$coefficients[1]
alpha_90=ts(alpha_90,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_91=reg_17$coefficients[2]
alpha_91=ts(alpha_91,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_92=reg_17$coefficients[3]
alpha_92=ts(alpha_92,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_93=reg_17$coefficients[4]
alpha_93=ts(alpha_93,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_yk= mean(yk)
mean_yk=ts(mean_yk,start = c(2005,2), end = c(2020,1), frequency = 4)

## wageindex

reg_18 = lm(wageindex~ sd1+ sd2 + sd3)

wageindex_ds= reg_18$residuals + mean(wageindex)

wageindex_ds=ts(wageindex_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## priv

reg_19 = lm(priv~ sd1+ sd2 + sd3)

priv_ds= reg_19$residuals + mean(priv)

priv_ds=ts(priv_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## mk

reg_20 = lm(mk~ sd1+ sd2 + sd3)

mk_ds= reg_20$residuals + mean(mk)

mk_ds=ts(mk_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_70 =reg_20$coefficients[1]
alpha_70=ts(alpha_70,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_71=reg_20$coefficients[2]
alpha_71=ts(alpha_71,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_72=reg_20$coefficients[3]
alpha_72=ts(alpha_72,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_73=reg_20$coefficients[4]
alpha_73=ts(alpha_73,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_mk= mean(mk)
mean_mk=ts(mean_mk,start = c(2005,2), end = c(2020,1), frequency = 4)

## prodk

reg_21 = lm(prodk~ sd1+ sd2 + sd3)

prodk_ds= reg_21$residuals + mean(prodk)

prodk_ds=ts(prodk_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## prate

##This one is shorter but not sure it also is in the Eviews code

reg_22 = lm(prate~ sd1[-1]+ sd2[-1] + sd3[-1])

prate_ds= reg_22$residuals + mean(prate)

prate_ds=ts(prate_ds,start = c(2005,3), end = c(2020,1), frequency = 4 )

alpha_120 =reg_22$coefficients[1]
alpha_120=ts(alpha_120,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_121=reg_22$coefficients[2]
alpha_121=ts(alpha_121,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_122=reg_22$coefficients[3]
alpha_122=ts(alpha_122,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_123=reg_22$coefficients[4]
alpha_123=ts(alpha_123,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_prate= mean(prate)
mean_prate=ts(mean_prate,start = c(2005,2), end = c(2020,1), frequency = 4)

## bd_nfc_k

reg_23 = lm(bd_nfc_k~ sd1+ sd2 + sd3)

bd_nfc_k_ds= reg_23$residuals + mean(bd_nfc_k)

bd_nfc_k_ds=ts(bd_nfc_k_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## xk

reg_24 = lm(xk~ sd1+ sd2 + sd3)

xk_ds= reg_24$residuals + mean(xk)

xk_ds=ts(xk_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_80 =reg_24$coefficients[1]
alpha_80=ts(alpha_80,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_81=reg_24$coefficients[2]
alpha_81=ts(alpha_81,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_82=reg_24$coefficients[3]
alpha_82=ts(alpha_82,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_83=reg_24$coefficients[4]
alpha_83=ts(alpha_83,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_xk= mean(xk)
mean_xk=ts(mean_xk,start = c(2005,2), end = c(2020,1), frequency = 4)
## rer

reg_25 = lm(rer~ sd1+ sd2 + sd3)

rer_ds= reg_25$residuals + mean(rer)

rer_ds=ts(rer_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## prod

reg_27 = lm(prod~ sd1+ sd2 + sd3)

prod_ds= reg_27$residuals + mean(prod)

prod_ds=ts(prod_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

## ps

reg_28 = lm(ps~ sd1+ sd2 + sd3)

ps_ds= reg_28$residuals + mean(ps)

ps_ds=ts(ps_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_130 =reg_28$coefficients[1]
alpha_130=ts(alpha_130,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_131=reg_28$coefficients[2]
alpha_131=ts(alpha_131,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_132=reg_28$coefficients[3]
alpha_132=ts(alpha_132,start = c(2005,2), end = c(2020,1), frequency = 4 )

alpha_133=reg_28$coefficients[4]
alpha_133=ts(alpha_133,start = c(2005,2), end = c(2020,1), frequency = 4 )

mean_ps= mean(ps)
mean_ps=ts(mean_ps,start = c(2005,2), end = c(2020,1), frequency = 4)

## y

reg_29 = lm(y~ sd1+ sd2 + sd3)

y_ds= reg_29$residuals + mean(y)

y_ds=ts(y_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )
```

```{r}
## Disse skal komme efter deseason

rw_ds = wage_ds/pc_ds #real wage rate based on deseasonalized variables
reprate_bd_h = (mean(delta_bd_h)*stats::lag(bd_h_k,-1))/i_bd_h_k_ds # replacement rate

#New wage equations
inflation= (pc/TSLAG(pc,4)) - 1

inflation_t= (pc/TSLAG(pc,4)) - 1
inflation_t = ts(c(0,0,0, inflation_t), start = c(2005,3), frequency = 4 )

wage_ds_t= sd1* wage_ds*(1+inflation_t) + sd2* TSLAG(wage_ds,1)*(1+TSLAG(inflation_t,1)) + sd3* TSLAG(wage_ds,2)*(1+TSLAG(inflation_t,2)) + sd4* TSLAG(wage_ds,3)*(1+TSLAG(inflation_t,3))

inflation_tt= inflation_t

# New price equations

pg_adj = log(pg) - (0.840852289293*log(pc) - 0.0401650251402)

p_bd_adj = log(p_bd) - (0.988897636487*log(pc) +
0.0286611434831)

p_equip_adj = log(p_equip) -( 0.304376220161*log(pc) +
0.0160136953595 )

px_adj= log(px) - (0.638216896739*log(pc) - 0.012650117362)
```

# Extra variables created to run the model:

```{r}
eq_nf_lx =   eq_nf_a_nf + eq_h_nf + eq_row_a_nf + eq_f_a_nf + eq_g_nf
yh2a = propinc_r_h + b2_h_r
yh2b = - propinc_p_h
aa_check = TSLAG(eq_h_f,1)*TSLAG(divd,1)
insu_h = TSLAG(ins_h,1)*TSLAG(insu,1) + ins_h_adj
yd_h1 = yh_h + tax_h + nben_h
sec_h_tr_g=sec_h_tr
eq_h_test = eq_h - (eq_h_nf + eq_h_f + eq_h_row)
lev_h = l_h/yd_h
fa_h = iba_h + sec_h + eq_h + ins_h
sk= pconk + ik + gk + xk
s = pcon + g + i + x
w_nf = w_h_r +  (w_row_r - w_row_p)
int_nf = (TSLAG(idep,1)*TSLAG(iba_nf,1) + TSLAG(ibd,1)*TSLAG(sec_nf,1) + TSLAG(iloan,1)*TSLAG(l_nf,1)) 
insu_nf =  TSLAG(ins_nf,1)*TSLAG(insu,1)
npropinc_nf = int_nf + int_nf_adj + div_nf + insu_nf + ins_nf_adj + (res_r_nf ) - (res_p_nf)
eq_nf_ax =  eq_nf_a_nf + eq_nf_a_f + eq_nf_a_row
eq_nf_a_test = eq_nf_a - eq_nf_ax
eq_f_lx =   eq_nf_a_f + eq_h_f + eq_row_a_f + eq_f_a_f + eq_g_f
eq_nf_l_test = eq_nf_l - eq_nf_lx 
fnw_nfk = fnw_nf/pc
int_f =(TSLAG(idep,1)*TSLAG(iba_f,1) + TSLAG(iloan,1)*TSLAG(l_f,1) + TSLAG(ibd,1)*TSLAG(sec_f_d,1) + TSLAG(iboa,1)*TSLAG(sec_f_a,1)) 
insu_f = TSLAG(ins_f,1)*TSLAG(insu,1)
npropinc_f = int_f + int_f_adj +  div_f + insu_f + ins_f_adj + (res_r_f) - (res_p_f)
nben_f = d8_f 
eq_f_ax =  eq_f_a_nf + eq_f_a_f + eq_f_a_row
eq_f_l_test = eq_f_l - (eq_f_l_f + eq_h_f + eq_row_a_f + eq_nf_a_f + eq_g_f)
sec_f_d_tr_nf=-sec_nf_tr
sec_f_d_tr_g=sec_f_d_tr-sec_f_d_tr_nf
fnw_fk = fnw_f/pc
int_g = (TSLAG(idep,1)*TSLAG(iba_g,1) + TSLAG(iloan,1)*TSLAG(l_g,1) + TSLAG(ibd,1)*TSLAG(sec_g,1))
insu_g = TSLAG(ins_g,1)*TSLAG(insu,1)
npropinc_g = int_g + int_g_adj + div_g + insu_g + ins_g_adj + (res_r_g) - ( res_p_g)
nben_row = scon_row_r - scon_row_p + sben_row_r - sben_row_p
nben_g = - (nben_row + nsben_h + nben_f)
eq_g_ax =  eq_g_row + eq_g_f + eq_g_nf
eq_g_test = eq_g - (eq_g_nf + eq_g_f + eq_g_row)
fnw_gk = fnw_g/pc
nx = x - m
cab =-nl_row
fab = (fnl_row)
bop = cab + fab
eq_row_ax =  eq_row_a_nf + eq_row_a_f
eq_row_a_test = eq_row_a - (eq_row_a_nf + eq_row_a_f)
eq_row_lx = eq_nf_a_row + eq_h_row + eq_f_a_row + eq_g_row
eq_row_l_test = eq_row_l - (eq_nf_a_row + eq_f_a_row + eq_g_row + eq_h_row)
fnw_rowk = fnw_row/pc
mkp = pc_ds / (wage_ds/prodk_ds + pm_ds)
acc_rate = ik / (bd_h_k+equip_h_k+bd_nfc_k+equip_nfc_k+bd_fc_k+equip_fc_k+bd_g_k+equip_g_k)
yh1 = w_h_r +  nsben_h + oth_h
yd_hk_ds = yd1_hk_ds +yd2a_hk_ds + yd2b_hk
propinc_p_nf = int_p_nf + div_p_nf + ins_p_nf + res_p_nf
propinc_r_nf = int_r_nf + div_r_nf + ins_r_nf + res_r_nf
```

# Shock variables

```{r}
## Creating the new shock variables
g_shock= g*1.02
prodk_ds_shock = prodk_ds
max_dp_shock= obs_dp_q
stat_reg_shock = ts(0.02, start = c(2005, 2), end =  c(2020,1), frequency = 4)

## Shock_1
pf_shock= pf+pf*0.1
pm_ds_shock= pm_ds+pm_ds*0.1

## Shock 2
iloan_shock= iloan + 0.03
idep_shock= idep + 0.03
ibd_shock= ibd + 0.03
iboa_shock= iboa + 0.03

## Shock 3

tax_rate1_shock= ts(c(0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369,0.369369), start = c(2005, 2), end =  c(2020,1), frequency = 4)

#### Shock variables 
shock_max_dp_model=ts(0,start = c(2005,2), end = c(2020,1), frequency = 4)
shock_g_model=ts(0,start = c(2005,2), end = c(2020,1), frequency = 4)
shock_reg_model=ts(0,start = c(2005,2), end = c(2020,1), frequency = 4)
shock_part_model=ts(0,start = c(2005,2), end = c(2020,1), frequency = 4)
shock_1_model=ts(0,start = c(2005,2), end = c(2020,1), frequency = 4)
shock_2_model=ts(0,start = c(2005,2), end = c(2020,1), frequency = 4)
shock_3_model=ts(0,start = c(2005,2), end = c(2020,1), frequency = 4)

#### Sensitivity analysis variables 
sens1=ts(0,start = c(2005,2), end = c(2020,1), frequency = 4)

```

# New variables for LF model

## Wage_trim

As some of the variables are yearly and stock variables we devide them by 4

```{r}
tyd_q= tyd/4
tydd_q= tydd/4
```

```{r, fig.width=10, fig.height=7}

w_h_r_trim1 = w_h_r
wage_trim1 = w_h_r_trim1/emp
## Deseasonalize

reg_30 = lm(wage_trim1~ sd1+ sd2 + sd3)

wage_trim1_ds= reg_30$residuals + mean(wage_trim1)

wage_trim1_ds=ts(wage_trim1_ds,start = c(2005,2), end = c(2020,1), frequency = 4 )

plot(wage_trim1_ds)

```

Vi kan se at den store forskel ligger i om vi bruger unemployment data fra vores model eller ADAM.

## Komp_r bruges

```{r}

library(magrittr)

#Yearly change in wage: (as max dagpenge kun ændres årligt?)

wage_growth=TSDELTA(wageindex,4)

wage_growth_1q=as.data.frame(as.numeric(wage_growth)[c(1,4,8,12,16,20,24,28,32,36,40,44,48,52,56)]) # Jeg bruger kvartal 2 fra 2006 som udviklling i 2006

wage_growth_1q %<>% 
  rename(Q1='as.numeric(wage_growth)[c(1, 4, 8, 12, 16, 20, 24, 28, 32, 36, 40, 44, 48, 52, 56)]')
```

De nævner selv at den løn stigning der regnes med er fratrukket arbejdsgiveradministrerede pensionsordninger

```{r}
#Udregninger af maximal dagpenge beløb er lavet i "udregning af dagpenge"

komp_r_ny = obs_dp_q/ wage_trim1
plot(komp_r_ny)

komp_r_ny1 = wage_dp_q/ wage_trim1
plot(komp_r_ny1)
```

## Udregning af dagpenge total

```{r}
adam_dp_person= (tyd/uld)/4
```

```{r}
## Måske brug IV-regression men passer nok ikke data godt så. 
total_dp_reg=lm(adam_dp_person ~ 0 +max_dp_quarter)

mean_dp_regul= 0.9507*obs_dp_q
nben_h_un_obs=mean_dp_regul*unemp*kuld
komp_r_obs=mean_dp_regul/wage_trim1
nben_h_rest_obs= nben_h - nben_h_un_obs

plot(komp_r_obs)
```

```{r}
mean_dp_wage= 0.9507*wage_dp_q
total_dp_wage= mean_dp_wage*unemp*kuld
komp_r_wage=mean_dp_wage/wage_trim1
```

```{r}
tilpas_pr = window(tilpas_pr_q, start= c(2006,1))
stat_tilpas = window(stat_tilpas_q, start= c(2006,1))
stat_reg = window(stat_reg_q, start= c(2006,1))
```

```{r}
max_dp_test= TSLAG(obs_dp_q,4) +(stat_reg_q+stat_tilpas_q) * TSLAG(obs_dp_q,4) 
```

# Txt. for model

```{r}

S_model.txt="MODEL

COMMENT> Max_dp

COMMENT> Lavet for at kunne støde til g 
IDENTITY> max_dp
EQ> max_dp = TSLAG(max_dp,4)+ (stat_reg+stat_tilpas)*TSLAG(max_dp,4)
IF> shock_max_dp_model <= 0 & SD1 >= 1
IDENTITY> max_dp
EQ> max_dp = TSLAG(max_dp,1)
IF>  SD2 >= 1
IDENTITY> max_dp
EQ> max_dp = TSLAG(max_dp,2)
IF>  SD3 >= 1
IDENTITY> max_dp
EQ> max_dp = TSLAG(max_dp,3)
IF>  SD4 >= 1
IDENTITY> max_dp
EQ> max_dp = TSLAG(max_dp,4)+ TSDELTA(TSLAG(wage,8),4)/TSLAG(wage,12)*TSLAG(max_dp,4)
IF> shock_max_dp_model > 0 & SD1 >= 1

COMMENT> Capital Stock with switches
IDENTITY> tax_rate1
EQ> tax_rate1 = tax_rate1*1
IF> shock_3_model <= 0
IDENTITY> tax_rate1
EQ> tax_rate1 = tax_rate1_shock
IF> shock_3_model > 0

COMMENT>  
IDENTITY> stat_tilpas
EQ> stat_tilpas = tilpas_pr
IF> tilpas_pr < 0 & SD1 >= 1
IDENTITY> stat_tilpas
EQ> stat_tilpas = 0
IF> tilpas_pr >= 0 & tilpas_pr <=0.003 & SD1 >= 1
IDENTITY> stat_tilpas
EQ> stat_tilpas = tilpas_pr - 0.003
IF> tilpas_pr >=0.003 & SD1 >= 1
IDENTITY> stat_tilpas
EQ> stat_tilpas = TSLAG(stat_tilpas,1)
IF>  SD2 >= 1
IDENTITY> stat_tilpas
EQ> stat_tilpas = TSLAG(stat_tilpas,2)
IF>  SD3 >= 1
IDENTITY> stat_tilpas
EQ> stat_tilpas = TSLAG(stat_tilpas,3)
IF>  SD4 >= 1

COMMENT> Lavet for at kunne støde til g 
IDENTITY> tilpas_pr
EQ> tilpas_pr = TSDELTA(TSLAG(wage,8),4)/TSLAG(wage,12) - 0.02
IF>  SD1 >= 1
IDENTITY> tilpas_pr
EQ> tilpas_pr = TSLAG(tilpas_pr,1)
IF>  SD2 >= 1
IDENTITY> tilpas_pr
EQ> tilpas_pr = TSLAG(tilpas_pr,2)
IF>  SD3 >= 1
IDENTITY> tilpas_pr
EQ> tilpas_pr = TSLAG(tilpas_pr,3)
IF>  SD4 >= 1

COMMENT> Pricing Nexus

COMMENT> 
IDENTITY> pg
EQ> LOG(pg) = 0.840852289293*LOG(pc) - 0.0401650251402 +
pg_adj

COMMENT> 
IDENTITY> p_bd
EQ> LOG(p_bd) = 0.988897636487*LOG(pc) + 0.0286611434831 +
p_bd_adj

COMMENT> 
IDENTITY> p_equip
EQ> LOG(p_equip) = 0.304376220161*LOG(pc) + 0.0160136953595
+ p_equip_adj

COMMENT> 
IDENTITY> px
EQ> LOG(px) = 0.638216896739*LOG(pc) - 0.012650117362 +
px_adj

COMMENT> Households

COMMENT> 
IDENTITY> int_r_h
EQ> int_r_h = TSLAG(idep,1)*TSLAG(iba_h,1)  + TSLAG(ibd,1)*TSLAG(sec_h,1)  + int_r_h_adj

COMMENT> 
IDENTITY> int_p_h
EQ> int_p_h =  - TSLAG(iloan,1)*TSLAG(l_h,1)  + int_p_h_adj

COMMENT> 
IDENTITY> div_h
EQ> div_h = d_20124*((TSLAG(eq_h_nf,1) + TSLAG(eq_h_f,1))*TSLAG(divd,1) + TSLAG(eq_h_row,1)*TSLAG(diva,1)) + (1-d_20124)*(TSLAG(eq_h,1)*TSLAG(divd,1)) +  div_h_adj

COMMENT> 
IDENTITY> aa_check
EQ> aa_check = TSLAG(eq_h_f,1)*TSLAG(divd,1)

COMMENT> 
IDENTITY> insu_h
EQ> insu_h = TSLAG(ins_h,1)*TSLAG(insu,1) + ins_h_adj

COMMENT>  property income H
IDENTITY> propinc_r_h
EQ> propinc_r_h = int_r_h + div_h + insu_h + res_r_h

COMMENT> ' property expenditure H
IDENTITY> propinc_p_h
EQ> propinc_p_h = int_p_h + res_p_h

COMMENT> 
IDENTITY> npropinc_h
EQ> npropinc_h = propinc_r_h - propinc_p_h

COMMENT> Gross income H
IDENTITY> yh_h
EQ> yh_h = w_h_r + b2_h_r + (npropinc_h )

COMMENT> 
IDENTITY> yh2a
EQ> yh2a = propinc_r_h + b2_h_r

COMMENT> 
IDENTITY> yh2b
EQ> yh2b = - propinc_p_h

COMMENT> 
IDENTITY> yh1
EQ> yh1 = w_h_r +  nsben_h + oth_h

COMMENT> 
IDENTITY> w_h_r
EQ> w_h_r = wage * emp

COMMENT> ' property income part of the disposable income (capitalists)
IDENTITY> yd2a_h
EQ> yd2a_h =(1 - tax_rate2)*(yh2a)

COMMENT> 
IDENTITY> yd2b_h
EQ> yd2b_h = (1 - tax_rate2)*(yh2b)

COMMENT> age and benefits part of the disposable income (workers) / tax_h_adj is added here because of the relative size versus yh2
IDENTITY> yd1_h
EQ> yd1_h =(1 - tax_rate1)*(yh1) - tax_h_adj

COMMENT> 
IDENTITY> tax_h
EQ> tax_h = -(tax_rate2*(yh2a+yh2b) +  tax_rate1*yh1 +  tax_h_adj)

COMMENT> 
IDENTITY> yd1_hk
EQ> yd1_hk = yd1_h/pc

COMMENT> 
IDENTITY> yd2a_hk
EQ> yd2a_hk = yd2a_h/pc

COMMENT> 
IDENTITY> yd2b_hk
EQ> yd2b_hk = yd2b_h/pc

COMMENT> Disposable income
IDENTITY> yd_h
EQ> yd_h = yd1_h + yd2a_h + yd2b_h

COMMENT> 
IDENTITY> nsben_h
EQ> nsben_h = nben_h - d8_h

COMMENT> Disposable income (redundant for now)
IDENTITY> yd_h1
EQ> yd_h1 = yh_h + tax_h + nben_h

COMMENT> 
IDENTITY> yd_hk
EQ> yd_hk = yd1_hk + yd2a_hk + yd2b_hk

COMMENT> 
IDENTITY> yd_hk_ds
EQ> yd_hk_ds = yd1_hk_ds + yd2a_hk_ds + yd2b_hk

COMMENT> 
IDENTITY> yd1_hk_ds
EQ> yd1_hk_ds = yd1_hk - (alpha_100 + alpha_101*SD1 + alpha_102*SD2 + alpha_103*SD3) + mean_yd1_hk

COMMENT> 
IDENTITY> yd2a_hk_ds
EQ> yd2a_hk_ds  = yd2a_hk  - (alpha_110 + alpha_111*SD1 + alpha_112*SD2 + alpha_113*SD3) + mean_yd2a_hk

COMMENT> Savings H
IDENTITY> s_h
EQ> s_h = yd_h - pcon + d8_h

COMMENT> Sectoral balance
IDENTITY> nl_h
EQ> nl_h = s_h - i_h - np_h + ctr_h

COMMENT> 
BEHAVIORAL> pconk_ds
TSRANGE 2006 2 2020 1
EQ> TSDELTALOG(pconk_ds) = C_1 + C_2*TSLAG(LOG(pconk_ds),1) + C_3*TSLAG(LOG(yd1_hk_ds),1) + C_4*LOG(TSLAG(yd2a_hk_ds,1) + TSLAG(yd2b_hk,1)) + C_5*TSLAG(LOG(fnw_hk),2) + C_6*TSDELTALOG(TSLAG(pconk_ds,2)) + C_7*TSDELTALOG(TSLAG(pconk_ds,3)) + C_8*TSDELTALOG(yd1_hk_ds)  + C_9*TSDELTALOG(yd2a_hk_ds + yd2b_hk) + C_10*d_2008q4 + C_11*d_2018q2 + C_12*d_2020q1 + C_13* time
COEFF> C_1 C_2 C_3 C_4 C_5 C_6 C_7 C_8 C_9 C_10 C_11 C_12 C_13
STORE> coe(1)

COMMENT> 
IDENTITY> pconk
EQ> pconk = pconk_ds + alpha_00 + alpha_01*SD1 + alpha_02*SD2 + alpha_03*SD3 - mean_pconk

COMMENT> nominal private consumption
IDENTITY> pcon
EQ> pcon = pconk*pc

COMMENT>  
BEHAVIORAL> i_bd_h_k_ds
TSRANGE 2006 3 2020 1
EQ> LOG(i_bd_h_k_ds) = P_620 + 
P_621*TSDELTALOG(TSLAG(i_bd_h_k_ds,1)/TSLAG(bd_h_k,2))+ 
P_622*TSDELTALOG(TSLAG(i_bd_h_k_ds,3)/TSLAG(bd_h_k,4))+ 
P_623*TSDELTALOG(TSLAG(p_bd,1)/TSLAG(pi_1,1)) + 
P_624*TSDELTALOG(TSLAG(p_bd,2)/TSLAG(pi_1,2)) + 
P_625*TSDELTALOG(TSLAG(yd_hk_ds,2)/TSLAG(bd_h_k,3))+ 
P_626*TSDELTALOG(TSLAG(-l_h,1)/TSLAG(bd_h,2)) + 
P_627*LOG(TSLAG(i_bd_h_k_ds,1)/TSLAG(bd_h_k,2)) + 
P_628*LOG(TSLAG(yd_hk_ds,1)/TSLAG(bd_h_k,2)) + 
P_629*LOG(TSLAG(p_bd,1)/TSLAG(pi_1,1)) +
P_630*LOG(TSLAG(-l_h,1)/TSLAG(bd_h,2)) + 
P_631*d_2006q4 + 
P_632*d_2014q4 + 
alpha_1*LOG(TSLAG(bd_h_k,1)) + 
alpha_2* TSLAG(LOG(i_bd_h_k_ds),1) + 
alpha_3* TSLAG(-LOG(bd_h_k),2)
COEFF> P_620 P_621 P_622 P_623 P_624 P_625 P_626 P_627 P_628 P_629 P_630 P_631 P_632 alpha_1 alpha_2 alpha_3
RESTRICT> alpha_1 = 1
alpha_2 = 1
alpha_3 = 1
STORE> coe(2)

COMMENT> 
IDENTITY> i_bd_h_k
EQ> i_bd_h_k = i_bd_h_k_ds + (alpha_10 + alpha_11*SD1 + alpha_12*SD2 + alpha_13*SD3) - mean_i_bd_h_k

COMMENT> 
BEHAVIORAL> i_equip_h_k_ds
TSRANGE 2006 2 2020 1
EQ> LOG(i_equip_h_k_ds) = p_636 + p_637 * TSDELTALOG(TSLAG(i_equip_h_k_ds,1)/TSLAG(equip_h_k,2)) + p_638 * TSDELTALOG(TSLAG(i_equip_h_k_ds,2)/TSLAG(equip_h_k,3)) + p_639*TSDELTALOG(yd_hk_ds/TSLAG(equip_h_k,1)) + p_640*dummy_4 + alpha_1* TSLAG(LOG(i_equip_h_k_ds),1) + alpha_2* TSLAG(LOG(equip_h_k),1) + alpha_3* TSLAG(-LOG(equip_h_k),2)
COEFF> p_636 p_637 p_638 p_639 p_640 alpha_1 alpha_2 alpha_3
RESTRICT> alpha_1 = 1
alpha_2 = 1
alpha_3 = 1 
STORE> coe(3)

COMMENT> 
IDENTITY> i_equip_h_k
EQ>  i_equip_h_k = i_equip_h_k_ds + alpha_20 + alpha_21*SD1 + alpha_22*SD2 + alpha_23*SD3 - mean_i_equip_h_k

COMMENT> tilføjet
IDENTITY> nben_h
EQ>  nben_h = nben_h_rest + nben_h_un

COMMENT> tilføjet
IDENTITY> komp_r
EQ>  komp_r = (dp_person/wage_trim)

COMMENT> tilføjet
IDENTITY> nben_h_un
EQ>  nben_h_un = dp_person*unemp*kuld

COMMENT> tilføjet
IDENTITY> dp_person
EQ>  dp_person = 0.9507*max_dp
IF>  sens1 <= 0
IDENTITY> dp_person
EQ> dp_person = 0.85*max_dp
IF>  sens1 >= 1

COMMENT> Sensitivitets analyse
IDENTITY> sens1
EQ> sens1 = sens1*1

COMMENT> tilføjet
IDENTITY> wage_trim
EQ>  wage_trim = w_h_r_trim/emp

COMMENT> tilføjet
IDENTITY> w_h_r_trim
EQ>  w_h_r_trim = w_h_r

COMMENT> tilføjet
BEHAVIORAL> nben_h_rest
TSRANGE 2005 4 2020 1
EQ> TSDELTALOG(nben_h_rest) = P_671 + P_672*TSDELTALOG(zaland_jesper) + P_673*TSDELTA(unemp) + P_674*TSDELTA(TSLAG(unemp,1)) +   P_675*LOG(TSLAG(nben_h_rest,1)) + P_676*TSLAG(unemp,1) + P_677*LOG(TSLAG(zaland_jesper,1)) 
COEFF> P_671 P_672 P_673 P_674 P_675 P_676 P_677 
STORE> coe(4)

COMMENT> Financial side

COMMENT> 
IDENTITY> l_h_tr
EQ> l_h_tr= l_h_tr_ratio * (-yd_hk_ds*pc)

COMMENT> Mangler stadig at få tilpasset til ik diff
BEHAVIORAL> l_h_tr_ratio
TSRANGE 2006 2 2020 1
EQ> TSDELTA(l_h_tr_ratio) = p_646 
+ p_641*TSDELTA(TSLAG(l_h_tr_ratio,2))
+ p_642*TSDELTA(iloan) 
+ p_643*TSDELTALOG(TSLAG(i_bd_h_k_ds,3)/TSLAG(yd_hk_ds,3)) 
+ p_644* TSLAG(l_h_tr_ratio,1) 
+ p_645*LOG(TSLAG(-l_h,2)/(TSLAG(yd_hk_ds,2)*TSLAG(pc,2)))
+ p_647*d_2007q34 
+ p_648*time
COEFF> p_646 p_641 p_642 p_643 p_644 p_645 p_647 p_648
STORE>  coe(5)

COMMENT> 
IDENTITY> iba_h
EQ> iba_h= TSLAG(iba_h,1) + iba_h_tr + iba_h_rvx

COMMENT> For the time being we take sec_h_tr as exogenous. It is assumed that households' holdings of securities are entirely issued by the government.
IDENTITY> sec_h
EQ> sec_h= TSLAG(sec_h,1) + sec_h_tr + sec_h_rvx

COMMENT> 
IDENTITY> sec_h_tr_g
EQ> sec_h_tr_g=sec_h_tr

COMMENT> 
IDENTITY> l_h
EQ> l_h= TSLAG(l_h,1) + l_h_tr + l_h_rvx

COMMENT> 
IDENTITY> eq_h
EQ> eq_h= eq_h_ratio*(TSLAG(eq_h,1) +TSLAG(sec_h,1)+  TSLAG(iba_h,1)) + eq_h_rvx

COMMENT> 
BEHAVIORAL> eq_h_ratio
TSRANGE 2006 1 2020 1
EQ> TSDELTA(eq_h_ratio) =  p_770  
+ p_771* TSDELTA(TSLAG(ibd,1))
+ p_772* TSDELTA(TSLAG(ratio_in_eq_h,1))
+ p_773* TSLAG(eq_h_ratio,1)
+ p_774* TSLAG(ibd,1) 
+ p_775* TSLAG(ratio_in_eq_h,2)
COEFF> p_770 p_771 p_772 p_773 p_774 p_775
STORE> coe(6)

COMMENT> 
IDENTITY> eq_h_tr
EQ> eq_h_tr = TSDELTA(eq_h) - eq_h_rvx

COMMENT> 
IDENTITY> iba_h_tr
EQ> iba_h_tr = nl_h - l_h_tr - eq_h_tr  - sec_h_tr - ins_h_tr + nl_h_adj

COMMENT> 
IDENTITY> eq_h_nf
EQ> eq_h_nf = d_20124*alpha_nf *eq_h

COMMENT> 
IDENTITY> eq_h_f
EQ> eq_h_f = d_20124*alpha_f *eq_h

COMMENT> 
IDENTITY> eq_h_row
EQ> eq_h_row = d_20124*alpha_row*eq_h

COMMENT> 
IDENTITY> eq_h_nf_tr
EQ> eq_h_nf_tr = TSDELTA(eq_h_nf) - eq_h_nf_rvx

COMMENT> 
IDENTITY> eq_h_f_tr
EQ> eq_h_f_tr = TSDELTA(eq_h_f) - eq_h_f_rvx

COMMENT> 
IDENTITY> eq_h_row_tr
EQ> eq_h_row_tr = TSDELTA(eq_h_row) - eq_h_row_rvx

COMMENT> 
IDENTITY> eq_h_test
EQ> eq_h_test = eq_h - (eq_h_nf + eq_h_f + eq_h_row)

COMMENT> 
IDENTITY> ins_h
EQ> ins_h= TSLAG(ins_h,1) + ins_h_tr + ins_h_rvx

COMMENT> 
IDENTITY> ins_h_tr
EQ> ins_h_tr = d8_h + ins_h_tr_excl_d8

COMMENT> 
BEHAVIORAL> d8_h
TSRANGE 2005 4 2020 1
EQ> TSDELTALOG(d8_h) = p_701*TSDELTALOG(TSLAG(d8_h,1)) + p_702*TSDELTALOG(w_h_r) + p_703*TSDELTALOG(TSLAG(old_age_ratio,1)) + p_704*LOG(TSLAG(d8_h,1)) + p_705*LOG(TSLAG(w_h_r,1)) + p_706*LOG(TSLAG(old_age_ratio,1)) + p_707*d_2014q3
COEFF> p_701 p_702 p_703 p_704 p_705 p_706 p_707
STORE> coe(7)

COMMENT> 
IDENTITY> lev_h
EQ> lev_h = l_h/yd_h

COMMENT> NET WEALTH AND FINANCIAL NET WEALTH

COMMENT> 
IDENTITY> fa_h
EQ> fa_h = iba_h + sec_h + eq_h + ins_h

COMMENT> 
IDENTITY> fnw_h
EQ> fnw_h = fa_h  + l_h

COMMENT> 
IDENTITY> fnw_hk
EQ> fnw_hk = fnw_h/pc

COMMENT> Capital Account

COMMENT>gross fixed capital formation (buildings and dwellings)
IDENTITY> i_bd_h
EQ> i_bd_h = i_bd_h_k * p_bd

COMMENT> gross fixed capital formation (equipment)
IDENTITY> i_equip_h
EQ> i_equip_h = i_equip_h_k * p_equip

COMMENT> Real net stock of buildings and dwellings
IDENTITY> bd_h_k
EQ> TSDELTA(bd_h_k) = i_bd_h_k - TSLAG(bd_h_k,1) * delta_bd_h

COMMENT>  Real net stock of equipment
IDENTITY> equip_h_k
EQ> TSDELTA(equip_h_k) = i_equip_h_k - TSLAG(equip_h_k,1) * delta_equip_h

COMMENT> Nominal net stock of buildings and dwellings
IDENTITY> bd_h
EQ> TSDELTA(bd_h) = i_bd_h + TSLAG(bd_h_k,1) * (1-delta_bd_h)*TSDELTA(p_bd) - TSLAG(bd_h_k,1) * delta_bd_h * p_bd

COMMENT> Nominal net stock of equipment
IDENTITY> equip_h
EQ> TSDELTA(equip_h) = i_equip_h + TSLAG(equip_h_k,1) * (1-delta_equip_h)*TSDELTA(p_equip) - TSLAG(equip_h_k,1) * delta_equip_h * p_equip

COMMENT> NON-FINANCIAL SECTOR

COMMENT> Real side

COMMENT> 
IDENTITY> y
EQ> y = pcon + g + i + x - m

COMMENT> real sales
IDENTITY> yk
EQ> yk = pconk + ik + gk + xk - mk

COMMENT> real sales
IDENTITY> yk_ds
EQ> yk_ds = yk - (alpha_90 + alpha_91*SD1 + alpha_92* SD2 + alpha_93*SD3)+ mean_yk

COMMENT> 
IDENTITY> sk
EQ> sk= pconk + ik + gk + xk

COMMENT> real government consumption
IDENTITY> gk
EQ> gk = g/pg

COMMENT> nominal export
IDENTITY> x
EQ> x = xk*px

COMMENT> 
IDENTITY> rer
EQ> rer = pc*xr/pf

COMMENT> 
BEHAVIORAL> xk_ds
TSRANGE 2006 3 2020 1
EQ> TSDELTALOG(xk_ds) =  p_652*TSDELTALOG(TSLAG(gdp_tp,4)) + p_653*TSDELTALOG(rer) + p_654*LOG(TSLAG(xk_ds,1)) + p_655*LOG(TSLAG(gdp_tp,1)) + p_656*LOG(TSLAG(rer,2)) + p_657*d_2008q2 + p_658*d_2018q1 + p_659*d_2019q3 + p_660* time
COEFF>  p_652 p_653 p_654 p_655 p_656 p_657 p_658 p_659 p_660
STORE> coe(8)

COMMENT> 
IDENTITY> xk
EQ> xk = xk_ds + alpha_80 + alpha_81*SD1 + alpha_82*SD2 + alpha_83*SD3 - mean_xk

COMMENT> nominal import
IDENTITY> m
EQ> m = mk*pm

COMMENT> 
BEHAVIORAL> mk_ds
TSRANGE 2006 2 2020 1
EQ> TSDELTALOG(mk_ds) = p_661 + p_662*TSDELTALOG(TSLAG(mk_ds,2)) + p_663*TSDELTALOG(TSLAG(rer,1)) + p_664*TSDELTALOG(TSLAG(rer,3)) + p_665*TSDELTALOG(yk_ds) + p_666*LOG(TSLAG(mk_ds,1)) + p_667*LOG(TSLAG(yk_ds,1)) + p_668*d_2009q1 + p_669*d_2009q4
COEFF> p_661 p_662 p_663 p_664 p_665 p_666 p_667 p_668 p_669
STORE> coe(9)

COMMENT> 
IDENTITY> mk
EQ> mk = mk_ds + alpha_70 + alpha_71*SD1 + alpha_72*SD2 + alpha_73*SD3 - mean_mk

COMMENT> GDP deflator
IDENTITY> py
EQ> py = y/yk

COMMENT> nominal sales
IDENTITY> s
EQ> s = pcon + g + i + x

COMMENT> wage bill paid by NFC
IDENTITY> w_nf
EQ> w_nf = w_h_r +  (w_row_r - w_row_p)

COMMENT> Gross operating surplus received by the NFC
IDENTITY> b2_nf_r
EQ> b2_nf_r = b2 - (b2_h_r + b2_f_r + b2_g_r)

COMMENT> 
IDENTITY> yf
EQ> yf = y-((p_tax+p_tax_row)-(p_sub+p_sub_row))

COMMENT> 
IDENTITY> b2
EQ> b2 = yf - w_nf

COMMENT> wage share
IDENTITY> ws
EQ> ws = w_nf/yf  * zz

COMMENT> profit share
IDENTITY> ps
EQ> ps = 1 - ws

COMMENT> Rate of capacity utilization
IDENTITY> u
EQ> u=yk/(bd_nfc_k+equip_nfc_k)

COMMENT> Profit rate
IDENTITY> prate
EQ> prate=s_nf/(TSLAG(bd_nfc,1)+TSLAG(equip_nfc,1))

COMMENT> Profit rate
IDENTITY> prate_ds
EQ> prate_ds = prate - (alpha_120 + alpha_121*SD1 + alpha_122*SD2 + alpha_123*SD3) + mean_prate

COMMENT> Profit rate
IDENTITY> ps_ds 
EQ> ps_ds  = ps - (alpha_130 + alpha_131*SD1 + alpha_132*SD2 + alpha_133*SD3) + mean_ps

COMMENT> 
IDENTITY> tobinq
EQ> tobinq = (eq_nf_l) / (equip_nfc + bd_nfc)

COMMENT> Er fixet
BEHAVIORAL> i_bd_nfc_k_ds
TSRANGE 2006 1 2020 1
EQ> LOG(i_bd_nfc_k_ds) = p_800 + 
p_801*TSDELTALOG(TSLAG(i_bd_nfc_k_ds,1)/TSLAG(bd_nfc_k,2)) + 
p_802*TSDELTALOG(ps_ds) + 
p_803*TSDELTALOG(u_ds) + 
p_804* LOG(TSLAG(i_bd_nfc_k_ds,1)/TSLAG(bd_nfc_k,2)) + 
p_805 * LOG(TSLAG(ps_ds,1)) + 
p_806 * LOG(TSLAG(u_ds,1)) + 
p_807 * TSDELTALOG(tobinq) +
p_808 * LOG(TSLAG(tobinq,1)) +
alpha_1*TSLAG(LOG(i_bd_nfc_k_ds),1) + 
alpha_2*TSLAG(LOG(bd_nfc_k),1) + 
alpha_3*TSLAG(-LOG(bd_nfc_k),2)
COEFF> p_800 p_801 p_802 p_803 p_804 p_805 p_806 p_807 p_808 alpha_1 alpha_2 alpha_3
RESTRICT> alpha_1 = 1
alpha_2 = 1
alpha_3 = 1
STORE> coe(10)

COMMENT> 
IDENTITY> i_bd_nfc_k
EQ> i_bd_nfc_k = i_bd_nfc_k_ds + (alpha_30 + alpha_31*SD1 + alpha_32*SD2 + alpha_33*SD3) - mean_i_bd_nfc_k

COMMENT> 
IDENTITY> i_equip_nfc_k
EQ> i_equip_nfc_k = i_equip_nfc_k_ds + (alpha_40 + alpha_41*SD1 + alpha_42*SD2 + alpha_43*SD3) - mean_i_equip_nfc_k

COMMENT> 
IDENTITY> ik
EQ> ik =  i_equip_h_k + i_equip_nfc_k  +  i_equip_g_k + i_equip_fc_k +  i_bd_h_k + i_bd_nfc_k  +  i_bd_g_k  + i_bd_fc_k + i_adj_h_k + i_adj_nfc_k + i_adj_fc_k + i_adj_g_k

COMMENT> 
IDENTITY> i_h
EQ> i_h = i_equip_h_k*p_equip +  i_bd_h_k*p_bd  + i_adj_h_k*p_equip

COMMENT> 
IDENTITY> i_nf
EQ> i_nf = i_equip_nfc_k*p_equip +  i_bd_nfc_k*p_bd  + i_adj_nfc_k*p_equip

COMMENT> 
IDENTITY> i_f
EQ> i_f = i_equip_fc_k*p_equip +  i_bd_fc_k*p_bd  + i_adj_fc_k*p_equip

COMMENT> 
IDENTITY> i_g
EQ> i_g = i_equip_g_k*p_equip +  i_bd_g_k*p_bd  + i_adj_g_k*p_equip

COMMENT> 
IDENTITY> i
EQ> i = i_nf + i_f + i_g + i_h

COMMENT> 
BEHAVIORAL> i_equip_nfc_k_ds
TSRANGE 2006 1 2020 1
EQ> LOG(i_equip_nfc_k_ds) = p_807 + p_808 * TSDELTALOG(TSLAG(i_equip_nfc_k_ds,1)/TSLAG(equip_nfc_k,2)) + p_809 * TSDELTALOG(ps_ds) + p_810 * TSDELTALOG(u_ds) + p_811 *  LOG(TSLAG(i_equip_nfc_k_ds,1)/TSLAG(equip_nfc_k,2)) + p_812 * LOG(TSLAG(ps_ds,1)) + p_813 * LOG(TSLAG(u_ds,1)) + p_814 * dummy_10 + p_815 * dummy_11 + p_890*TSDELTALOG(tobinq) + p_891*LOG(TSLAG(tobinq,1)) + alpha_1*TSLAG(LOG(i_equip_nfc_k_ds),1) + alpha_2*TSLAG(LOG(equip_nfc_k),1) + alpha_3*TSLAG(-LOG(equip_nfc_k),2)
COEFF> p_807 p_808 p_809 p_810 p_811 p_812 p_813 p_814 p_815 p_890 p_891 alpha_1 alpha_2 alpha_3
RESTRICT> alpha_1 = 1
alpha_2 = 1
alpha_3 = 1
STORE> coe(11)

COMMENT> 
IDENTITY> int_nf
EQ> int_nf = (TSLAG(idep,1)*TSLAG(iba_nf,1) + TSLAG(ibd,1)*TSLAG(sec_nf,1) + TSLAG(iloan,1)*TSLAG(l_nf,1))

COMMENT> 
IDENTITY> div_nf
EQ> div_nf =d_20124*(((TSLAG(eq_nf_a_nf,1) + TSLAG(eq_nf_a_f,1) - TSLAG(eq_nf_l_nf,1) + TSLAG(eq_h_nf,1)+ TSLAG(eq_row_a_nf,1) + TSLAG(eq_f_a_nf,1) + TSLAG(eq_g_nf,1)))* TSLAG(divd,1) + TSLAG(eq_nf_a_row,1)*TSLAG(diva,1)) + (1-d_20124)*(TSLAG(neq_nf,1)*TSLAG(divd,1)) + div_nf_adj

COMMENT> 
IDENTITY> insu_nf
EQ> insu_nf =  TSLAG(ins_nf,1)*TSLAG(insu,1)

COMMENT> 
IDENTITY> npropinc_nf
EQ> npropinc_nf = int_nf + int_nf_adj + div_nf + insu_nf + ins_nf_adj + (res_r_nf ) - (res_p_nf)

COMMENT> gross income NFC Ændret yil den anden i modellen 
IDENTITY> yh_nf
EQ> yh_nf = y + b2_nf_r - b2 - (p_tax + p_tax_row) + (p_sub+p_sub_row) - w_nf + (propinc_r_nf - propinc_p_nf)

COMMENT> savings NFC TAX_NF is negative
IDENTITY> s_nf
EQ> s_nf = yh_nf + tax_nf + oth_nf

COMMENT> 
IDENTITY> tax_nf
EQ> tax_nf = -tax_rate_nf*y

COMMENT> Sectoral balance NFC
IDENTITY> nl_nf
EQ> nl_nf = s_nf - i_nf - np_nf + ctr_nf

COMMENT> Financial side

COMMENT> Change in stocks

COMMENT> iba
IDENTITY> iba_nf
EQ> iba_nf = TSLAG(iba_nf,1) + iba_nf_tr + iba_nf_rvx

COMMENT> 
IDENTITY> iba_nf_tr
EQ> iba_nf_tr = nl_nf + nl_nf_adj - neq_nf_tr - l_nf_tr - sec_nf_tr - ins_nf_tr

COMMENT> SEC . For the time being, we take sec_nf_tr as exogenous.
IDENTITY> sec_nf
EQ> sec_nf = TSLAG(sec_nf,1) + sec_nf_tr + sec_nf_rvx

COMMENT>  Loans
IDENTITY> l_nf
EQ> l_nf = TSLAG(l_nf,1) + l_nf_tr + l_nf_rvx

COMMENT> CHANGE in STOCKS OF EQUITIES:

COMMENT> Equities held as assets by NFC
IDENTITY> eq_nf_a
EQ> eq_nf_a = (1-d_20124)*(TSLAG(eq_nf_a,1) + eq_nf_a_tr + eq_nf_a_rvx)  + d_20124*eq_nf_ax

COMMENT> 
IDENTITY> eq_nf_a_nf
EQ> eq_nf_a_nf = d_20124 * (TSLAG(eq_nf_a_nf,1) + eq_nf_a_nf_tr + eq_nf_a_nf_rv)

COMMENT> equiteies NFC-NFC asset is obviously equal to equities NFC-NFC liability
IDENTITY> eq_nf_l_nf
EQ> eq_nf_l_nf = eq_nf_a_nf

COMMENT> this is the data for quities as assets for NFC viz-a-viz counter parties, e.g., NFC purchase equities from NFC, FC, and RoW
IDENTITY> eq_nf_ax
EQ> eq_nf_ax =  eq_nf_a_nf + eq_nf_a_f + eq_nf_a_row

COMMENT> 
IDENTITY> eq_nf_a_f
EQ> eq_nf_a_f = d_20124 * (TSLAG(eq_nf_a_f,1) + eq_nf_a_f_tr + eq_nf_a_f_rv)

COMMENT> 
IDENTITY> eq_nf_a_row
EQ> eq_nf_a_row = d_20124 * (TSLAG(eq_nf_a_row,1) + eq_nf_a_row_tr + eq_nf_a_row_rv)

COMMENT> total gross liabilities of equities on firms balance sheet is defined as the sum of NFC-NFC asset + the sum of NFC equities held by other sectors
IDENTITY> eq_nf_l
EQ> eq_nf_l = (1- d_20124)*(TSLAG(eq_nf_l,1) + eq_nf_l_tr + eq_nf_l_rvx)   +  d_20124*eq_nf_lx

COMMENT> total gross liabilities of equities on firms balance sheet is defined as the sum of NFC-NFC asset + the sum of NFC equities held by other sectors
IDENTITY> eq_nf_lx
EQ> eq_nf_lx =   eq_nf_a_nf + eq_h_nf + eq_row_a_nf + eq_f_a_nf + eq_g_nf

COMMENT> 
IDENTITY> neq_nf
EQ> neq_nf =d_20123*(-alpha_neq_nf * i_nf) + d_20124*( (eq_nf_a_f+ eq_nf_a_row)-(eq_f_a_nf + eq_g_nf + eq_h_nf + eq_row_a_nf))

COMMENT> 
IDENTITY> neq_nf_tr
EQ> neq_nf_tr = TSDELTA(neq_nf)-(neq_nf_rvx)

COMMENT>Some consistency check eqs:

COMMENT> this variable should be zero after 2012 and onwards as it captures the difference between equities as assets for nfc (aggregate) data minus equities as assets for NFC held at difference sectors after 2012. In short, after 2012, we can identify the counter parties where NFC are holding stocks.
IDENTITY> eq_nf_a_test
EQ> eq_nf_a_test = eq_nf_a - eq_nf_ax

COMMENT> this variable should be zero after 2012 and onwards as it captures the difference between equities as liabilities for nfc (aggregate) data minus equities as liabilities for NFC held at difference sectors after 2012. 
IDENTITY> eq_nf_l_test
EQ> eq_nf_l_test = eq_nf_l - eq_nf_lx

COMMENT> insurance NFC
IDENTITY> ins_nf
EQ> ins_nf = TSLAG(ins_nf,1) + ins_nf_tr + ins_nf_rvx

COMMENT> FINANCIAL NET WEALTH

COMMENT> net financial wealth
IDENTITY> fnw_nf
EQ> fnw_nf = iba_nf + eq_nf_a - eq_nf_l + sec_nf +l_nf + ins_nf

COMMENT> 
IDENTITY> fnw_nfk
EQ> fnw_nfk = fnw_nf/pc

COMMENT> Capital Account

COMMENT> gross fixed capital formation (buildings and dwellings)
IDENTITY> i_bd_nfc
EQ> i_bd_nfc = i_bd_nfc_k * p_bd

COMMENT> 
IDENTITY> i_equip_nfc
EQ> i_equip_nfc = i_equip_nfc_k * p_equip

COMMENT>  Real net stock of buildings and dwellings
IDENTITY> bd_nfc_k
EQ> TSDELTA(bd_nfc_k) = i_bd_nfc_k - TSLAG(bd_nfc_k,1) * delta_bd_nfc

COMMENT> Real net stock of equipment
IDENTITY> equip_nfc_k
EQ> TSDELTA(equip_nfc_k) = i_equip_nfc_k - TSLAG(equip_nfc_k,1) * delta_equip_nfc

COMMENT> Nominal net stock of buildings and dwellings
IDENTITY> bd_nfc
EQ> TSDELTA(bd_nfc) = i_bd_nfc + TSLAG(bd_nfc_k,1) * (1-delta_bd_nfc)*TSDELTA(p_bd) - TSLAG(bd_nfc_k,1) * delta_bd_nfc * p_bd

COMMENT> Nominal net stock of equipment
IDENTITY> equip_nfc
EQ> TSDELTA(equip_nfc) = i_equip_nfc + TSLAG(equip_nfc_k,1) * (1-delta_equip_nfc)*TSDELTA(p_equip) - TSLAG(equip_nfc_k,1) * delta_equip_nfc * p_equip

COMMENT> FINANCIAL CORP

COMMENT> Real side

COMMENT> 
IDENTITY> int_f
EQ> int_f = (TSLAG(idep,1)*TSLAG(iba_f,1) + TSLAG(iloan,1)*TSLAG(l_f,1) + TSLAG(ibd,1)*TSLAG(sec_f_d,1) + TSLAG(iboa,1)*TSLAG(sec_f_a,1))

COMMENT> 
IDENTITY> div_f
EQ> div_f = d_20124 *(((TSLAG(eq_f_a_f,1) + TSLAG(eq_nf_a_nf,1)) - (TSLAG(eq_f_l_f,1) + TSLAG(eq_h_f ,1)+ TSLAG(eq_row_a_f,1) + TSLAG(eq_nf_a_f,1) + TSLAG(eq_g_f,1)))* TSLAG(divd,1) + TSLAG(eq_f_a_row,1)*TSLAG(diva,1)) + (1-d_20124)*((TSLAG(neq_f,1)*TSLAG(divd,1))) +  div_f_adj

COMMENT> 
IDENTITY> insu_f
EQ> insu_f = TSLAG(ins_f,1)*TSLAG(insu,1)

COMMENT> 
IDENTITY> propinc_r_f
EQ> propinc_r_f = int_r_f + div_r_f + ins_r_f + res_r_f

COMMENT> 
IDENTITY> propinc_p_f
EQ> propinc_p_f = int_p_f + div_p_f + ins_p_f + res_p_f

COMMENT> 
IDENTITY> npropinc_f
EQ> npropinc_f = int_f + int_f_adj +  div_f + insu_f + ins_f_adj + (res_r_f) - (res_p_f)

COMMENT> 
IDENTITY> yh_f
EQ> yh_f = b2_f_r + (npropinc_f)

COMMENT> 
IDENTITY> yd_f
EQ> yd_f = yh_f + tax_f + nben_f + oth_f

COMMENT> This is based on the fact that d8_h = d8_f in the data.
IDENTITY> d8_f
EQ> d8_f = d8_h

COMMENT> 
IDENTITY> nben_f
EQ> nben_f = d8_f

COMMENT> Savings F
IDENTITY> s_f
EQ> s_f = yd_f - d8_f

COMMENT> Sectoral balance F
IDENTITY> nl_f
EQ> nl_f = s_f - i_f - np_f + ctr_f

COMMENT> FInancial side

COMMENT> Change in stocks of ibas

COMMENT> 
IDENTITY> iba_f
EQ>  iba_f = TSLAG(iba_f,1) + iba_f_tr + iba_f_rvx

COMMENT> 
IDENTITY> iba_f_tr
EQ> iba_f_tr = -(iba_nf_tr + iba_h_tr + iba_g_tr + iba_row_tr)

COMMENT> Change in stocks of equities:

COMMENT> 
IDENTITY> eq_f_a
EQ> eq_f_a = (1-d_20124)*(TSLAG(eq_f_a,1) + eq_f_a_tr + eq_f_a_rvx)  + d_20124*eq_f_ax

COMMENT> 
IDENTITY> eq_f_a_nf
EQ> eq_f_a_nf = d_20124 * (TSLAG(eq_f_a_nf,1) + eq_f_a_nf_tr + eq_f_a_nf_rv)

COMMENT> 
IDENTITY> eq_f_a_f
EQ> eq_f_a_f = d_20124 * (TSLAG(eq_f_a_f,1) + eq_f_a_f_tr + eq_f_a_f_rv)

COMMENT> 
IDENTITY> eq_f_l_f
EQ> eq_f_l_f = eq_f_a_f

COMMENT> 
IDENTITY> eq_f_a_row
EQ> eq_f_a_row = d_20124 * (TSLAG(eq_f_a_row,1) + eq_f_a_row_tr + eq_f_a_row_rv)

COMMENT> this is the data for quities as assets for NFC viz-a-viz counter parties, e.g., NFC purchase equities from NFC, FC, and RoW
IDENTITY> eq_f_ax
EQ> eq_f_ax =  eq_f_a_nf + eq_f_a_f + eq_f_a_row

COMMENT> 
IDENTITY> eq_f_l
EQ> eq_f_l = (1- d_20124)*(TSLAG(eq_f_l,1) + eq_f_l_tr + eq_f_l_rvx)   +  d_20124*eq_f_lx

COMMENT> 
IDENTITY> eq_f_lx
EQ> eq_f_lx =   eq_nf_a_f + eq_h_f + eq_row_a_f + eq_f_a_f + eq_g_f

COMMENT> 
IDENTITY> neq_f
EQ> neq_f = eq_f_a - eq_f_l

COMMENT> 
IDENTITY> neq_f_tr
EQ> neq_f_tr = TSDELTA(neq_f) - neq_f_rvx

COMMENT> 
IDENTITY> eq_f_l_test
EQ> eq_f_l_test = eq_f_l - (eq_f_l_f + eq_h_f + eq_row_a_f + eq_nf_a_f + eq_g_f)

COMMENT> Change in stocks of securities

COMMENT> 
IDENTITY> sec_f_d
EQ> sec_f_d = TSLAG(sec_f_d,1) + sec_f_d_tr + sec_f_d_rvx

COMMENT> 
IDENTITY> sec_f_d_tr
EQ> sec_f_d_tr =  nl_f + nl_f_adj - iba_f_tr - ins_f_tr - l_f_tr - neq_f_tr - sec_f_a_tr

COMMENT> 
IDENTITY> sec_f_d_tr_nf
EQ> sec_f_d_tr_nf=-sec_nf_tr

COMMENT> 
IDENTITY> sec_f_d_tr_g
EQ> sec_f_d_tr_g=sec_f_d_tr-sec_f_d_tr_nf

COMMENT> 
IDENTITY> sec_f_a
EQ> sec_f_a = TSLAG(sec_f_a,1) + sec_f_a_tr + sec_f_a_rvx

COMMENT> 
IDENTITY> sec_f_a_tr
EQ> sec_f_a_tr=-sec_row_tr

COMMENT> 
IDENTITY> l_f
EQ> l_f = -(l_h + l_nf + l_g + l_row)

COMMENT> 
IDENTITY> l_f_tr
EQ> l_f_tr = TSDELTA(l_f) -l_f_rvx

COMMENT> 
IDENTITY> ins_f
EQ> ins_f = -(ins_h + ins_row + ins_nf + ins_g)

COMMENT> 
IDENTITY> ins_f_tr
EQ> ins_f_tr = TSDELTA(ins_f) - ins_f_rvx

COMMENT> 'WEALTH

COMMENT> 
IDENTITY> fnw_f
EQ> fnw_f = iba_f + sec_f_a + sec_f_d +l_f + eq_f_a - eq_f_l + ins_f

COMMENT> 
IDENTITY> fnw_fk
EQ> fnw_fk = fnw_f/pc

COMMENT> Capital Account

COMMENT> gross fixed capital formation (buildings and dwellings)
IDENTITY> i_bd_fc
EQ> i_bd_fc = i_bd_fc_k * p_bd

COMMENT> gross fixed capital formation (equipment)
IDENTITY> i_equip_fc
EQ> i_equip_fc = i_equip_fc_k * p_equip

COMMENT> 
IDENTITY> bd_fc_k
EQ> TSDELTA(bd_fc_k) = i_bd_fc_k - TSLAG(bd_fc_k,1) * delta_bd_fc

COMMENT> 
IDENTITY> equip_fc_k
EQ> TSDELTA(equip_fc_k) = i_equip_fc_k - TSLAG(equip_fc_k,1) * delta_equip_fc

COMMENT> 
IDENTITY> bd_fc
EQ> TSDELTA(bd_fc) = i_bd_fc + TSLAG(bd_fc_k,1) * (1-delta_bd_fc)*TSDELTA(p_bd) - TSLAG(bd_fc_k,1) * delta_bd_fc * p_bd

COMMENT> 
IDENTITY> equip_fc
EQ> TSDELTA(equip_fc) = i_equip_fc + TSLAG(equip_fc_k,1) * (1-delta_equip_fc)*TSDELTA(p_equip) - TSLAG(equip_fc_k,1) * delta_equip_fc * p_equip

COMMENT> GOVT

COMMENT> Real side

COMMENT> 
IDENTITY> int_g
EQ> int_g = (TSLAG(idep,1)*TSLAG(iba_g,1) + TSLAG(iloan,1)*TSLAG(l_g,1) + TSLAG(ibd,1)*TSLAG(sec_g,1))

COMMENT> 
IDENTITY> div_g
EQ> div_g = d_20124 * ( (TSLAG(eq_g_nf,1) + TSLAG(eq_g_f,1))*TSLAG(divd,1) + TSLAG(eq_g_row,1)*TSLAG(diva,1) ) + (1-d_20124) * ((TSLAG(eq_g,1)*TSLAG(divd,1))) + div_g_adj

COMMENT> 
IDENTITY> insu_g
EQ> insu_g = TSLAG(ins_g,1)*TSLAG(insu,1)

COMMENT> 
IDENTITY> propinc_r_g
EQ> propinc_r_g = int_r_g + div_r_g + res_r_g

COMMENT> 
IDENTITY> propinc_p_g
EQ> propinc_p_g = int_p_g + res_p_g

COMMENT> 
IDENTITY> npropinc_g
EQ> npropinc_g = int_g + int_g_adj + div_g + insu_g + ins_g_adj + (res_r_g) - ( res_p_g)

COMMENT> 
IDENTITY> yh_g
EQ> yh_g = b2_g_r + p_tax - p_sub + (npropinc_g)

COMMENT> 
IDENTITY> yd_g
EQ> yd_g = yh_g + tax_g + nben_g + oth_g

COMMENT> 
IDENTITY> nben_g
EQ> nben_g = - (nben_row + nsben_h + nben_f)

COMMENT> 
IDENTITY> tax_g
EQ> tax_g = -(tax_h + tax_nf + tax_f + tax_row)

COMMENT> 
IDENTITY> s_g
EQ> s_g = yd_g - g

COMMENT> 
IDENTITY> nl_g
EQ> nl_g = s_g - i_g - np_g + ctr_g

COMMENT> Financial side

COMMENT> Change in stocks ' we need to have an accomodating asset

COMMENT> 
IDENTITY> iba_g
EQ> iba_g= TSLAG(iba_g,1) + iba_g_tr + iba_g_rvx

COMMENT> 
IDENTITY> eq_g
EQ> eq_g = (1-d_20124)*(TSLAG(eq_g,1) + eq_g_tr + eq_g_rvx) + d_20124*eq_g_ax

COMMENT> 
IDENTITY> eq_g_nf
EQ> eq_g_nf = d_20124 * (TSLAG(eq_g_nf,1) + eq_g_nf_tr + eq_g_nf_rv)

COMMENT> 
IDENTITY> eq_g_f
EQ> eq_g_f = d_20124 * (TSLAG(eq_g_f,1) + eq_g_f_tr + eq_g_f_rv)

COMMENT> 
IDENTITY> eq_g_row
EQ> eq_g_row = d_20124 * (TSLAG(eq_g_row,1) + eq_g_row_tr + eq_g_row_rv)

COMMENT> 
IDENTITY> eq_g_ax
EQ> eq_g_ax =  eq_g_row + eq_g_f + eq_g_nf

COMMENT> 
IDENTITY> eq_g_test
EQ> eq_g_test = eq_g - (eq_g_nf + eq_g_f + eq_g_row)

COMMENT> 
IDENTITY> sec_g
EQ> sec_g = TSLAG(sec_g,1) + sec_g_tr + sec_g_rvx

COMMENT> 
IDENTITY> sec_g_tr
EQ> sec_g_tr = nl_g + nl_g_adj - iba_g_tr - l_g_tr - eq_g_tr - ins_g_tr

COMMENT> SKAL REDDUNDANT EQ MED HER???? redundant = (sec_g_tr_0+(sec_h_tr_g_0+sec_f_d_tr_g_0))  'Redundant equation

COMMENT> 
IDENTITY> l_g
EQ> l_g = TSLAG(l_g,1) + l_g_tr + l_g_rvx

COMMENT> 
IDENTITY> ins_g
EQ> ins_g = TSLAG(ins_g,1) + ins_g_tr + ins_g_rvx

COMMENT> WEALTH

COMMENT> 
IDENTITY> fnw_g
EQ> fnw_g = iba_g + eq_g + sec_g +l_g + ins_g

COMMENT> 
IDENTITY> fnw_gk
EQ> fnw_gk = fnw_g/pc

COMMENT> Capital Account

COMMENT> gross fixed capital formation (buildings and dwellings)
IDENTITY> i_bd_g
EQ> i_bd_g = i_bd_g_k * p_bd

COMMENT> gross fixed capital formation (equipment)
IDENTITY> i_equip_g
EQ> i_equip_g = i_equip_g_k * p_equip

COMMENT> Real net stock of buildings and dwellings
IDENTITY> bd_g_k
EQ> TSDELTA(bd_g_k) = i_bd_g_k - TSLAG(bd_g_k,1) * delta_bd_g

COMMENT> 
IDENTITY> equip_g_k
EQ> TSDELTA(equip_g_k) = i_equip_g_k - TSLAG(equip_g_k,1) * delta_equip_g

COMMENT> 
IDENTITY> bd_g
EQ> TSDELTA(bd_g) = i_bd_g + TSLAG(bd_g_k,1) * (1-delta_bd_g)*TSDELTA(p_bd) - TSLAG(bd_g_k,1) * delta_bd_g * p_bd

COMMENT> 
IDENTITY> equip_g
EQ> TSDELTA(equip_g) = i_equip_g + TSLAG(equip_g_k,1) * (1-delta_equip_g)*TSDELTA(p_equip) - TSLAG(equip_g_k,1) * delta_equip_g * p_equip

COMMENT> REST OF THE WORLD

COMMENT> Real side

COMMENT> 
IDENTITY> int_row
EQ> int_row = (TSLAG(idep,1)*TSLAG(iba_row,1) + TSLAG(iboa,1)*TSLAG(sec_row,1) + TSLAG(iloan,1)*TSLAG(l_row,1))

COMMENT> 
IDENTITY> div_row
EQ> div_row = d_20124*(  (TSLAG(eq_row_a_nf,1) + TSLAG(eq_row_a_f,1))*TSLAG(divd,1) - ((TSLAG(eq_nf_a_row,1)+TSLAG(eq_f_a_row,1)+TSLAG(eq_g_row,1)+TSLAG(eq_h_row,1))*TSLAG(diva,1))) + (1-d_20124) * ((TSLAG(neq_row,1)*TSLAG(divd,1))) + div_row_adj

COMMENT> 
IDENTITY> insu_row
EQ> insu_row = TSLAG(ins_row,1)*TSLAG(insu,1)

COMMENT> 
IDENTITY> propinc_r_row
EQ> propinc_r_row = int_r_row + div_r_row + ins_r_row + res_r_row

COMMENT> 
IDENTITY> propinc_p_row
EQ> propinc_p_row = int_p_row + div_p_row + ins_p_row + res_p_row

COMMENT> 
IDENTITY> npropinc_row
EQ> npropinc_row = int_row + int_row_adj + div_row + insu_row + ins_row_adj + (res_r_row) - (res_p_row)

COMMENT> 
IDENTITY> s_row
EQ> s_row = m - x + p_tax_row - p_sub_row + (w_row_r - w_row_p) +  (npropinc_row) + (tax_row) + nben_row + oth_row

COMMENT> 
IDENTITY> nl_row
EQ> nl_row = s_row + ctr_row - np_row

COMMENT> 
IDENTITY> nben_row
EQ> nben_row = scon_row_r - scon_row_p + sben_row_r - sben_row_p

COMMENT> 
IDENTITY> nx
EQ> nx = x - m

COMMENT> 
IDENTITY> cab
EQ> cab =-nl_row

COMMENT> 
IDENTITY> bop
EQ> bop = cab + fab

COMMENT> 
IDENTITY> fab
EQ> fab = (fnl_row)

COMMENT> Financial side

COMMENT> 
IDENTITY> iba_row
EQ> iba_row= TSLAG(iba_row,1) + iba_row_tr + iba_row_rvx

COMMENT> 
IDENTITY> iba_row_tr
EQ> iba_row_tr = nl_row + nl_row_adj - neq_row_tr - sec_row_tr - l_row_tr - ins_row_tr

COMMENT> 
IDENTITY> sec_row
EQ> sec_row = TSLAG(sec_row,1) + sec_row_tr + sec_row_rvx

COMMENT> 
IDENTITY> l_row
EQ> l_row = TSLAG(l_row,1) +l_row_tr +l_row_rvx

COMMENT> CHANGE in equities

COMMENT> 
IDENTITY> eq_row_a
EQ> eq_row_a = (1-d_20124)*(TSLAG(eq_row_a,1) + eq_row_a_tr + eq_row_a_rvx)  + d_20124*eq_row_ax

COMMENT> 
IDENTITY> eq_row_a_nf
EQ> eq_row_a_nf = d_20124 * (TSLAG(eq_row_a_nf,1) + eq_row_a_nf_tr + eq_row_a_nf_rv)

COMMENT> 
IDENTITY> eq_row_a_f
EQ> eq_row_a_f = d_20124 * (TSLAG(eq_row_a_f,1) + eq_row_a_f_tr + eq_row_a_f_rv)

COMMENT> 
IDENTITY> eq_row_ax
EQ> eq_row_ax =  eq_row_a_nf + eq_row_a_f

COMMENT> 
IDENTITY> eq_row_a_test
EQ> eq_row_a_test = eq_row_a - (eq_row_a_nf + eq_row_a_f)

COMMENT> 
IDENTITY> eq_row_l
EQ> eq_row_l = (1-d_20124)*(TSLAG(eq_row_l,1) + eq_row_l_tr + eq_row_l_rvx) + d_20124*(eq_row_lx)

COMMENT> 
IDENTITY> eq_row_lx
EQ> eq_row_lx = eq_nf_a_row + eq_h_row + eq_f_a_row + eq_g_row

COMMENT> 
IDENTITY> eq_row_l_test
EQ> eq_row_l_test = eq_row_l - (eq_nf_a_row + eq_f_a_row + eq_g_row + eq_h_row)

COMMENT> 
IDENTITY> neq_row
EQ> neq_row =-d_20123*(neq_nf + eq_h + neq_f + eq_g) + d_20124*( eq_row_a_nf + eq_row_a_f - (eq_nf_a_row + eq_f_a_row + eq_g_row + eq_h_row))

COMMENT> 
IDENTITY> neq_row_tr
EQ> neq_row_tr = TSDELTA(neq_row) - neq_row_rvx

COMMENT> 
IDENTITY> ins_row
EQ> ins_row = TSLAG(ins_row,1) + ins_row_tr + ins_row_rvx

COMMENT> WEALTH

COMMENT> 
IDENTITY> fnw_row
EQ> fnw_row = iba_row + eq_row_a - eq_row_l + sec_row +l_row + ins_row

COMMENT> 
IDENTITY> fnw_rowk
EQ> fnw_rowk = fnw_row/pc

COMMENT> Labour market

COMMENT> 
IDENTITY> unadj
EQ> unadj = lf - empadj

COMMENT> 
IDENTITY> unemp
EQ> unemp = lf - emp

COMMENT> 
IDENTITY> ur
EQ> ur = unemp/lf

COMMENT> 
IDENTITY> ur_ds
EQ> ur_ds = ur - (alpha_140 + alpha_141*SD1 + alpha_142*SD2 + alpha_143*SD3) + mean_ur

COMMENT> 
IDENTITY> uradj
EQ> uradj = unadj/lf

COMMENT> 
IDENTITY> lf
EQ> lf = (part*pop)/1000
IF> shock_part_model <=0
IDENTITY> lf
EQ> lf = (part1*pop*(1-old_age_ratio))/1000
IF> shock_part_model > 0

COMMENT> The pricing equation implies a mark-up of (1 + 0.38) in the long-run, i.e., the long-run (co-integrating vector) impact of the indepenent variable on pc_ds = 0.029/0.021

COMMENT> 
BEHAVIORAL> pc_ds
TSRANGE 2006 3 2020 1
EQ> TSDELTALOG(pc_ds) = C_21*TSDELTALOG(TSLAG(pc_ds,1)) + C_22*TSDELTALOG(TSLAG(pc_ds,2))+ C_23*TSDELTALOG(TSLAG(pc_ds,3)) + C_24*TSDELTALOG(TSLAG(pc_ds,4)) + C_25*TSDELTALOG(wage_ds)+ C_26*TSDELTALOG(pm_ds) + C_27*TSDELTALOG(TSLAG(pm_ds,2)) + C_28*LOG(TSLAG(pc_ds,1)) + C_29*LOG(TSLAG(wage_ds,1)) + C_30*LOG(TSLAG(prod_ds,1)) + C_31*LOG(TSLAG(pm_ds,1)) + C_32* d_2007q3 + C_33*d_2017q23
COEFF> C_21 C_22 C_23 C_24 C_25 C_26 C_27 C_28 C_29 C_30 C_31 C_32 C_33 
STORE> coe(12)

COMMENT> 
IDENTITY> pc
EQ> pc = pc_ds + alpha_60 + alpha_61*SD1 + alpha_62*SD2 + alpha_63*SD3 - mean_pc

COMMENT> 
IDENTITY> mkp
EQ> mkp = pc_ds / (wage_ds/prodk_ds + pm_ds)

COMMENT> 
BEHAVIORAL> emp
EQ> emp = p_250 * (yk/prodk) 
COEFF> p_250 
STORE> coe(14)

COMMENT> 
IDENTITY> urterm
EQ> urterm = ur - urs

COMMENT> 
IDENTITY> rw_ds
EQ> rw_ds= wage_ds/pc_ds

COMMENT> 
IDENTITY> wage
EQ> wage = wage_ds  + (alpha_50 + alpha_51*SD1 + alpha_52*SD2 + alpha_53*SD3) - mean_wage

COMMENT> 
IDENTITY> wage_ds_t
EQ> wage_ds_t= SD1* wage_ds*(1+inflation_t) + SD2* TSLAG(wage_ds,1)*(1+TSLAG(inflation_t,1)) + SD3* TSLAG(wage_ds,2)*(1+TSLAG(inflation_t,2)) + SD4* TSLAG(wage_ds,3)*(1+TSLAG(inflation_t,3))

COMMENT> 
IDENTITY> inflation
EQ> inflation = (pc/TSLAG(pc,4)) - 1

COMMENT> 
IDENTITY> inflation_t
EQ> inflation_t = inflation_tt
IF> inflation_tt <= 0
IDENTITY> inflation_t
EQ> inflation_t = inflation
IF> inflation_tt > 0

COMMENT>  
IDENTITY> wageindex
EQ> wageindex = wage/wage_2010q3

COMMENT> 
IDENTITY> mkp
EQ> mkp = pc_ds / (wage_ds/prodk_ds + pm_ds)

COMMENT> 
BEHAVIORAL> wage_ds
TSRANGE 2006 3 2020 1
EQ> TSDELTALOG(wage_ds) = p_522* TSDELTA(TSLAG(ur,4)) + p_523* TSDELTALOG(prod_ds) + p_max* TSDELTALOG(TSLAG(max_dp,1))  + p_524* LOG(TSLAG(wage_ds,1)) + p_525* LOG(TSLAG(wage_ds_t,1)) + p_526* LOG(TSLAG(prod_ds,1))  + p_527 * (d_2009q1 + d_2009q2)
COEFF> p_522 p_523 p_max p_524 p_525 p_526 p_527
STORE> coe(20)

COMMENT> 
IDENTITY> prod
EQ> prod=prodk * (y/yk)

COMMENT> 
IDENTITY> prod_ds
EQ> prod_ds=prodk_ds * (y_ds/yk_ds)

COMMENT> 
IDENTITY> acc_rate
EQ> acc_rate = ik / (bd_h_k + equip_h_k + bd_nfc_k + equip_nfc_k + bd_fc_k + equip_fc_k + bd_g_k + equip_g_k)

COMMENT> CONSISTENCY CHEKCS

COMMENT> 
IDENTITY> nl_check
EQ> nl_check= (nl_h + nl_f + nl_nf + nl_g + nl_row)

COMMENT> 
IDENTITY> fnl_check
EQ> fnl_check= (fnl_h + fnl_f + fnl_nf + fnl_g + fnl_row)

COMMENT> 
IDENTITY> check_np
EQ> check_np = np_row + np_h + np_f + np_nf + np_g

COMMENT> 
IDENTITY> check_ctr
EQ> check_ctr = ctr_row + ctr_h + ctr_f + ctr_nf + ctr_g

COMMENT> 
IDENTITY> check_invest
EQ> check_invest = i - i_f - i_nf - i_g - i_h

COMMENT> 
IDENTITY> check_iba
EQ> check_iba = iba_h + iba_f + iba_nf + iba_g + iba_row

COMMENT> 
IDENTITY> check_tax
EQ> check_tax = tax_g + tax_nf + tax_f + tax_h + tax_row

COMMENT> 
IDENTITY> check_iba_rv
EQ> check_iba_rv = iba_h_rv + iba_f_rv + iba_nf_rv + iba_g_rv + iba_row_rv

COMMENT> 
IDENTITY> check_iba_tr
EQ> check_iba_tr = iba_h_tr + iba_f_tr + iba_nf_tr + iba_g_tr + iba_row_tr

COMMENT> 
IDENTITY> check_eq
EQ> check_eq = eq_h + (eq_f_a - eq_f_l) + (eq_nf_a - eq_nf_l) + eq_g + (eq_row_a - eq_row_l)

COMMENT> 
IDENTITY> check_eq_tr
EQ> check_eq_tr = eq_h_tr + neq_nf_tr  + neq_f_tr + eq_g_tr + neq_row_tr + eq_g_tr

COMMENT> 
IDENTITY> check_div
EQ> check_div = div_h + div_f + div_g + div_nf + div_row

COMMENT> 
IDENTITY> check_l
EQ> check_l = l_h + l_f + l_nf + l_g + l_row

COMMENT> 
IDENTITY> check_l_tr
EQ> check_l_tr = l_f_tr + l_h_tr + l_nf_tr + l_g_tr + l_row_tr

COMMENT> 
IDENTITY> check_ins
EQ> check_ins = ins_h + ins_f + ins_nf + ins_g + ins_row

COMMENT> 
IDENTITY> check_ins_tr
EQ> check_ins_tr =  ins_f_tr  + ins_h_tr + ins_nf_tr + ins_g_tr + ins_row_tr

COMMENT> 
IDENTITY> check_sec
EQ> check_sec = sec_h + sec_f_a + sec_f_d + sec_nf + sec_g + sec_row

COMMENT> 
IDENTITY> check_sec
EQ> check_sec = sec_h + sec_f_a + sec_f_d + sec_nf + sec_g + sec_row

IDENTITY> part
EQ> part = part*1
IF> shock_part_model <= 0
IDENTITY> part
EQ> part = lf*1000/pop
IF> shock_part_model > 0

BEHAVIORAL> part1
TSRANGE 2006 3 2020 1
EQ> TSDELTALOG(part1) = beta_0 + beta_1*TSDELTALOG(TSLAG(part,2)) + beta_2*TSDELTALOG(TSLAG(part,4)) + beta_3*TSDELTALOG(komp_r) + beta_4*TSDELTALOG(TSLAG(komp_r,3)) +beta_5*TSDELTALOG(rw_ds) + beta_6*TSDELTALOG(TSLAG(rw_ds,1)) + beta_7*TSDELTA(TSLAG(ur_ds,3)) + beta_8*LOG(TSLAG(part1,1))  + beta_9*log(TSLAG(rw_ds,1)) + beta_10*log(TSLAG(komp_r,1)) + beta_11*TSLAG(ur_ds,1) + beta_12* d_2013q4
COEFF> beta_0 beta_1 beta_2 beta_3 beta_4 beta_5 beta_6 beta_7 beta_8 beta_9 beta_10 beta_11 beta_12
STORE> coe(16)

COMMENT> Capital Stock with switches
IDENTITY> g
EQ> g = g*1
IF> shock_g_model <= 0
IDENTITY> g
EQ> g = g_shock
IF> shock_g_model > 0

COMMENT> Capital Stock with switches
IDENTITY> stat_reg
EQ> stat_reg = stat_reg*1
IF> shock_reg_model <= 0
IDENTITY> stat_reg
EQ> stat_reg = stat_reg_shock
IF> shock_reg_model > 0

COMMENT> Capital Stock with switches
IDENTITY> pf
EQ> pf = pf*1
IF> shock_1_model <= 0
IDENTITY> pf
EQ> pf = pf_shock
IF> shock_1_model > 0

COMMENT> Capital Stock with switches
IDENTITY> pm_ds
EQ> pm_ds = pm_ds*1
IF> shock_1_model <= 0
IDENTITY> pm_ds
EQ> pm_ds = pm_ds_shock
IF> shock_1_model > 0

COMMENT> Capital Stock with switches
IDENTITY> iloan
EQ> iloan = iloan*1
IF> shock_2_model <= 0
IDENTITY> iloan
EQ> iloan = iloan_shock
IF> shock_2_model > 0

COMMENT> Capital Stock with switches
IDENTITY> idep
EQ> idep = idep*1
IF> shock_2_model <= 0
IDENTITY> idep
EQ> idep = idep_shock
IF> shock_2_model > 0

COMMENT> Capital Stock with switches
IDENTITY> ibd
EQ> ibd = ibd*1
IF> shock_2_model <= 0
IDENTITY> ibd
EQ> ibd = ibd_shock
IF> shock_2_model > 0

COMMENT> Capital Stock with switches
IDENTITY> iboa
EQ> iboa = iboa*1
IF> shock_2_model <= 0
IDENTITY> iboa
EQ> iboa = iboa_shock
IF> shock_2_model > 0

COMMENT> Capital Stock with switches
IDENTITY> tax_rate1
EQ> tax_rate1 = tax_rate1*1
IF> shock_3_model <= 0
IDENTITY> tax_rate1
EQ> tax_rate1 = tax_rate1_shock
IF> shock_3_model > 0

COMMENT> Lavet for at kunne støde til g 
IDENTITY> shock_3_model
EQ> shock_3_model = shock_3_model*1

COMMENT> Lavet for at kunne støde til g 
IDENTITY> shock_2_model
EQ> shock_2_model = shock_2_model*1

COMMENT> Lavet for at kunne støde til g 
IDENTITY> shock_1_model
EQ> shock_1_model = shock_1_model*1

COMMENT> Lavet for at kunne støde til g 
IDENTITY> shock_g_model
EQ> shock_g_model = shock_g_model*1

COMMENT> Lavet for at kunne støde til g 
IDENTITY> shock_reg_model
EQ> shock_reg_model = shock_reg_model*1

COMMENT> Lavet for at kunne støde til g 
IDENTITY> shock_max_dp_model
EQ> shock_max_dp_model = shock_max_dp_model*1

COMMENT> Lavet for at kunne støde til g 
IDENTITY> shock_part_model
EQ> shock_part_model = shock_part_model*1

END"
```

```{r}
S_model=LOAD_MODEL(modelText = S_model.txt)
```

## Equations brugt til behavioural equations

```{r}
l_h_tr_ratio= -l_h_tr/(yd_hk_ds*pc)
eq_h_ratio= (eq_h-eq_h_rvx)/(TSLAG(eq_h,1) +TSLAG(sec_h,1)+ TSLAG(iba_h,1))
ratio_in_eq_h= (div_r_h+eq_h_rvx)/TSLAG(eq_h)
```

```{r}

# I think I should create the timeseries for the coeff.

S_modelData=list(

  ##Sensitivites analyse

   sens1 = TIMESERIES(c(sens1),   
                       START=c(2006,1),FREQ=4),

  ## max_dp

  stat_reg = TIMESERIES(c(stat_reg),   
                       START=c(2006,1),FREQ=4),
  stat_tilpas = TIMESERIES(c(stat_tilpas),   
                       START=c(2006,1),FREQ=4),
  tilpas_pr = TIMESERIES(c(tilpas_pr),   
                       START=c(2006,1),FREQ=4),

  ## Price adjustments
  pg_adj = TIMESERIES(c(pg_adj),   
                       START=c(2005,2),FREQ=4),
  p_bd_adj = TIMESERIES(c(p_bd_adj),   
                       START=c(2005,2),FREQ=4),
  p_equip_adj = TIMESERIES(c(p_equip_adj),   
                     START=c(2005,2),FREQ=4),
  px_adj = TIMESERIES(c(px_adj),   
                        START=c(2005,2),FREQ=4),
  ## Households

  max_dp_shock = TIMESERIES(c(max_dp_shock),   
                       START=c(2005,2),FREQ=4),
  part1 = TIMESERIES(c(part1),   
                       START=c(2005,2),FREQ=4),

  int_r_h = TIMESERIES(c(int_r_h),   
                       START=c(2005,2),FREQ=4),
  int_p_h = TIMESERIES(c(int_p_h),   
                       START=c(2005,2),FREQ=4),
  div_h = TIMESERIES(c(div_h),   
                     START=c(2005,3),FREQ=4),
  aa_check = TIMESERIES(c(aa_check),   
                        START=c(2005,2),FREQ=4),
  insu_h = TIMESERIES(c(insu_h),   
                      START=c(2005,3),FREQ=4),
  propinc_r_h = TIMESERIES(c(propinc_r_h),   
                           START=c(2005,2),FREQ=4),
  propinc_r_nf = TIMESERIES(c(propinc_r_nf ),   
                           START=c(2005,2),FREQ=4),
  propinc_p_nf= TIMESERIES(c(propinc_p_nf),   
                           START=c(2005,2),FREQ=4),
  propinc_p_h = TIMESERIES(c(propinc_p_h),   
                           START=c(2005,2),FREQ=4),
  npropinc_h = TIMESERIES(c(npropinc_h),   
                          START=c(2005,2),FREQ=4),
  yh_h = TIMESERIES(c(yh_h),   
                    START=c(2005,2),FREQ=4),
  yh2a = TIMESERIES(c(yh2a),   
                    START=c(2005,2),FREQ=4),
  yh2b = TIMESERIES(c(yh2b),   
                    START=c(2005,2),FREQ=4),
  yh1 = TIMESERIES(c(yh1),   
                   START=c(2005,2),FREQ=4),
  w_h_r = TIMESERIES(c(w_h_r),   
                     START=c(2005,2),FREQ=4),
  yd2a_h = TIMESERIES(c(yd2a_h),

                      START=c(2005,2),FREQ=4),
  yd2b_h = TIMESERIES(c(yd2b_h),

                      START=c(2005,2),FREQ=4),
  yd1_h = TIMESERIES(c(yd1_h),

                     START=c(2005,2),FREQ=4),
  tax_h = TIMESERIES(c(tax_h),

                     START=c(2005,2),FREQ=4),
  yd1_hk = TIMESERIES(c(yd1_hk),   
                      START=c(2005,2),FREQ=4),

  yd2a_hk = TIMESERIES(c(yd2a_hk),   
                       START=c(2005,2),FREQ=4),

  yd2b_hk = TIMESERIES(c(yd2b_hk),   
                       START=c(2005,2),FREQ=4),

  yd_h = TIMESERIES(c(yd_h),   
                    START=c(2005,2),FREQ=4),

  nsben_h = TIMESERIES(c(nsben_h),   
                       START=c(2005,2),FREQ=4),
  yd_h1 = TIMESERIES(c(yd_h1),   
                     START=c(2005,2),FREQ=4),

  yd_hk = TIMESERIES(c(yd_hk),   
                     START=c(2005,2),FREQ=4),
  nsben_h = TIMESERIES(c(nsben_h),   
                       START=c(2005,2),FREQ=4),

  s_h = TIMESERIES(c(s_h),   
                   START=c(2005,2),FREQ=4),

  nl_h = TIMESERIES(c(nl_h),   
                    START=c(2005,2),FREQ=4),

  pconk = TIMESERIES(c(pconk),   
                     START=c(2005,2),FREQ=4),

  pcon = TIMESERIES(c(pcon),   
                    START=c(2005,2),FREQ=4),
  i_bd_h_k_ds = TIMESERIES(c(i_bd_h_k_ds),   
                           START=c(2005,2),FREQ=4),
  i_bd_h_k = TIMESERIES(c(i_bd_h_k),   
                        START=c(2005,2),FREQ=4),
  i_equip_h_k_ds = TIMESERIES(c(i_equip_h_k_ds),   
                              START=c(2005,2),FREQ=4),
  i_equip_h_k = TIMESERIES(c(i_equip_h_k),   
                           START=c(2005,2),FREQ=4),

  propinc_p_h = TIMESERIES(c(propinc_p_h),   
                           START=c(2005,2),FREQ=4),
  nben_h = TIMESERIES(c(nben_h),   
                      START=c(2005,2),FREQ=4), ##Skal laves om!!!

  ## Financial site

  sec_h_tr_g = TIMESERIES(c(sec_h_tr_g),   
                          START=c(2005,2),FREQ=4),

  l_h = TIMESERIES(c(l_h),   
                   START=c(2005,2),FREQ=4),
  iba_h_tr = TIMESERIES(c(iba_h_tr),   
                        START=c(2005,2),FREQ=4),

  eq_h_nf = TIMESERIES(c(eq_h_nf),   
                       START=c(2005,2),FREQ=4),
  eq_h_f = TIMESERIES(c(eq_h_f),   
                      START=c(2005,2),FREQ=4),

  eq_h_row = TIMESERIES(c(eq_h_row),   
                        START=c(2005,2),FREQ=4),

  eq_h_nf_tr = TIMESERIES(c(eq_h_nf_tr),   
                          START=c(2005,2),FREQ=4),

  eq_h_f_tr = TIMESERIES(c(eq_h_f_tr),   
                         START=c(2005,2),FREQ=4),

  eq_h_row_tr = TIMESERIES(c(eq_h_row_tr),   
                           START=c(2005,2),FREQ=4),

  eq_h_test = TIMESERIES(c(eq_h_test),   
                         START=c(2005,2),FREQ=4),
  ins_h = TIMESERIES(c(ins_h),   
                     START=c(2005,2),FREQ=4),
  ins_h_tr = TIMESERIES(c(ins_h_tr),   
                        START=c(2005,2),FREQ=4),
  lev_h = TIMESERIES(c(lev_h),   
                     START=c(2005,3),FREQ=4),
  ## NET WEALTH AND FINANCIAL NET WEALTH

  fa_h = TIMESERIES(c(fa_h),   
                    START=c(2005,2),FREQ=4),

  fnw_h = TIMESERIES(c(fnw_h),   
                     START=c(2005,2),FREQ=4),
  fnw_hk = TIMESERIES(c(fnw_hk),   
                      START=c(2005,2),FREQ=4),

  ## Capital Account
  i_bd_h = TIMESERIES(c(i_bd_h),   
                      START=c(2005,2),FREQ=4),

  i_equip_h = TIMESERIES(c(i_equip_h),   
                         START=c(2005,2),FREQ=4),
  bd_h_k = TIMESERIES(c(bd_h_k),   
                      START=c(2005,2),FREQ=4),

  equip_h_k = TIMESERIES(c(equip_h_k),   
                         START=c(2005,2),FREQ=4),
  bd_h = TIMESERIES(c(bd_h),   
                    START=c(2005,2),FREQ=4),

  equip_h = TIMESERIES(c(equip_h),   
                       START=c(2005,2),FREQ=4),

  ## Non-Financial sector

  y = TIMESERIES(c(y),   
                 START=c(2005,2),FREQ=4),

  yk = TIMESERIES(c(yk),   
                  START=c(2005,2),FREQ=4),

  sk = TIMESERIES(c(sk),   
                  START=c(2005,2),FREQ=4),

  gk = TIMESERIES(c(gk),   
                  START=c(2005,2),FREQ=4),
  x = TIMESERIES(c(x),   
                 START=c(2005,2),FREQ=4),
  rer = TIMESERIES(c(rer),   
                   START=c(2005,2),FREQ=4),
  xk_ds = TIMESERIES(c(xk_ds),   
                     START=c(2005,2),FREQ=4),
  xk = TIMESERIES(c(xk),   
                  START=c(2005,2),FREQ=4),

  m = TIMESERIES(c(m),   
                 START=c(2005,2),FREQ=4),

  mk_ds = TIMESERIES(c(mk_ds),   
                     START=c(2005,2),FREQ=4),

  mk = TIMESERIES(c(mk),   
                  START=c(2005,2),FREQ=4),
  py = TIMESERIES(c(py),   
                  START=c(2005,2),FREQ=4),
  s = TIMESERIES(c(s),   
                 START=c(2005,2),FREQ=4),
  w_nf = TIMESERIES(c(w_nf),   
                    START=c(2005,2),FREQ=4),
  b2_nf_r = TIMESERIES(c(b2_nf_r),   
                       START=c(2005,2),FREQ=4),

  yf = TIMESERIES(c(yf),   
                  START=c(2005,2),FREQ=4),

  b2 = TIMESERIES(c(b2),   
                  START=c(2005,2),FREQ=4),

  ws = TIMESERIES(c(ws),   
                  START=c(2005,2),FREQ=4),
  ps = TIMESERIES(c(ps),   
                  START=c(2005,2),FREQ=4),
  u = TIMESERIES(c(u),   
                 START=c(2005,2),FREQ=4),
  prate = TIMESERIES(c(prate),   
                     START=c(2005,3),FREQ=4),
  tobinq = TIMESERIES(c(tobinq),   
                      START=c(2005,2),FREQ=4),

  i_bd_nfc_k_ds = TIMESERIES(c(i_bd_nfc_k_ds),   
                             START=c(2005,2),FREQ=4),

  i_bd_nfc_k = TIMESERIES(c(i_bd_nfc_k),   
                          START=c(2005,2),FREQ=4),

  i_equip_nfc_k = TIMESERIES(c(i_equip_nfc_k),   
                             START=c(2005,2),FREQ=4),
  ik = TIMESERIES(c(ik),   
                  START=c(2005,2),FREQ=4),
  i_h = TIMESERIES(c(i_h),   
                   START=c(2005,2),FREQ=4),
  i_nf = TIMESERIES(c(i_nf),   
                    START=c(2005,2),FREQ=4),
  i_f = TIMESERIES(c(i_f),   
                   START=c(2005,2),FREQ=4),

  i_g = TIMESERIES(c(i_g),   
                   START=c(2005,2),FREQ=4),

  i = TIMESERIES(c(i),   
                 START=c(2005,2),FREQ=4),

  i_equip_nfc_k_ds = TIMESERIES(c(i_equip_nfc_k_ds),   
                                START=c(2005,2),FREQ=4),
  int_nf = TIMESERIES(c(int_nf),   
                      START=c(2005,3),FREQ=4),
  div_nf = TIMESERIES(c(div_nf),   
                      START=c(2005,3),FREQ=4),
  insu_nf = TIMESERIES(c(insu_nf),   
                       START=c(2005,2),FREQ=4),
  npropinc_nf = TIMESERIES(c(npropinc_nf),   
                           START=c(2005,3),FREQ=4),

  yh_nf = TIMESERIES(c(yh_nf),   
                     START=c(2005,2),FREQ=4),

  s_nf = TIMESERIES(c(s_nf),   
                    START=c(2005,2),FREQ=4),

  tax_nf = TIMESERIES(c(tax_nf),   
                      START=c(2005,2),FREQ=4),
  nl_nf = TIMESERIES(c(nl_nf),   
                     START=c(2005,2),FREQ=4),

  ## Financial site

  iba_nf = TIMESERIES(c(iba_nf),   
                      START=c(2005,2),FREQ=4),
  iba_nf_tr = TIMESERIES(c(iba_nf_tr),   
                         START=c(2005,2),FREQ=4),
  sec_nf = TIMESERIES(c(sec_nf),   
                      START=c(2005,2),FREQ=4),

  l_nf = TIMESERIES(c(l_nf),   
                    START=c(2005,2),FREQ=4),

  eq_nf_a = TIMESERIES(c(eq_nf_a),   
                       START=c(2005,2),FREQ=4),

  eq_nf_a_nf = TIMESERIES(c(eq_nf_a_nf),   
                          START=c(2005,2),FREQ=4),
  eq_nf_l_nf = TIMESERIES(c(eq_nf_l_nf),   
                          START=c(2005,2),FREQ=4),
  eq_nf_ax = TIMESERIES(c(eq_nf_ax),   
                        START=c(2005,2),FREQ=4),
  eq_nf_a_f = TIMESERIES(c(eq_nf_a_f),   
                         START=c(2005,2),FREQ=4),
  eq_nf_a_row = TIMESERIES(c(eq_nf_a_row),   
                           START=c(2005,2),FREQ=4),

  eq_nf_l = TIMESERIES(c(eq_nf_l),   
                       START=c(2005,2),FREQ=4),

  neq_nf = TIMESERIES(c(Neq_nf),   
                      START=c(2005,2),FREQ=4),

  neq_nf_tr = TIMESERIES(c(Neq_nf_tr),   
                         START=c(2005,3),FREQ=4),
  eq_nf_a_test = TIMESERIES(c(eq_nf_a_test),   
                            START=c(2005,2),FREQ=4),
  eq_nf_l_test = TIMESERIES(c(eq_nf_l_test),   
                            START=c(2005,2),FREQ=4),
  ins_nf = TIMESERIES(c(ins_nf),   
                      START=c(2005,2),FREQ=4),
  fnw_nf = TIMESERIES(c(fnw_nf),   
                      START=c(2005,2),FREQ=4),
  fnw_nfk = TIMESERIES(c(fnw_nfk),   
                       START=c(2005,2),FREQ=4),

  ## Capital account

  i_bd_nfc = TIMESERIES(c(i_bd_nfc),   
                        START=c(2005,2),FREQ=4),
  i_equip_nfc = TIMESERIES(c(i_equip_nfc),   
                           START=c(2005,2),FREQ=4),
  bd_nfc_k = TIMESERIES(c(bd_nfc_k),   
                        START=c(2005,2),FREQ=4),
  equip_nfc_k = TIMESERIES(c(equip_nfc_k),   
                           START=c(2005,2),FREQ=4),
  bd_nfc = TIMESERIES(c(bd_nfc),   
                      START=c(2005,2),FREQ=4),
  equip_nfc = TIMESERIES(c(equip_nfc),   
                         START=c(2005,2),FREQ=4),

  ## Financial corp

  int_f = TIMESERIES(c(int_f),   
                     START=c(2005,3),FREQ=4),
  div_f = TIMESERIES(c(div_f),   
                     START=c(2005,3),FREQ=4),
  insu_f = TIMESERIES(c(insu_f),   
                      START=c(2005,2),FREQ=4),

  propinc_r_f = TIMESERIES(c(propinc_r_f),   
                           START=c(2005,2),FREQ=4),

  propinc_p_f = TIMESERIES(c(propinc_p_f),   
                           START=c(2005,2),FREQ=4),

  npropinc_f = TIMESERIES(c(npropinc_f),   
                          START=c(2005,3),FREQ=4),
  yh_f = TIMESERIES(c(yh_f),   
                    START=c(2005,2),FREQ=4),
  yd_f = TIMESERIES(c(yd_f),   
                    START=c(2005,2),FREQ=4),
  d8_f = TIMESERIES(c(d8_f),   
                    START=c(2005,2),FREQ=4),
  nben_f = TIMESERIES(c(nben_f),   
                      START=c(2005,2),FREQ=4),

  s_f = TIMESERIES(c(s_f),   
                   START=c(2005,2),FREQ=4),

  nl_f = TIMESERIES(c(nl_f),   
                    START=c(2005,2),FREQ=4),

  iba_f = TIMESERIES(c(iba_f),   
                     START=c(2005,2),FREQ=4),
  iba_f_tr = TIMESERIES(c(iba_f_tr),   
                        START=c(2005,2),FREQ=4),
  eq_f_a = TIMESERIES(c(eq_f_a),   
                      START=c(2005,2),FREQ=4),
  eq_f_a_nf = TIMESERIES(c(eq_f_a_nf),   
                         START=c(2005,2),FREQ=4),
  eq_f_a_f = TIMESERIES(c(eq_f_a_f),   
                        START=c(2005,2),FREQ=4),
  eq_f_l_f = TIMESERIES(c(eq_f_l_f),   
                        START=c(2005,2),FREQ=4),

  eq_f_a_row = TIMESERIES(c(eq_f_a_row),   
                          START=c(2005,2),FREQ=4),

  eq_f_ax = TIMESERIES(c(eq_f_ax),   
                       START=c(2005,2),FREQ=4),
  eq_f_l = TIMESERIES(c(eq_f_l),   
                      START=c(2005,2),FREQ=4),
  eq_f_lx = TIMESERIES(c(eq_f_lx),   
                       START=c(2005,2),FREQ=4),
  neq_f = TIMESERIES(c(Neq_f),   
                     START=c(2005,2),FREQ=4),
  neq_f_tr = TIMESERIES(c(Neq_f_tr),   
                        START=c(2005,3),FREQ=4),
  eq_f_l_test = TIMESERIES(c(eq_f_l_test),   
                           START=c(2005,2),FREQ=4),
  sec_f_d = TIMESERIES(c(sec_f_d),   
                       START=c(2005,2),FREQ=4),
  sec_f_d_tr = TIMESERIES(c(sec_f_d_tr),   
                          START=c(2005,2),FREQ=4),

  sec_f_d_tr_nf = TIMESERIES(c(sec_f_d_tr_nf),   
                             START=c(2005,2),FREQ=4),
  sec_f_d_tr_g = TIMESERIES(c(sec_f_d_tr_g),   
                            START=c(2005,2),FREQ=4),
  sec_f_a = TIMESERIES(c(sec_f_a),   
                       START=c(2005,2),FREQ=4),
  sec_f_a_tr = TIMESERIES(c(sec_f_a_tr),   
                          START=c(2005,2),FREQ=4),

  l_f = TIMESERIES(c(l_f),   
                   START=c(2005,2),FREQ=4),

  l_f_tr = TIMESERIES(c(l_f_tr),   
                      START=c(2005,2),FREQ=4),

  ins_f = TIMESERIES(c(ins_f),   
                     START=c(2005,2),FREQ=4),

  ins_f_tr = TIMESERIES(c(ins_f_tr),   
                        START=c(2005,2),FREQ=4),
  fnw_f = TIMESERIES(c(fnw_f),   
                     START=c(2005,2),FREQ=4),
  fnw_fk = TIMESERIES(c(fnw_fk),   
                      START=c(2005,2),FREQ=4),
  i_bd_fc = TIMESERIES(c(i_bd_fc),   
                       START=c(2005,2),FREQ=4),
  i_equip_fc = TIMESERIES(c(i_equip_fc),   
                          START=c(2005,2),FREQ=4),
  bd_fc_k = TIMESERIES(c(bd_fc_k),   
                       START=c(2005,2),FREQ=4),
  equip_fc_k = TIMESERIES(c(equip_fc_k),   
                          START=c(2005,2),FREQ=4),
  bd_fc = TIMESERIES(c(bd_fc),   
                     START=c(2005,2),FREQ=4),

  equip_fc = TIMESERIES(c(equip_fc),   
                        START=c(2005,2),FREQ=4),

  ## Govt

  int_g = TIMESERIES(c(int_g),   
                     START=c(2005,3),FREQ=4),

  div_g = TIMESERIES(c(div_g),   
                     START=c(2005,3),FREQ=4),
  insu_g = TIMESERIES(c(insu_g),   
                      START=c(2005,2),FREQ=4),
  propinc_r_g = TIMESERIES(c(propinc_r_g),   
                           START=c(2005,2),FREQ=4),
  propinc_p_g = TIMESERIES(c(propinc_p_g),   
                           START=c(2005,2),FREQ=4),
  npropinc_g = TIMESERIES(c(npropinc_g),   
                          START=c(2005,3),FREQ=4),
  yh_g = TIMESERIES(c(yh_g),   
                    START=c(2005,2),FREQ=4),
  yd_g = TIMESERIES(c(yd_g),   
                    START=c(2005,2),FREQ=4),
  nben_g = TIMESERIES(c(nben_g),   
                      START=c(2005,2),FREQ=4),

  tax_g = TIMESERIES(c(tax_g),   
                     START=c(2005,2),FREQ=4),
  s_g = TIMESERIES(c(s_g),   
                   START=c(2005,2),FREQ=4),
  nl_g = TIMESERIES(c(nl_g),   
                    START=c(2005,2),FREQ=4),
  iba_g = TIMESERIES(c(iba_g),   
                     START=c(2005,2),FREQ=4),

  eq_g = TIMESERIES(c(eq_g),   
                    START=c(2005,2),FREQ=4),

  eq_g_nf = TIMESERIES(c(eq_g_nf),   
                       START=c(2005,2),FREQ=4),

  eq_g_f = TIMESERIES(c(eq_g_f),   
                      START=c(2005,2),FREQ=4),
  eq_g_row = TIMESERIES(c(eq_g_row),   
                        START=c(2005,2),FREQ=4),

  eq_g_ax = TIMESERIES(c(eq_g_ax),   
                       START=c(2005,2),FREQ=4),
  eq_g_test = TIMESERIES(c(eq_g_test),   
                         START=c(2005,2),FREQ=4),
  sec_g = TIMESERIES(c(sec_g),   
                     START=c(2005,2),FREQ=4),

  sec_g_tr = TIMESERIES(c(sec_g_tr),   
                        START=c(2005,2),FREQ=4),

  l_g = TIMESERIES(c(l_g),   
                   START=c(2005,2),FREQ=4),

  ins_g = TIMESERIES(c(ins_g),   
                     START=c(2005,2),FREQ=4),

  fnw_g = TIMESERIES(c(fnw_g),   
                     START=c(2005,2),FREQ=4),
  fnw_gk = TIMESERIES(c(fnw_gk),   
                      START=c(2005,2),FREQ=4),

  i_bd_g  = TIMESERIES(c(i_bd_g ),   
                       START=c(2005,2),FREQ=4),

  i_equip_g = TIMESERIES(c(i_equip_g),   
                         START=c(2005,2),FREQ=4),
  bd_g_k = TIMESERIES(c(bd_g_k),   
                      START=c(2005,2),FREQ=4),
  equip_g_k = TIMESERIES(c(equip_g_k),   
                         START=c(2005,2),FREQ=4),

  bd_g = TIMESERIES(c(bd_g),   
                    START=c(2005,2),FREQ=4),
  equip_g = TIMESERIES(c(equip_g),   
                       START=c(2005,2),FREQ=4),

  ## ROW

  int_row = TIMESERIES(c(int_row),   
                       START=c(2005,3),FREQ=4),

  div_row = TIMESERIES(c(div_row),   
                       START=c(2005,3),FREQ=4),
  insu_row = TIMESERIES(c(insu_row),   
                        START=c(2005,2),FREQ=4),
  propinc_r_row = TIMESERIES(c(propinc_r_row),   
                             START=c(2005,2),FREQ=4),
  propinc_p_row = TIMESERIES(c(propinc_p_row),   
                             START=c(2005,2),FREQ=4),
  npropinc_row = TIMESERIES(c(npropinc_row),   
                            START=c(2005,3),FREQ=4),
  s_row = TIMESERIES(c(s_row),   
                     START=c(2005,2),FREQ=4),
  nl_row = TIMESERIES(c(nl_row),   
                      START=c(2005,3),FREQ=4),

  nben_row = TIMESERIES(c(nben_row),   
                        START=c(2005,2),FREQ=4),

  nx = TIMESERIES(c(nx),   
                  START=c(2005,2),FREQ=4),

  cab = TIMESERIES(c(cab),   
                   START=c(2005,3),FREQ=4),
  bop = TIMESERIES(c(bop),   
                   START=c(2005,3),FREQ=4),
  fab = TIMESERIES(c(fab),   
                   START=c(2005,3),FREQ=4),
  iba_row = TIMESERIES(c(iba_row),   
                       START=c(2005,2),FREQ=4),
  iba_row_tr = TIMESERIES(c(iba_row_tr),   
                          START=c(2005,2),FREQ=4),
  sec_row = TIMESERIES(c(sec_row),   
                       START=c(2005,2),FREQ=4),
  l_row = TIMESERIES(c(l_row),   
                     START=c(2005,2),FREQ=4),
  eq_row_a = TIMESERIES(c(eq_row_a),   
                        START=c(2005,2),FREQ=4),

  eq_row_a_nf = TIMESERIES(c(eq_row_a_nf),   
                           START=c(2005,2),FREQ=4),
  eq_row_a_f = TIMESERIES(c(eq_row_a_f),   
                          START=c(2005,2),FREQ=4),
  eq_row_ax = TIMESERIES(c(eq_row_ax),   
                         START=c(2005,2),FREQ=4),
  eq_row_a_test = TIMESERIES(c(eq_row_a_test),   
                             START=c(2005,2),FREQ=4),

  eq_row_l = TIMESERIES(c(eq_row_l),   
                        START=c(2005,2),FREQ=4),

  eq_row_lx = TIMESERIES(c(eq_row_lx),   
                         START=c(2005,2),FREQ=4),

  eq_row_l_test = TIMESERIES(c(eq_row_l_test),   
                             START=c(2005,2),FREQ=4),
  neq_row = TIMESERIES(c(Neq_row),   
                       START=c(2005,2),FREQ=4),

  neq_row_tr = TIMESERIES(c(Neq_row_tr),   
                          START=c(2005,3),FREQ=4),
  ins_row = TIMESERIES(c(ins_row),   
                       START=c(2005,2),FREQ=4),
  fnw_row = TIMESERIES(c(fnw_row),   
                       START=c(2005,2),FREQ=4),

  fnw_rowk = TIMESERIES(c(fnw_rowk),   
                        START=c(2005,2),FREQ=4),

  ## Labour market
  unadj = TIMESERIES(c(unadj),   
                     START=c(2005,2),FREQ=4),

  unemp = TIMESERIES(c(unemp),   
                     START=c(2005,2),FREQ=4),
  ur = TIMESERIES(c(ur),   
                  START=c(2005,2),FREQ=4),
  uradj = TIMESERIES(c(uradj),   
                     START=c(2005,2),FREQ=4),
  lf = TIMESERIES(c(LF),   
                  START=c(2005,2),FREQ=4),
  pc_ds = TIMESERIES(c(pc_ds),   
                     START=c(2005,2),FREQ=4),
  pc = TIMESERIES(c(pc),   
                  START=c(2005,2),FREQ=4),
  mkp = TIMESERIES(c(mkp),   
                   START=c(2005,2),FREQ=4),

  wage_ds = TIMESERIES(c(wage_ds),   
                       START=c(2005,2),FREQ=4),
  wage_ds_t = TIMESERIES(c(wage_ds_t),   
                       START=c(2006,2),FREQ=4),
  inflation = TIMESERIES(c(inflation),   
                       START=c(2006,2),FREQ=4),
  inflation_t = TIMESERIES(c(inflation_t),   
                       START=c(2005,3),FREQ=4),
   inflation_tt = TIMESERIES(c(inflation_tt),   
                       START=c(2005,3),FREQ=4),
  pm_ds = TIMESERIES(c(pm_ds),   
                     START=c(2005,2),FREQ=4),
  mk_ds = TIMESERIES(c(mk_ds),   
                     START=c(2005,2),FREQ=4),

  emp = TIMESERIES(c(emp),   
                   START=c(2005,2),FREQ=4),

  urterm = TIMESERIES(c(urterm),   
                      START=c(2005,2),FREQ=4),

  rw_ds = TIMESERIES(c(rw_ds),   
                     START=c(2005,2),FREQ=4),
  wage = TIMESERIES(c(wage1),   
                    START=c(2003,1),FREQ=4),

  wageindex = TIMESERIES(c(wageindex),   
                         START=c(2005,2),FREQ=4),
  mkp = TIMESERIES(c(mkp),   
                   START=c(2005,2),FREQ=4),

  prod = TIMESERIES(c(prod),   
                    START=c(2005,2),FREQ=4),
  acc_rate = TIMESERIES(c(acc_rate),   
                        START=c(2005,2),FREQ=4),

   nben_h_rest = TIMESERIES(c(nben_h_rest_obs),   
                       START=c(2005,2),FREQ=4),
  nben_h_un = TIMESERIES(c(nben_h_un_obs),   
                        START=c(2005,2),FREQ=4),
  komp_r = TIMESERIES(c(komp_r_obs),   
                        START=c(2005,2),FREQ=4),
  shock = TIMESERIES(c(shock),   
                        START=c(2005,2),FREQ=4),
  g_shock = TIMESERIES(c(g_shock),   
                        START=c(2005,2),FREQ=4),
  stat_reg_shock = TIMESERIES(c(stat_reg_shock),   
                        START=c(2005,2),FREQ=4),

  wage_trim = TIMESERIES(c(wage_trim1),   
                        START=c(2005,2),FREQ=4),
   dp_person = TIMESERIES(c(mean_dp_regul),   
                        START=c(2005,2),FREQ=4),
  max_dp = TIMESERIES(c(obs_dp_q),   
                        START=c(2005,2),FREQ=4),
  kuld = TIMESERIES(c(kuld),   
                        START=c(2005,2),FREQ=4),
  abm = TIMESERIES(c(abm),   
                        START=c(2005,2),FREQ=4),
  w_h_r_trim = TIMESERIES(c(w_h_r_trim1),   
                        START=c(2005,2),FREQ=4),

  ## Dummies

  d_2006q1 = TIMESERIES(c(d_2006q1),   
                        START=c(2005,2),FREQ=4),

  d_2006q2 = TIMESERIES(c(d_2006q2),   
                        START=c(2005,2),FREQ=4),
  d_2006q3 = TIMESERIES(c(d_2006q3),   
                        START=c(2005,2),FREQ=4),
  d_2006q4 = TIMESERIES(c(d_2006q4),   
                        START=c(2005,2),FREQ=4),
  d_2007q34 = TIMESERIES(c(d_2007q34),   
                         START=c(2005,2),FREQ=4),
  d_2008q2 = TIMESERIES(c(d_2008q2),   
                        START=c(2005,2),FREQ=4),
  d_2008q3 = TIMESERIES(c(d_2008q3),   
                        START=c(2005,2),FREQ=4),
  d_2008q4 = TIMESERIES(c(d_2008q4),   
                        START=c(2005,2),FREQ=4),

  d_2009q4 = TIMESERIES(c(d_2009q4),   
                        START=c(2005,2),FREQ=4),
  d_2018q2 = TIMESERIES(c(d_2018q2),   
                        START=c(2005,2),FREQ=4),
  d_20123 = TIMESERIES(c(d_20123),   
                       START=c(2005,2),FREQ=4),
  d_20124 = TIMESERIES(c(d_20124),   
                       START=c(2005,2),FREQ=4),

  d_2009q1 = TIMESERIES(c(d_2009q1),   
                        START=c(2005,2),FREQ=4),

  d_2011q1 = TIMESERIES(c(d_2011q1),   
                        START=c(2005,2),FREQ=4),

  d_2014q3 = TIMESERIES(c(d_2014q3),   
                        START=c(2005,2),FREQ=4),
  d_2014q4 = TIMESERIES(c(d_2014q4),   
                        START=c(2005,2),FREQ=4),

  d_2013q4 = TIMESERIES(c(d_2013q4),   
                        START=c(2005,2),FREQ=4),
  d_2018q1 = TIMESERIES(c(d_2018q1),   
                        START=c(2005,2),FREQ=4),

  d_2016q3 = TIMESERIES(c(d_2016q3),   
                        START=c(2005,2),FREQ=4),

  d_2019q3 = TIMESERIES(c(d_2019q3),   
                        START=c(2005,2),FREQ=4),

  d_2019q4 = TIMESERIES(c(d_2019q4),   
                        START=c(2005,2),FREQ=4),
  d_2020q1 = TIMESERIES(c(d_2020q1),   
                        START=c(2005,2),FREQ=4),
  d_2017q1 = TIMESERIES(c(d_2017q1),   
                        START=c(2005,2),FREQ=4),
  d_2013q4 = TIMESERIES(c(d_2013q4),   
                        START=c(2005,2),FREQ=4),
  dummy_2018q34 = TIMESERIES(c(dummy_2018q34),   
                             START=c(2005,2),FREQ=4),
  dummy_2011q1 = TIMESERIES(c(dummy_2011q1),   
                            START=c(2005,2),FREQ=4),

  dummy_2009q2 = TIMESERIES(c(dummy_2009q2),   
                            START=c(2005,2),FREQ=4),
  time = TIMESERIES(c(time),   
                    START=c(2005,2),FREQ=4),

  dummy_4 = TIMESERIES(c(DUMMY_4),   
                       START=c(2005,2),FREQ=4),

  dummy_10 = TIMESERIES(c(DUMMY_10),   
                        START=c(2005,2),FREQ=4),

  dummy_11 = TIMESERIES(c(DUMMY_11),   
                        START=c(2005,2),FREQ=4),

  ## Deseason variables

  alpha_00 = TIMESERIES(c(alpha_00),   
                       START=c(2005,2),FREQ=4),

  alpha_01 = TIMESERIES(c(alpha_01),   
                       START=c(2005,2),FREQ=4),
  alpha_02 = TIMESERIES(c(alpha_02),   
                       START=c(2005,2),FREQ=4),
  alpha_03 = TIMESERIES(c(alpha_03),   
                       START=c(2005,2),FREQ=4),
  alpha_10 = TIMESERIES(c(alpha_10),   
                        START=c(2005,2),FREQ=4),
  alpha_11 = TIMESERIES(c(alpha_11),   
                        START=c(2005,2),FREQ=4),
  alpha_12 = TIMESERIES(c(alpha_12),   
                        START=c(2005,2),FREQ=4),

  alpha_13 = TIMESERIES(c(alpha_13),   
                        START=c(2005,2),FREQ=4),
  mean_i_bd_h_k = TIMESERIES(c(mean_i_bd_h_k),   
                             START=c(2005,2),FREQ=4),
  alpha_20 = TIMESERIES(c(alpha_20 ),   
                        START=c(2005,2),FREQ=4),
  alpha_21 = TIMESERIES(c(alpha_21),   
                        START=c(2005,2),FREQ=4),

  alpha_22 = TIMESERIES(c(alpha_22),   
                        START=c(2005,2),FREQ=4),

  alpha_23 = TIMESERIES(c(alpha_23),   
                        START=c(2005,2),FREQ=4),

  mean_i_equip_h_k = TIMESERIES(c(mean_i_equip_h_k),   
                                START=c(2005,2),FREQ=4),
  alpha_30 = TIMESERIES(c(alpha_30),   
                        START=c(2005,2),FREQ=4),

  alpha_31 = TIMESERIES(c(alpha_31),   
                        START=c(2005,2),FREQ=4),
  alpha_32 = TIMESERIES(c(alpha_32),   
                        START=c(2005,2),FREQ=4),

  alpha_33 = TIMESERIES(c(alpha_33),   
                        START=c(2005,2),FREQ=4),

  mean_i_bd_nfc_k = TIMESERIES(c(mean_i_bd_nfc_k),   
                               START=c(2005,2),FREQ=4),

  alpha_40 = TIMESERIES(c(alpha_40),   
                        START=c(2005,2),FREQ=4),
  alpha_41 = TIMESERIES(c(alpha_41),   
                        START=c(2005,2),FREQ=4),
  alpha_42 = TIMESERIES(c(alpha_42),   
                        START=c(2005,2),FREQ=4),
  alpha_43 = TIMESERIES(c(alpha_43),   
                        START=c(2005,2),FREQ=4),

  mean_i_equip_nfc_k = TIMESERIES(c(mean_i_equip_nfc_k),   
                                  START=c(2005,2),FREQ=4),
  alpha_50 = TIMESERIES(c(alpha_50),   
                        START=c(2005,2),FREQ=4),

  alpha_51 = TIMESERIES(c(alpha_51),   
                        START=c(2005,2),FREQ=4),

  alpha_52 = TIMESERIES(c(alpha_52),   
                        START=c(2005,2),FREQ=4),

  alpha_53 = TIMESERIES(c(alpha_53),   
                        START=c(2005,2),FREQ=4),
  mean_wage = TIMESERIES(c(mean_wage),   
                         START=c(2005,2),FREQ=4),

  alpha_60 = TIMESERIES(c(alpha_60),   
                        START=c(2005,2),FREQ=4),
  alpha_61 = TIMESERIES(c(alpha_61),   
                        START=c(2005,2),FREQ=4),
  alpha_62 = TIMESERIES(c(alpha_62),   
                        START=c(2005,2),FREQ=4),
  alpha_63 = TIMESERIES(c(alpha_63),   
                        START=c(2005,2),FREQ=4),
  mean_pc = TIMESERIES(c(mean_pc),   
                       START=c(2005,2),FREQ=4),
  alpha_70 = TIMESERIES(c(alpha_70),   
                        START=c(2005,2),FREQ=4),
  alpha_71 = TIMESERIES(c(alpha_71),   
                        START=c(2005,2),FREQ=4),

  alpha_72 = TIMESERIES(c(alpha_72),   
                        START=c(2005,2),FREQ=4),
  alpha_73 = TIMESERIES(c(alpha_73),   
                        START=c(2005,2),FREQ=4),
  mean_mk = TIMESERIES(c(mean_mk),   
                       START=c(2005,2),FREQ=4),
  alpha_80 = TIMESERIES(c(alpha_80),   
                        START=c(2005,2),FREQ=4),

  alpha_81 = TIMESERIES(c(alpha_81),   
                        START=c(2005,2),FREQ=4),

  alpha_82 = TIMESERIES(c(alpha_82),   
                        START=c(2005,2),FREQ=4),

  alpha_83 = TIMESERIES(c(alpha_83),   
                        START=c(2005,2),FREQ=4),
  mean_xk = TIMESERIES(c(mean_xk),   
                       START=c(2005,2),FREQ=4),

   alpha_90 = TIMESERIES(c(alpha_90),   
                        START=c(2005,2),FREQ=4),

  alpha_91 = TIMESERIES(c(alpha_91),   
                        START=c(2005,2),FREQ=4),

  alpha_92 = TIMESERIES(c(alpha_92),   
                        START=c(2005,2),FREQ=4),

  alpha_93 = TIMESERIES(c(alpha_93),   
                        START=c(2005,2),FREQ=4),
  mean_yd1_hk = TIMESERIES(c(mean_yd1_hk),   
                       START=c(2005,2),FREQ=4),

   alpha_100 = TIMESERIES(c(alpha_100),   
                        START=c(2005,2),FREQ=4),

  alpha_101 = TIMESERIES(c(alpha_101),   
                        START=c(2005,2),FREQ=4),

  alpha_102 = TIMESERIES(c(alpha_102),   
                        START=c(2005,2),FREQ=4),

  alpha_103 = TIMESERIES(c(alpha_103),   
                        START=c(2005,2),FREQ=4),
  mean_prate = TIMESERIES(c(mean_prate),   
                       START=c(2005,2),FREQ=4),

   alpha_120 = TIMESERIES(c(alpha_120),   
                        START=c(2005,2),FREQ=4),

  alpha_121 = TIMESERIES(c(alpha_121),   
                        START=c(2005,2),FREQ=4),

  alpha_122 = TIMESERIES(c(alpha_122),   
                        START=c(2005,2),FREQ=4),

  alpha_123 = TIMESERIES(c(alpha_123),   
                        START=c(2005,2),FREQ=4),
   mean_ps = TIMESERIES(c(mean_ps),   
                       START=c(2005,2),FREQ=4),

   alpha_130 = TIMESERIES(c(alpha_130),   
                        START=c(2005,2),FREQ=4),

  alpha_131 = TIMESERIES(c(alpha_131),   
                        START=c(2005,2),FREQ=4),

  alpha_132 = TIMESERIES(c(alpha_132),   
                        START=c(2005,2),FREQ=4),

  alpha_133 = TIMESERIES(c(alpha_133),   
                        START=c(2005,2),FREQ=4),
   mean_ur = TIMESERIES(c(mean_ur),   
                       START=c(2005,2),FREQ=4),

   alpha_140 = TIMESERIES(c(alpha_140),   
                        START=c(2005,2),FREQ=4),

  alpha_141 = TIMESERIES(c(alpha_141),   
                        START=c(2005,2),FREQ=4),

  alpha_142 = TIMESERIES(c(alpha_142),   
                        START=c(2005,2),FREQ=4),

  alpha_143 = TIMESERIES(c(alpha_143),   
                        START=c(2005,2),FREQ=4),
   mean_yd2a_hk = TIMESERIES(c(mean_yd2a_hk),   
                       START=c(2005,2),FREQ=4),

   alpha_110 = TIMESERIES(c(alpha_110),   
                        START=c(2005,2),FREQ=4),

  alpha_111 = TIMESERIES(c(alpha_111),   
                        START=c(2005,2),FREQ=4),

  alpha_112 = TIMESERIES(c(alpha_112),   
                        START=c(2005,2),FREQ=4),

  alpha_113 = TIMESERIES(c(alpha_113),   
                        START=c(2005,2),FREQ=4),
  mean_yk = TIMESERIES(c(mean_yk),   
                       START=c(2005,2),FREQ=4),
  mean_pconk = TIMESERIES(c(mean_pconk),   
                       START=c(2005,2),FREQ=4),

  yd_hk_ds = TIMESERIES(c(yd_hk_ds),   
                        START=c(2005,2),FREQ=4),
  i_equip_h_k_ds = TIMESERIES(c(i_equip_h_k_ds),   
                              START=c(2005,2),FREQ=4),
  i_bd_nfc_k_ds = TIMESERIES(c(i_bd_nfc_k_ds),   
                             START=c(2005,2),FREQ=4),
  u_ds = TIMESERIES(c(u_ds),   
                    START=c(2005,2),FREQ=4),
  ps_ds = TIMESERIES(c(ps_ds),   
                     START=c(2005,2),FREQ=4),
  i_equip_nfc_k_ds = TIMESERIES(c(i_equip_nfc_k_ds),   
                                START=c(2005,2),FREQ=4),

  pconk_ds = TIMESERIES(c(pconk_ds),   
                        START=c(2005,2),FREQ=4),
  b2_hk_ds = TIMESERIES(c(b2_hk_ds),   
                        START=c(2005,2),FREQ=4),
  yd1_hk_ds = TIMESERIES(c(yd1_hk_ds ),   
                         START=c(2005,2),FREQ=4),
  yd2a_hk_ds = TIMESERIES(c(yd2a_hk_ds),   
                          START=c(2005,2),FREQ=4),

  w_h_rk_ds = TIMESERIES(c(w_h_rk_ds),   
                         START=c(2005,2),FREQ=4),

  pc_ds = TIMESERIES(c(pc_ds),   
                     START=c(2005,2),FREQ=4),
  pm_ds = TIMESERIES(c(pm_ds),   
                     START=c(2005,2),FREQ=4),
  ur_ds = TIMESERIES(c(ur_ds),   
                     START=c(2005,2),FREQ=4),
  yk_ds = TIMESERIES(c(yk_ds),   
                     START=c(2005,2),FREQ=4),

  wageindex_ds = TIMESERIES(c(wageindex_ds),   
                            START=c(2005,2),FREQ=4),

  priv_ds = TIMESERIES(c(priv_ds),   
                       START=c(2005,2),FREQ=4),

  mk_ds = TIMESERIES(c(mk_ds),   
                     START=c(2005,2),FREQ=4),
  prodk_ds = TIMESERIES(c(prodk_ds),   
                        START=c(2005,2),FREQ=4),
  prate_ds = TIMESERIES(c(prate_ds),   
                        START=c(2005,3),FREQ=4),
  bd_nfc_k_ds = TIMESERIES(c(bd_nfc_k_ds),   
                           START=c(2005,2),FREQ=4),

  xk_ds = TIMESERIES(c(xk_ds),   
                     START=c(2005,2),FREQ=4),
  rer_ds = TIMESERIES(c(rer_ds),   
                      START=c(2005,2),FREQ=4),

  prod_ds = TIMESERIES(c(prod_ds),   
                       START=c(2005,2),FREQ=4),

  ps_ds = TIMESERIES(c(ps_ds),   
                     START=c(2005,2),FREQ=4),

  y_ds = TIMESERIES(c(y_ds),   
                    START=c(2005,2),FREQ=4),

  SD1 = TIMESERIES(c(sd1),   
                       START=c(2005,2),FREQ=4),

  SD2 = TIMESERIES(c(sd2),   
                     START=c(2005,2),FREQ=4),

  SD3 = TIMESERIES(c(sd3),   
                    START=c(2005,2),FREQ=4),
  SD4 = TIMESERIES(c(sd4),   
                    START=c(2005,2),FREQ=4),

  ## globaldemand1

  oecd = TIMESERIES(c(oecd),   
                    START=c(2005,2),FREQ=4),
  oecde = TIMESERIES(c(oecde),   
                     START=c(2005,2),FREQ=4),

  g_20 = TIMESERIES(c(g_20),   
                    START=c(2005,2),FREQ=4),
  ea19 = TIMESERIES(c(ea19),   
                    START=c(2005,2),FREQ=4),
  eu = TIMESERIES(c(eu),   
                  START=c(2005,2),FREQ=4),
  gdp_tp = TIMESERIES(c(gdp_tp),   
                      START=c(2005,2),FREQ=4),
  rer_old = TIMESERIES(c(rer_old),   
                       START=c(2005,2),FREQ=4),
  xr = TIMESERIES(c(xr),   
                  START=c(2005,2),FREQ=4),
  p_trade = TIMESERIES(c(p_trade),   
                       START=c(2005,2),FREQ=4),

  ## Exogene

  p_bd = TIMESERIES(c(p_bd),   
                    START=c(2005,2),FREQ=4),
  pi_1 = TIMESERIES(c(pi_1),   
                  START=c(2005,2),FREQ=4),
  zaland_jesper = TIMESERIES(c(zaland_jesper ),   
                             START=c(2005,2),FREQ=4),
  ibd = TIMESERIES(c(ibd),   
                   START=c(2005,2),FREQ=4),

  iba_h = TIMESERIES(c(iba_h),   
                     START=c(2005,2),FREQ=4),

  sec_h = TIMESERIES(c(sec_h),   
                     START=c(2005,2),FREQ=4),

  div_r_h = TIMESERIES(c(div_r_h),   
                       START=c(2005,2),FREQ=4),
  eq_h_rvx = TIMESERIES(c(eq_h_rvx),   
                        START=c(2005,3),FREQ=4),

  old_age_ratio = TIMESERIES(c(old_age_ratio),   
                             START=c(2005,2),FREQ=4),
  iloan = TIMESERIES(c(iloan),   
                     START=c(2005,2),FREQ=4),
  l_h_tr = TIMESERIES(c(l_h_tr),   
                      START=c(2005,2),FREQ=4),
  l_h_tr_ratio = TIMESERIES(c(l_h_tr_ratio),   
                      START=c(2005,2),FREQ=4),
  eq_h = TIMESERIES(c(eq_h),   
                    START=c(2005,2),FREQ=4),
  d8_h = TIMESERIES(c(d8_h),   
                    START=c(2005,2),FREQ=4),
  eq_h_tr = TIMESERIES(c(eq_h_tr),   
                       START=c(2005,3),FREQ=4),
  eq_nf_lx = TIMESERIES(c(eq_nf_lx),   
                        START=c(2005,2),FREQ=4),
  idep = TIMESERIES(c(idep),   
                    START=c(2005,2),FREQ=4),
  int_r_h_adj = TIMESERIES(c(int_r_h_adj),   
                           START=c(2005,3),FREQ=4),
  int_p_h_adj = TIMESERIES(c(int_p_h_adj),   
                           START=c(2005,3),FREQ=4),
  div_h_adj = TIMESERIES(c(div_h_adj),   
                         START=c(2005,3),FREQ=4),
  diva = TIMESERIES(c(diva),   
                    START=c(2005,2),FREQ=4),
  divd = TIMESERIES(c(divd),   
                    START=c(2005,2),FREQ=4),

  ins_h_adj = TIMESERIES(c(ins_h_adj),   
                    START=c(2005,3),FREQ=4),
  res_r_h = TIMESERIES(c(res_r_h),   
                  START=c(2005,2),FREQ=4),
  res_p_h = TIMESERIES(c(res_p_h ),   
                             START=c(2005,2),FREQ=4),

  oth_h = TIMESERIES(c(oth_h),   
                     START=c(2005,2),FREQ=4),

  tax_rate2 = TIMESERIES(c(tax_rate2),   
                     START=c(2005,2),FREQ=4),

  tax_rate1 = TIMESERIES(c(tax_rate1),   
                       START=c(2005,2),FREQ=4),
  tax_h_adj = TIMESERIES(c(tax_h_adj),   
                        START=c(2005,2),FREQ=4),

  np_h = TIMESERIES(c(np_h),   
                             START=c(2005,2),FREQ=4),
  ctr_h = TIMESERIES(c(ctr_h),   
                     START=c(2005,2),FREQ=4),
  iba_h_rvx = TIMESERIES(c(iba_h_rvx),   
                      START=c(2005,3),FREQ=4),
  sec_h_tr = TIMESERIES(c(sec_h_tr),   
                    START=c(2005,2),FREQ=4),
  sec_h_rvx = TIMESERIES(c(sec_h_rvx),   
                    START=c(2005,3),FREQ=4),
  l_h_rvx = TIMESERIES(c(l_h_rvx),   
                       START=c(2005,3),FREQ=4),
  sec_h_tr = TIMESERIES(c(sec_h_tr),   
                        START=c(2005,2),FREQ=4),
  nl_h_adj = TIMESERIES(c(nl_h_adj),   
                    START=c(2005,3),FREQ=4),
  alpha_nf = TIMESERIES(c(alpha_nf),   
                           START=c(2005,2),FREQ=4),
  alpha_f = TIMESERIES(c(alpha_f),   
                           START=c(2005,2),FREQ=4),
  alpha_row = TIMESERIES(c(alpha_row),   
                         START=c(2005,2),FREQ=4),
  eq_h_nf_rvx = TIMESERIES(c(eq_h_nf_rvx),   
                    START=c(2005,3),FREQ=4),
  eq_h_row_rvx = TIMESERIES(c(eq_h_row_rvx),   
                    START=c(2005,3),FREQ=4),

  ins_h_rvx = TIMESERIES(c(ins_h_rvx),   
                         START=c(2005,3),FREQ=4),
  ins_h_tr_excl_d8 = TIMESERIES(c(ins_h_tr_excl_d8),   
                       START=c(2005,2),FREQ=4),
  p_equip = TIMESERIES(c(p_equip ),   
                       START=c(2005,2),FREQ=4),
  delta_bd_h = TIMESERIES(c(delta_bd_h),   
                      START=c(2005,2),FREQ=4),

  delta_equip_h = TIMESERIES(c(delta_equip_h),   
                     START=c(2005,2),FREQ=4),

  w_row_r = TIMESERIES(c(w_row_r),   
                         START=c(2005,2),FREQ=4),

  w_row_p = TIMESERIES(c(w_row_p),   
                         START=c(2005,2),FREQ=4),
  b2_h_r = TIMESERIES(c(b2_h_r),   
                         START=c(2005,2),FREQ=4),

  b2_f_r = TIMESERIES(c(b2_f_r),   
                    START=c(2005,2),FREQ=4),
  b2_g_r = TIMESERIES(c(b2_g_r),   
                     START=c(2005,2),FREQ=4),
  p_tax = TIMESERIES(c(p_tax),   
                         START=c(2005,2),FREQ=4),
  p_tax_row = TIMESERIES(c(p_tax_row),   
                        START=c(2005,2),FREQ=4),
  p_sub = TIMESERIES(c(p_sub),   
                         START=c(2005,2),FREQ=4),
  p_sub_row = TIMESERIES(c(p_sub_row),   
                       START=c(2005,2),FREQ=4),
  zz = TIMESERIES(c(zz),   
                        START=c(2005,2),FREQ=4),
  i_equip_g_k = TIMESERIES(c(i_equip_g_k),   
                        START=c(2005,2),FREQ=4),
  i_equip_fc_k = TIMESERIES(c(i_equip_fc_k),   
                        START=c(2005,2),FREQ=4),
  i_bd_g_k = TIMESERIES(c(i_bd_g_k),   
                       START=c(2005,2),FREQ=4),
  i_bd_fc_k = TIMESERIES(c(i_bd_fc_k),   
                         START=c(2005,2),FREQ=4),
  i_adj_h_k = TIMESERIES(c(i_adj_h_k),   
                           START=c(2005,2),FREQ=4),
  i_adj_nfc_k = TIMESERIES(c(i_adj_nfc_k),   
                            START=c(2005,2),FREQ=4),

  i_adj_fc_k = TIMESERIES(c(i_adj_fc_k),   
                     START=c(2005,2),FREQ=4),
  i_adj_g_k = TIMESERIES(c(i_adj_g_k),   
                         START=c(2005,2),FREQ=4),
  insu = TIMESERIES(c(insu),   
                     START=c(2005,2),FREQ=4),
  int_nf_adj = TIMESERIES(c(int_nf_adj),   
                         START=c(2005,3),FREQ=4),
  ins_f_adj = TIMESERIES(c(ins_f_adj),   
                  START=c(2005,3),FREQ=4),

  oth_nf = TIMESERIES(c(oth_nf),   
                         START=c(2005,2),FREQ=4),
  tax_rate_nf = TIMESERIES(c(tax_rate_nf),   
                                START=c(2005,2),FREQ=4),
  np_nf = TIMESERIES(c(np_nf ),   
                       START=c(2005,2),FREQ=4),
  ctr_nf = TIMESERIES(c(ctr_nf),   
                          START=c(2005,2),FREQ=4),

  iba_nf_rvx = TIMESERIES(c(iba_nf_rvx),   
                             START=c(2005,3),FREQ=4),

  nl_nf_adj = TIMESERIES(c(nl_nf_adj),   
                       START=c(2005,3),FREQ=4),

  l_nf_tr = TIMESERIES(c(l_nf_tr),   
                       START=c(2005,2),FREQ=4),
  sec_nf_tr = TIMESERIES(c(sec_nf_tr),   
                      START=c(2005,2),FREQ=4),

  ins_nf_tr = TIMESERIES(c(ins_nf_tr),   
                      START=c(2005,2),FREQ=4),
  sec_nf_tr = TIMESERIES(c(sec_nf_tr),   
                      START=c(2005,2),FREQ=4),
  sec_nf_rvx = TIMESERIES(c(sec_nf_rvx),   
                     START=c(2005,3),FREQ=4),
  l_nf_rvx = TIMESERIES(c(l_nf_rvx),   
                     START=c(2005,3),FREQ=4),
  eq_nf_a_tr = TIMESERIES(c(eq_nf_a_tr),   
                         START=c(2005,2),FREQ=4),
  eq_nf_a_rvx = TIMESERIES(c(eq_nf_a_rvx),   
                  START=c(2005,3),FREQ=4),
  eq_nf_a_nf_tr = TIMESERIES(c(eq_nf_a_nf_tr),   
                           START=c(2005,2),FREQ=4),
  eq_nf_a_nf_rv = TIMESERIES(c(eq_nf_a_nf_rv),   
                            START=c(2005,2),FREQ=4),
  eq_nf_a_f_tr = TIMESERIES(c(eq_nf_a_f_tr),   
                        START=c(2005,2),FREQ=4),
  eq_nf_a_f_rv = TIMESERIES(c(eq_nf_a_f_rv),   
                         START=c(2005,2),FREQ=4),
  eq_nf_a_row_tr = TIMESERIES(c(eq_nf_a_row_tr),   
                         START=c(2005,2),FREQ=4),
  eq_nf_a_row_rv = TIMESERIES(c(eq_nf_a_row_rv),   
                           START=c(2005,2),FREQ=4),

  eq_nf_l_tr = TIMESERIES(c(eq_nf_l_tr),   
                          START=c(2005,2),FREQ=4),
  eq_nf_l_rvx = TIMESERIES(c(eq_nf_l_rvx),   
                         START=c(2005,3),FREQ=4),
  alpha_neq_nf = TIMESERIES(c(alpha_neq_nf),   
                    START=c(2005,2),FREQ=4),
  ins_nf_rvx = TIMESERIES(c(ins_nf_rvx),   
                          START=c(2005,3),FREQ=4),
  delta_bd_nfc = TIMESERIES(c(delta_bd_nfc),   
                         START=c(2005,2),FREQ=4),
  delta_equip_nfc = TIMESERIES(c(delta_equip_nfc),   
                          START=c(2005,2),FREQ=4),

  iboa = TIMESERIES(c(iboa),   
                          START=c(2005,2),FREQ=4),
  div_f_adj = TIMESERIES(c(div_f_adj),   
                         START=c(2005,3),FREQ=4),
  div_r_f = TIMESERIES(c(div_r_f),   
                    START=c(2005,2),FREQ=4),
  ins_r_f = TIMESERIES(c(ins_r_f),   
                          START=c(2005,2),FREQ=4),
  res_r_f = TIMESERIES(c(res_r_f),   
                         START=c(2005,2),FREQ=4),

  int_p_f = TIMESERIES(c(int_p_f),   
                      START=c(2005,2),FREQ=4),
  div_p_f = TIMESERIES(c(div_p_f),   
                           START=c(2005,2),FREQ=4),
  ins_p_f = TIMESERIES(c(ins_p_f ),   
                     START=c(2005,2),FREQ=4),
  res_p_f = TIMESERIES(c(res_p_f),   
                      START=c(2005,2),FREQ=4),

  int_f_adj = TIMESERIES(c(int_f_adj),   
                          START=c(2005,3),FREQ=4),

  tax_f = TIMESERIES(c(tax_f),   
                         START=c(2005,2),FREQ=4),

  oth_f = TIMESERIES(c(oth_f),   
                       START=c(2005,2),FREQ=4),
  ctr_f = TIMESERIES(c(ctr_f),   
                         START=c(2005,2),FREQ=4),

  iba_f_rvx = TIMESERIES(c(iba_f_rvx),   
                         START=c(2005,3),FREQ=4),
  iba_g_tr = TIMESERIES(c(iba_g_tr),   
                         START=c(2005,2),FREQ=4),
  eq_f_a_rvx = TIMESERIES(c(eq_f_a_rvx),   
                          START=c(2005,3),FREQ=4),
  eq_f_a_tr = TIMESERIES(c(eq_f_a_tr),   
                       START=c(2005,2),FREQ=4),
  eq_f_a_nf_tr = TIMESERIES(c(eq_f_a_nf_tr),   
                        START=c(2005,2),FREQ=4),
  eq_f_a_nf_rv = TIMESERIES(c(eq_f_a_nf_rv),   
                          START=c(2005,2),FREQ=4),
  eq_f_a_f_tr = TIMESERIES(c(eq_f_a_f_tr),   
                           START=c(2005,2),FREQ=4),
  eq_f_a_f_rv = TIMESERIES(c(eq_f_a_f_rv),   
                             START=c(2005,2),FREQ=4),

  eq_f_a_row_tr = TIMESERIES(c(eq_f_a_row_tr),   
                             START=c(2005,2),FREQ=4),
  eq_f_a_row_rv = TIMESERIES(c(eq_f_a_row_rv),   
                            START=c(2005,2),FREQ=4),
  eq_f_l_tr = TIMESERIES(c(eq_f_l_tr),   
                            START=c(2005,2),FREQ=4),
  eq_f_l_rvx = TIMESERIES(c(eq_f_l_rvx),   
                              START=c(2005,3),FREQ=4),
  neq_f_rvx = TIMESERIES(c(Neq_f_rvx),   
                              START=c(2005,3),FREQ=4),

  sec_f_d_rvx = TIMESERIES(c(sec_f_d_rvx),   
                          START=c(2005,3),FREQ=4),
  nl_f_adj = TIMESERIES(c(nl_f_adj),   
                           START=c(2005,2),FREQ=4),
  sec_f_a_rvx = TIMESERIES(c(sec_f_a_rvx),   
                            START=c(2005,3),FREQ=4),
  sec_row_tr = TIMESERIES(c(sec_row_tr),   
                          START=c(2005,2),FREQ=4),
  l_f_rvx = TIMESERIES(c(l_f_rvx),   
                            START=c(2005,3),FREQ=4),
  ins_f_rvx = TIMESERIES(c(ins_f_rvx),   
                               START=c(2005,3),FREQ=4),
  delta_bd_fc = TIMESERIES(c(delta_bd_fc),   
                       START=c(2005,2),FREQ=4),
  delta_equip_fc = TIMESERIES(c(delta_equip_fc),   
                         START=c(2005,2),FREQ=4),

  div_g_adj = TIMESERIES(c(div_g_adj),   
                         START=c(2005,3),FREQ=4),

  int_r_g = TIMESERIES(c(int_r_g),   
                     START=c(2005,2),FREQ=4),

  div_r_g = TIMESERIES(c(div_r_g),   
                     START=c(2005,2),FREQ=4),
  res_r_g = TIMESERIES(c(res_r_g),   
                     START=c(2005,2),FREQ=4),

  int_p_g = TIMESERIES(c(int_p_g),   
                         START=c(2005,2),FREQ=4),
  res_p_g = TIMESERIES(c(res_p_g),   
                        START=c(2005,2),FREQ=4),
  int_g_adj = TIMESERIES(c(int_g_adj),   
                          START=c(2005,3),FREQ=4),
  ins_g_adj = TIMESERIES(c(ins_g_adj),   
                         START=c(2005,3),FREQ=4),
  oth_g = TIMESERIES(c(oth_g),   
                            START=c(2005,2),FREQ=4),
  tax_row = TIMESERIES(c(tax_row),   
                            START=c(2005,2),FREQ=4),
  ctr_g = TIMESERIES(c(ctr_g),   
                           START=c(2005,2),FREQ=4),
  iba_g_rvx = TIMESERIES(c(iba_g_rvx),   
                           START=c(2005,3),FREQ=4),

  eq_g_tr = TIMESERIES(c(eq_g_tr),   
                             START=c(2005,2),FREQ=4),
  eq_g_rvx = TIMESERIES(c(eq_g_rvx),   
                             START=c(2005,3),FREQ=4),
  eq_g_nf_tr = TIMESERIES(c(eq_g_nf_tr),   
                         START=c(2005,2),FREQ=4),
  eq_g_f_tr = TIMESERIES(c(eq_g_f_tr),   
                          START=c(2005,2),FREQ=4),
  eq_g_f_rv = TIMESERIES(c(eq_g_f_rv),   
                         START=c(2005,2),FREQ=4),

  eq_g_row_tr = TIMESERIES(c(eq_g_row_tr),   
                           START=c(2005,2),FREQ=4),
  eq_g_row_rv = TIMESERIES(c(eq_g_row_rv),   
                        START=c(2005,2),FREQ=4),
  sec_g_rvx = TIMESERIES(c(sec_g_rvx),   
                           START=c(2005,3),FREQ=4),
  nl_g_adj = TIMESERIES(c(nl_g_adj),   
                          START=c(2005,2),FREQ=4),
  l_g_tr = TIMESERIES(c(l_g_tr),   
                       START=c(2005,2),FREQ=4),
  eq_g_tr = TIMESERIES(c(eq_g_tr),   
                         START=c(2005,2),FREQ=4),
  ins_g_tr = TIMESERIES(c(ins_g_tr),   
                           START=c(2005,2),FREQ=4),

  l_g_rvx = TIMESERIES(c(l_g_rvx),   
                           START=c(2005,3),FREQ=4),
  ins_g_rvx = TIMESERIES(c(ins_g_rvx),   
                        START=c(2005,3),FREQ=4),
  delta_bd_g = TIMESERIES(c(delta_bd_g),   
                      START=c(2005,2),FREQ=4),
  i_equip_g_k = TIMESERIES(c(i_equip_g_k),   
                       START=c(2005,2),FREQ=4),
  delta_equip_g = TIMESERIES(c(delta_equip_g),   
                        START=c(2005,2),FREQ=4),
  delta_bd_g = TIMESERIES(c(delta_bd_g),   
                      START=c(2005,2),FREQ=4),

  div_row_adj = TIMESERIES(c(div_row_adj),   
                         START=c(2005,3),FREQ=4),

  int_r_row = TIMESERIES(c(int_r_row),   
                       START=c(2005,2),FREQ=4),

  div_r_row = TIMESERIES(c(div_r_row),   
                       START=c(2005,2),FREQ=4),
  ins_r_row = TIMESERIES(c(ins_r_row),   
                       START=c(2005,2),FREQ=4),

  res_r_row = TIMESERIES(c(res_r_row),   
                       START=c(2005,2),FREQ=4),
  int_p_row = TIMESERIES(c(int_p_row),   
                       START=c(2005,2),FREQ=4),
  div_p_row = TIMESERIES(c(div_p_row),   
                         START=c(2005,2),FREQ=4),
  ins_p_row = TIMESERIES(c(ins_p_row),   
                         START=c(2005,2),FREQ=4),
  res_p_row = TIMESERIES(c(res_p_row),   
                     START=c(2005,2),FREQ=4),
  int_row_adj = TIMESERIES(c(int_row_adj),   
                       START=c(2005,3),FREQ=4),
  ins_row_adj = TIMESERIES(c(ins_row_adj),   
                     START=c(2005,3),FREQ=4),
  oth_row = TIMESERIES(c(oth_row),   
                         START=c(2005,2),FREQ=4),

  ctr_row = TIMESERIES(c(ctr_row),   
                       START=c(2005,2),FREQ=4),
  np_row = TIMESERIES(c(np_row),   
                        START=c(2005,2),FREQ=4),
  scon_row_r = TIMESERIES(c(scon_row_r),   
                          START=c(2005,2),FREQ=4),
  scon_row_p = TIMESERIES(c(scon_row_p),   
                         START=c(2005,2),FREQ=4),
  sben_row_r = TIMESERIES(c(sben_row_r),   
                         START=c(2005,2),FREQ=4),

  sben_row_p = TIMESERIES(c(sben_row_p),   
                           START=c(2005,2),FREQ=4),
  iba_row_rvx = TIMESERIES(c(iba_row_rvx),   
                         START=c(2005,3),FREQ=4),
  nl_row_adj = TIMESERIES(c(nl_row_adj),   
                        START=c(2005,3),FREQ=4),
  l_row_tr = TIMESERIES(c(l_row_tr),   
                      START=c(2005,2),FREQ=4),
  ins_row_tr = TIMESERIES(c(ins_row_tr),   
                       START=c(2005,2),FREQ=4),
  sec_row_rvx = TIMESERIES(c(sec_row_rvx),   
                        START=c(2005,3),FREQ=4),

  l_row_tr = TIMESERIES(c(l_row_tr),   
                       START=c(2005,2),FREQ=4),
  l_row_rvx = TIMESERIES(c(l_row_rvx),   
                         START=c(2005,3),FREQ=4),
  eq_row_a_tr = TIMESERIES(c(eq_row_a_tr),   
                          START=c(2005,2),FREQ=4),
  eq_row_a_rvx = TIMESERIES(c(eq_row_a_rvx),   
                           START=c(2005,3),FREQ=4),
  eq_row_a_nf_rv = TIMESERIES(c(eq_row_a_nf_rv),   
                             START=c(2005,2),FREQ=4),
  eq_row_a_nf_tr = TIMESERIES(c(eq_row_a_nf_tr),   
                          START=c(2005,2),FREQ=4),

  eq_row_a_f_tr = TIMESERIES(c(eq_row_a_f_tr),   
                          START=c(2005,2),FREQ=4),
  eq_row_a_f_rv = TIMESERIES(c(eq_row_a_f_rv),   
                       START=c(2005,2),FREQ=4),
  eq_row_l_tr = TIMESERIES(c(eq_row_l_tr),   
                           START=c(2005,2),FREQ=4),
  eq_row_l_rvx = TIMESERIES(c(eq_row_l_rvx),   
                          START=c(2005,3),FREQ=4),
  neq_row_rvx = TIMESERIES(c(Neq_row_rvx),   
                        START=c(2005,3),FREQ=4),
  ins_row_tr = TIMESERIES(c(ins_row_tr),   
                          START=c(2005,2),FREQ=4),
  ins_row_rvx = TIMESERIES(c(ins_row_rvx),   
                           START=c(2005,3),FREQ=4),

  empadj = TIMESERIES(c(empadj),   
                        START=c(2005,2),FREQ=4),
  part = TIMESERIES(c(part),   
                         START=c(2005,2),FREQ=4),
  pop = TIMESERIES(c(pop),   
                           START=c(2005,2),FREQ=4),
  urs = TIMESERIES(c(urs),   
                            START=c(2005,2),FREQ=4),
  wage_2010q3 = TIMESERIES(c(wage_2010q3),   
                              START=c(2005,2),FREQ=4),
  neq_nf_rvx = TIMESERIES(c(Neq_nf_rvx),   
                           START=c(2005,3),FREQ=4),
  eq_h_f_rvx = TIMESERIES(c(eq_h_f_rvx),   
                           START=c(2005,3),FREQ=4),
  g = TIMESERIES(c(g),   
                           START=c(2005,2),FREQ=4),
  pg = TIMESERIES(c(pg),   
                           START=c(2005,2),FREQ=4),
  px = TIMESERIES(c(px),   
                           START=c(2005,2),FREQ=4),
  pf = TIMESERIES(c(pf),   
                           START=c(2005,2),FREQ=4),
  pm = TIMESERIES(c(pm),   
                           START=c(2005,2),FREQ=4),
  div_nf_adj = TIMESERIES(c(div_nf_adj),   
                           START=c(2005,3),FREQ=4),
  ins_nf_adj = TIMESERIES(c(ins_nf_adj),   
                           START=c(2005,3),FREQ=4),
  res_p_nf = TIMESERIES(c(res_p_nf),   
                           START=c(2005,2),FREQ=4),
  res_r_nf = TIMESERIES(c(res_r_nf),   
                           START=c(2005,2),FREQ=4),
  int_r_f = TIMESERIES(c(int_r_f),   
                           START=c(2005,2),FREQ=4),
  np_f = TIMESERIES(c(np_f),   
                           START=c(2005,2),FREQ=4),
  np_g = TIMESERIES(c(np_g),   
                           START=c(2005,2),FREQ=4),
  eq_g_nf_rv = TIMESERIES(c(eq_g_nf_rv),   
                           START=c(2005,2),FREQ=4),
  prodk = TIMESERIES(c(prodk),   
                           START=c(2005,2),FREQ=4),

  ### Checks

  nl_check = TIMESERIES(c(nl_check),   
                           START=c(2005,3),FREQ=4),
  fnl_check = TIMESERIES(c(fnl_check),   
                           START=c(2005,3),FREQ=4),
  check_np  = TIMESERIES(c(check_np),   
                           START=c(2005,2),FREQ=4),
  check_ctr  = TIMESERIES(c(check_ctr),   
                           START=c(2005,2),FREQ=4),
  check_invest  = TIMESERIES(c(check_invest ),   
                           START=c(2005,2),FREQ=4),
  check_iba  = TIMESERIES(c(check_iba ),   
                           START=c(2005,2),FREQ=4),
  check_tax  = TIMESERIES(c(check_tax ),   
                           START=c(2005,2),FREQ=4),
  check_iba_rv  = TIMESERIES(c(check_iba_rv ),   
                           START=c(2005,2),FREQ=4),
  check_iba_tr  = TIMESERIES(c(check_iba_tr ),   
                           START=c(2005,2),FREQ=4),
  check_eq  = TIMESERIES(c(check_eq ),   
                           START=c(2005,2),FREQ=4),
  check_eq_tr  = TIMESERIES(c(check_eq_tr ),   
                           START=c(2005,3),FREQ=4),
  check_div  = TIMESERIES(c(check_div ),   
                           START=c(2005,3),FREQ=4),
  check_l  = TIMESERIES(c(check_l ),   
                           START=c(2005,2),FREQ=4),
  check_l_tr  = TIMESERIES(c(check_l_tr ),   
                           START=c(2005,2),FREQ=4),
  check_ins  = TIMESERIES(c(check_ins ),   
                           START=c(2005,2),FREQ=4),
  check_ins_tr  = TIMESERIES(c(check_ins_tr ),   
                           START=c(2005,2),FREQ=4),
  check_sec  = TIMESERIES(c(check_sec ),   
                           START=c(2005,2),FREQ=4),

  fnl_f  = TIMESERIES(c(fnl_f ),   
                           START=c(2005,2),FREQ=4),
  fnl_nf  = TIMESERIES(c(fnl_nf ),   
                           START=c(2005,3),FREQ=4),
  fnl_g   = TIMESERIES(c(fnl_g ),   
                           START=c(2005,2),FREQ=4),
  fnl_row   = TIMESERIES(c(fnl_row),   
                           START=c(2005,3),FREQ=4),
  iba_h_rv   = TIMESERIES(c(iba_h_rv),   
                           START=c(2005,2),FREQ=4),
  iba_f_rv   = TIMESERIES(c(iba_f_rv  ),   
                           START=c(2005,2),FREQ=4),
  iba_nf_rv   = TIMESERIES(c(iba_nf_rv  ),   
                           START=c(2005,2),FREQ=4),
  iba_g_rv   = TIMESERIES(c(iba_g_rv  ),   
                           START=c(2005,2),FREQ=4),
  iba_row_rv  = TIMESERIES(c(iba_row_rv ),   
                           START=c(2005,2),FREQ=4),
  fnl_h  = TIMESERIES(c(fnl_h ),   
                           START=c(2005,3),FREQ=4),
  eq_h_ratio  = TIMESERIES(c(eq_h_ratio ),   
                           START=c(2005,3),FREQ=4),
  ratio_in_eq_h  = TIMESERIES(c(ratio_in_eq_h ),   
                           START=c(2005,3),FREQ=4),
  d_2007q3  = TIMESERIES(c(d_2007q3 ),   
                           START=c(2005,3),FREQ=4),
  d_2017q23  = TIMESERIES(c(d_2017q23),   
                           START=c(2005,3),FREQ=4),
  d_2009q2  = TIMESERIES(c(d_2009q2),   
                           START=c(2005,3),FREQ=4),
  ## Test

   shock_reg_model  = TIMESERIES(c(shock_reg_model ),   
                           START=c(2005,2),FREQ=4),

   shock_g_model  = TIMESERIES(c(shock_g_model ),   
                           START=c(2005,2),FREQ=4),
  shock_max_dp_model  = TIMESERIES(c(shock_max_dp_model ),   
                           START=c(2005,2),FREQ=4),
  shock_part_model  = TIMESERIES(c(shock_part_model ),   
                           START=c(2005,2),FREQ=4),

  shock_1_model  = TIMESERIES(c(shock_1_model ),   
                           START=c(2005,2),FREQ=4),
  shock_2_model  = TIMESERIES(c(shock_2_model ),   
                           START=c(2005,2),FREQ=4),
  shock_3_model  = TIMESERIES(c(shock_3_model ),   
                           START=c(2005,2),FREQ=4),
   pf_shock = TIMESERIES(c(pf_shock),   
                        START=c(2005,2),FREQ=4),
  g_shock = TIMESERIES(c(g_shock),   
                        START=c(2005,2),FREQ=4),
  pm_ds_shock = TIMESERIES(c(pm_ds_shock),   
                        START=c(2005,2),FREQ=4),
  iloan_shock = TIMESERIES(c(iloan_shock),   
                        START=c(2005,2),FREQ=4),
  idep_shock = TIMESERIES(c(idep_shock),   
                        START=c(2005,2),FREQ=4),
   ibd_shock = TIMESERIES(c(ibd_shock),   
                        START=c(2005,2),FREQ=4),
  iboa_shock = TIMESERIES(c(iboa_shock),   
                        START=c(2005,2),FREQ=4),
   tax_rate1_shock = TIMESERIES(c(tax_rate1_shock),   
                        START=c(2005,2),FREQ=4)
)
```

```{r}
S_model=LOAD_MODEL_DATA(S_model,S_modelData)

```

# Estimate

```{r}

#Estimate model coefficients
S_model=ESTIMATE(S_model
                 ,TSRANGE=c(2006,3,2020,1)
)
```

# Simulate

```{r}
# In-sample prediction (no add factors)
S_model <- SIMULATE(S_model
                    ,simType='DYNAMIC'
                    ,TSRANGE=c(2006,3,2020,1)
                    ,simConvergence=0.00001
                    ,simIterLimit=100
                   )
```

```{r, echo=FALSE}
wage_trim_00 = S_model$simulation$wage_trim
dp_person_00 = S_model$simulation$dp_person
max_dp_00 = S_model$simulation$max_dp
stat_tilpas_00 = S_model$simulation$stat_tilpas
stat_reg_00= S_model$simulation$stat_tilpas
tilpas_pr_00 = S_model$simulation$tilpas_pr

yk_ds_00 = S_model$simulation$yk_ds
part_00 = S_model$simulation$part
lf_00 = S_model$simulation$lf
pconk_ds_00 = S_model$simulation$pconk_ds
yd_h1_00 = S_model$simulation$yd_h1 
nben_h_rest_00 = S_model$simulation$nben_h_rest
nben_h_un_00 = S_model$simulation$nben_h_un
komp_r_00 = S_model$simulation$komp_r
nben_h_00 = S_model$simulation$nben_h
wageindex_00 = S_model$simulation$wageindex
yh1_00 = S_model$simulation$yh1
yd1_h_00 = S_model$simulation$yd1_h
yd_hk_00 = S_model$simulation$yd_hk
yd_h_00 = S_model$simulation$yd_h
l_h_tr_00 = S_model$simulation$l_h_tr
s_h_00 = S_model$simulation$s_h
eq_h_00 = S_model$simulation$eq_h
iba_h_00 = S_model$simulation$iba_h
yd1_hk_ds_00 = S_model$simulation$yd1_hk_ds
ur_00 = S_model$simulation$ur
ur_ds_00 = S_model$simulation$ur_ds
iba_f_tr_00 = S_model$simulation$iba_f_tr
iba_h_tr_00 = S_model$simulation$iba_h_tr
iba_nf_tr_00 = S_model$simulation$iba_nf_tr
nl_nf_00 = S_model$simulation$nl_nf
neq_nf_tr_00 = S_model$simulation$neq_nf_tr
l_nf_tr_00 = S_model$simulation$l_nf_tr
sec_nf_tr_00 = S_model$simulation$sec_nf_tr
ins_nf_tr_00 = S_model$simulation$ins_nf_tr
neq_nf_00 = S_model$simulation$neq_nf
i_nf_00 = S_model$simulation$i_nf
np_nf_00 = S_model$simulation$np_nf
s_nf_00 = S_model$simulation$s_nf
yh_nf_00 = S_model$simulation$yh_nf
y_00 = S_model$simulation$y
b2_nf_r_00 = S_model$simulation$b2_nf_r
p_tax_00 = S_model$simulation$p_tax
b2_00 = S_model$simulation$b2
w_nf_00 = S_model$simulation$w_nf
npropinc_nf_00 = S_model$simulation$npropinc_nf
int_nf_00 = S_model$simulation$int_nf
div_nf_00 = S_model$simulation$div_nf
eq_nf_a_f_00 = S_model$simulation$eq_nf_a_f
eq_nf_l_nf_00 = S_model$simulation$eq_nf_l_nf
eq_nf_a_nf_00 = S_model$simulation$eq_nf_a_nf
eq_h_nf_00 = S_model$simulation$eq_h_nf
eq_row_a_nf_00 = S_model$simulation$eq_row_a_nf
eq_f_a_nf_00 = S_model$simulation$eq_f_a_nf
eq_g_nf_00 = S_model$simulation$eq_g_nf
eq_nf_a_row_00 = S_model$simulation$eq_nf_a_row
neq_nf_00 = S_model$simulation$neq_nf
check_div_00 = S_model$simulation$check_div
xk_ds_00 = S_model$simulation$xk_ds
g_00 = S_model$simulation$g
rw_ds_00 = S_model$simulation$rw_ds
b2_00 = S_model$simulation$b2
test = S_model$simulation$test

### Gengivelse
yk_ds_00 = S_model$simulation$yk_ds
emp_00 = S_model$simulation$emp
wage_ds_00 = S_model$simulation$wage_ds
unemp_00 = S_model$simulation$unemp
ik_00 = S_model$simulation$ik
pc_ds_00 = S_model$simulation$pc_ds
yd_hk_ds_00 = S_model$simulation$yd_hk_ds ## kunne godt se ud som om noget er galt med den her.

### Test

yd1_hk_ds_00 = S_model$simulation$yd1_hk_ds
yd2a_hk_ds_00 = S_model$simulation$yd2a_hk_ds
yd2b_hk_00 = S_model$simulation$yd2b_hk
yd1_hk_00 = S_model$simulation$yd1_hk
yd2a_hk_00 = S_model$simulation$yd2a_hk
yd1_h_00 = S_model$simulation$yd1_h
yh1_00 = S_model$simulation$yh1
sec_g_tr_00 = S_model$simulation$sec_g_tr
sec_f_d_tr_g_00 = S_model$simulation$sec_f_d_tr_g
sec_h_tr_g_00 = S_model$simulation$sec_h_tr_g
l_h_00 = S_model$simulation$l_h
yd2b_h_00 = S_model$simulation$yd2b_h
yh2b_00 = S_model$simulation$yh2b
propinc_p_h_00 = S_model$simulation$propinc_p_h
int_p_h_00 = S_model$simulation$int_p_h
iloan_00 = S_model$simulation$iloan
int_p_h_adj_00 = S_model$simulation$int_p_h_adj

## RESCHECK
yk_00 = S_model$simulation$yk
prodk_00 = S_model$simulation$prodk
pcon_00 = S_model$simulation$pcon
i_00 = S_model$simulation$i
x_00 = S_model$simulation$x
m_00 = S_model$simulation$m
i_h_00 = S_model$simulation$i_h
w_nf_00 = S_model$simulation$w_nf
s_nf_00 = S_model$simulation$s_nf
pconk_00 = S_model$simulation$pconk
ik_00 = S_model$simulation$ik
gk_00 = S_model$simulation$gk
xk_00 = S_model$simulation$xk
mk_00 = S_model$simulation$mk
g_00 = S_model$simulation$g
ur_00 = S_model$simulation$ur
ur_ds_00 = S_model$simulation$ur_ds
prod_ds_00 = S_model$simulation$prod_ds
int_r_h_00 = S_model$simulation$int_r_h
int_p_h_00 = S_model$simulation$int_p_h
div_h_00 = S_model$simulation$div_h
aa_check_00 = S_model$simulation$aa_check
insu_h_00 = S_model$simulation$insu_h
propinc_r_h_00 = S_model$simulation$propinc_r_h
propinc_p_h_00 = S_model$simulation$propinc_p_h
yh_h_00 = S_model$simulation$yh_h
yh2a_00 = S_model$simulation$yh2a
yh1_00 = S_model$simulation$yh1
w_h_r_00 = S_model$simulation$w_h_r
yd2a_h_00 = S_model$simulation$yd2a_h
yd2b_h_00 = S_model$simulation$yd2b_h 
yd1_h_00 = S_model$simulation$yd1_h
tax_h_00 = S_model$simulation$tax_h
yd1_hk_00 = S_model$simulation$yd1_hk
yd2a_hk_00 = S_model$simulation$yd2a_hk
yd2b_hk_00 = S_model$simulation$yd2b_hk
yd_h_00 = S_model$simulation$yd_h
nsben_h_00 = S_model$simulation$nsben_h
yd_h1_00 = S_model$simulation$yd_h1
nl_h_00 = S_model$simulation$nl_h
pconk_ds_00 = S_model$simulation$pconk_ds
pconk_00 = S_model$simulation$pconk
pcon_00 = S_model$simulation$pcon
i_bd_h_k_00 = S_model$simulation$i_bd_h_k
i_equip_h_k_00 = S_model$simulation$i_equip_h_k
iba_h_00 = S_model$simulation$iba_h
sec_h_00 = S_model$simulation$sec_h
sec_h_tr_g_00 = S_model$simulation$sec_h_tr_g
l_h_00 = S_model$simulation$l_h
eq_h_tr_00 = S_model$simulation$eq_h_tr
iba_h_tr_00 = S_model$simulation$iba_h_tr
eq_h_nf_00 = S_model$simulation$eq_h_nf
eq_h_f_00 = S_model$simulation$eq_h_f
eq_h_row_00 = S_model$simulation$eq_h_row
eq_h_nf_tr_00 = S_model$simulation$eq_h_nf_tr
eq_h_f_tr_00 = S_model$simulation$eq_h_f_tr
eq_h_row_tr_00 = S_model$simulation$eq_h_row_tr
eq_h_test_00 = S_model$simulation$eq_h_test
ins_h_00 = S_model$simulation$ins_h
ins_h_tr_00 = S_model$simulation$ins_h_tr
lev_h_00 = S_model$simulation$lev_h
fa_h_00 = S_model$simulation$fa_h
fnw_h_00 = S_model$simulation$fnw_h
fnw_hk_00 = S_model$simulation$fnw_hk
i_bd_h_00 = S_model$simulation$i_bd_h
i_equip_h_00 = S_model$simulation$i_equip_h
bd_h_k_00 = S_model$simulation$bd_h_k
equip_h_k_00 = S_model$simulation$equip_h_k 
bd_h_00 = S_model$simulation$bd_h
equip_h_00 = S_model$simulation$equip_h
rer_00 = S_model$simulation$rer 
py_00 = S_model$simulation$py
s_00 = S_model$simulation$s
w_nf_00 = S_model$simulation$w_nf
yf_00 = S_model$simulation$yf
ws_00 = S_model$simulation$ws
ps_00 = S_model$simulation$ps 
u_00 = S_model$simulation$u
prate_00 = S_model$simulation$prate 
prate_ds_00 = S_model$simulation$prate_ds 
ps_ds_00 = S_model$simulation$ps_ds
tobinq_00 = S_model$simulation$tobinq
i_bd_nfc_k_00 = S_model$simulation$i_bd_nfc_k
i_equip_nfc_k_00 = S_model$simulation$i_equip_nfc_k
i_h_00 = S_model$simulation$i_h 
i_nf_00 = S_model$simulation$i_nf
i_f_00 = S_model$simulation$i_f 
i_g_00 = S_model$simulation$i_g 
int_nf_00 = S_model$simulation$int_nf
div_nf_00 = S_model$simulation$div_nf
insu_nf_00 = S_model$simulation$insu_nf
npropinc_nf_00 = S_model$simulation$npropinc_nf
yh_nf_00 = S_model$simulation$yh_nf
tax_nf_00 = S_model$simulation$tax_nf
nl_nf_00 = S_model$simulation$nl_nf 
iba_nf_00 = S_model$simulation$iba_nf 
iba_nf_tr_00 = S_model$simulation$iba_nf_tr
sec_nf_00 = S_model$simulation$sec_nf
l_nf_00 = S_model$simulation$l_nf
eq_nf_a_00 = S_model$simulation$eq_nf_a
eq_nf_a_nf_00 = S_model$simulation$eq_nf_a_nf 
eq_nf_l_nf_00 = S_model$simulation$eq_nf_l_nf
eq_nf_ax_00 = S_model$simulation$eq_nf_ax 
eq_nf_a_f_00 = S_model$simulation$eq_nf_a_f 
eq_nf_a_row_00 = S_model$simulation$eq_nf_a_row
eq_nf_l_00 = S_model$simulation$eq_nf_l
eq_nf_lx_00 = S_model$simulation$eq_nf_lx
neq_nf_00 = S_model$simulation$neq_nf
neq_nf_tr_00 = S_model$simulation$neq_nf_tr 
eq_nf_a_test_00 = S_model$simulation$eq_nf_a_test
eq_nf_l_test_00 = S_model$simulation$eq_nf_l_test 
ins_nf_00 = S_model$simulation$ins_nf 
fnw_nf_00 = S_model$simulation$fnw_nf
fnw_nfk_00 = S_model$simulation$fnw_nfk 
i_bd_nfc_00 = S_model$simulation$i_bd_nfc
i_equip_nfc_00 = S_model$simulation$i_equip_nfc 
bd_nfc_k_00 = S_model$simulation$bd_nfc_k 
equip_nfc_k_00 = S_model$simulation$equip_nfc_k 
bd_nfc_00 = S_model$simulation$bd_nfc
equip_nfc_00 = S_model$simulation$equip_nfc 
int_f_00 = S_model$simulation$int_f
div_f_00 = S_model$simulation$div_f
insu_f_00 = S_model$simulation$insu_f
propinc_r_f_00 = S_model$simulation$propinc_r_f
propinc_p_f_00 = S_model$simulation$propinc_p_f
npropinc_f_00 = S_model$simulation$npropinc_f
yh_f_00 = S_model$simulation$yh_f
yd_f_00 = S_model$simulation$yd_f
d8_f_00 = S_model$simulation$d8_f
nben_f_00 = S_model$simulation$nben_f 
s_f_00 = S_model$simulation$s_f
nl_f_00 = S_model$simulation$nl_f 
iba_f_00 = S_model$simulation$iba_f 
eq_f_a_00 = S_model$simulation$eq_f_a 
eq_f_a_nf_00 = S_model$simulation$eq_f_a_nf 
eq_f_a_f_00 = S_model$simulation$eq_f_a_f 
eq_f_l_f_00 = S_model$simulation$eq_f_l_f
eq_f_a_row_00 = S_model$simulation$eq_f_a_row 
eq_f_ax_00 = S_model$simulation$eq_f_ax
eq_f_l_00 = S_model$simulation$eq_f_l
eq_f_lx_00 = S_model$simulation$eq_f_lx 
neq_f_00 = S_model$simulation$neq_f 
neq_f_tr_00 = S_model$simulation$neq_f_tr 
eq_f_l_test_00 = S_model$simulation$eq_f_l_test 
sec_f_d_00 = S_model$simulation$sec_f_d 
sec_f_d_tr_00 = S_model$simulation$sec_f_d_tr 
sec_f_d_tr_nf_00 = S_model$simulation$sec_f_d_tr_nf
sec_f_d_tr_g_00 = S_model$simulation$sec_f_d_tr_g 
sec_f_a_00 = S_model$simulation$sec_f_a 
sec_f_a_tr_00 = S_model$simulation$sec_f_a_tr 
l_f_00 = S_model$simulation$l_f 
l_f_tr_00 = S_model$simulation$l_f_tr 
ins_f_00 = S_model$simulation$ins_f 
ins_f_tr_00 = S_model$simulation$ins_f_tr 
fnw_f_00 = S_model$simulation$fnw_f
fnw_fk_00 = S_model$simulation$fnw_fk 
i_bd_fc_00 = S_model$simulation$i_bd_fc 
i_equip_fc_00 = S_model$simulation$i_equip_fc 
bd_fc_k_00 = S_model$simulation$bd_fc_k 
equip_fc_k_00 = S_model$simulation$equip_fc_k 
bd_fc_00 = S_model$simulation$bd_fc
equip_fc_00 = S_model$simulation$equip_fc
int_g_00 = S_model$simulation$int_g 
div_g_00 = S_model$simulation$div_g 
insu_g_00 = S_model$simulation$insu_g 
propinc_r_g_00 = S_model$simulation$propinc_r_g 
propinc_p_g_00 = S_model$simulation$propinc_p_g 
npropinc_g_00 = S_model$simulation$npropinc_g 
yh_g_00 = S_model$simulation$yh_g 
yd_g_00 = S_model$simulation$yd_g 
nben_g_00 = S_model$simulation$nben_g 
tax_g_00 = S_model$simulation$tax_g 
s_g_00 = S_model$simulation$s_g 
nl_g_00 = S_model$simulation$nl_g 
iba_g_00 = S_model$simulation$iba_g 
eq_g_00 = S_model$simulation$eq_g 
eq_g_nf_00 = S_model$simulation$eq_g_nf
eq_g_f_00 = S_model$simulation$eq_g_f 
eq_g_row_00 = S_model$simulation$eq_g_row 
eq_g_ax_00 = S_model$simulation$eq_g_ax 
eq_g_test_00 = S_model$simulation$eq_g_test 
sec_g_00 = S_model$simulation$sec_g 
sec_g_tr_00 = S_model$simulation$sec_g_tr 
l_g_00 = S_model$simulation$l_g 
ins_g_00 = S_model$simulation$ins_g 
fnw_g_00 = S_model$simulation$fnw_g 
fnw_gk_00 = S_model$simulation$fnw_gk 
i_bd_g_00 = S_model$simulation$i_bd_g 
i_equip_g_00 = S_model$simulation$i_equip_g 
bd_g_k_00 = S_model$simulation$bd_g_k 
equip_g_k_00 = S_model$simulation$equip_g_k 
bd_g_00 = S_model$simulation$bd_g 
equip_g_00 = S_model$simulation$equip_g 
int_row_00 = S_model$simulation$int_row 
div_row_00 = S_model$simulation$div_row 
insu_row_00 = S_model$simulation$insu_row 
propinc_r_row_00 = S_model$simulation$propinc_r_row 
propinc_p_row_00 = S_model$simulation$propinc_p_row 
npropinc_row_00 = S_model$simulation$npropinc_row
s_row_00 = S_model$simulation$s_row
nl_row_00 = S_model$simulation$nl_row
nben_row_00 = S_model$simulation$nben_row 
nx_00 = S_model$simulation$nx 
cab_00 = S_model$simulation$cab 
bop_00 = S_model$simulation$bop 
fab_00 = S_model$simulation$fab 
iba_row_00 = S_model$simulation$iba_row 
iba_row_tr_00 = S_model$simulation$iba_row_tr 
sec_row_00 = S_model$simulation$sec_row 
l_row_00 = S_model$simulation$l_row 
eq_row_a_00 = S_model$simulation$eq_row_a 
eq_row_a_nf_00 = S_model$simulation$eq_row_a_nf 
eq_row_a_f_00 = S_model$simulation$eq_row_a_f
eq_row_ax_00 = S_model$simulation$eq_row_ax 
eq_row_a_test_00 = S_model$simulation$eq_row_a_test 
eq_row_l_00 = S_model$simulation$eq_row_l
eq_row_lx_00 = S_model$simulation$eq_row_lx 
eq_row_l_test_00 = S_model$simulation$eq_row_l_test 
neq_row_00 = S_model$simulation$neq_row
neq_row_tr_00 = S_model$simulation$neq_row_tr 
ins_row_00 = S_model$simulation$ins_row 
fnw_row_00 = S_model$simulation$fnw_row
fnw_rowk_00 = S_model$simulation$fnw_rowk
unadj_00 = S_model$simulation$unadj
unemp_00 = S_model$simulation$unemp 
ur_00 = S_model$simulation$ur 
ur_ds_00 = S_model$simulation$ur_ds
uradj_00 = S_model$simulation$uradj 
lf_00 = S_model$simulation$lf 
pc_00 = S_model$simulation$pc 
mkp_00 = S_model$simulation$mkp
emp_00 = S_model$simulation$emp
urterm_00 = S_model$simulation$urterm 
wage_00 = S_model$simulation$wage 
wageindex_00 = S_model$simulation$wageindex 
wage_ds_00 = S_model$simulation$wage_ds 
prod_00 = S_model$simulation$prod
prod_ds_00 = S_model$simulation$prod_ds 
acc_rate_00 = S_model$simulation$acc_rate 
nl_check_00 = S_model$simulation$nl_check
fnl_check_00 = S_model$simulation$fnl_check
check_np_00 = S_model$simulation$check_np
check_ctr_00 = S_model$simulation$check_ctr 
check_invest_00 = S_model$simulation$check_invest 
check_iba_00 = S_model$simulation$check_iba
check_tax_00 = S_model$simulation$check_tax_00
check_iba_rv_00 = S_model$simulation$check_iba_rv 
check_iba_tr_00 = S_model$simulation$check_iba_tr
check_eq_00 = S_model$simulation$check_eq
check_eq_tr_00 = S_model$simulation$check_eq_tr
check_div_00 = S_model$simulation$check_div
check_l_00 = S_model$simulation$check_l 
check_l_tr_00 = S_model$simulation$check_l_tr
check_ins_00 = S_model$simulation$check_ins
check_ins_tr_00 = S_model$simulation$check_ins_tr
check_sec_00 = S_model$simulation$check_sec
npropinc_h_00 = S_model$simulation$npropinc_h

sens1_00 = S_model$simulation$sens1

part1_00 = S_model$simulation$part1
```

# Tjek

```{r}
plot(tilpas_pr_00, col= "green"); lines(tilpas_pr, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(wage_trim_00, col= "green"); lines(wage_trim1, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(dp_person_00, col= "green"); lines(mean_dp_regul, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(stat_tilpas_00, col= "green", ylim= c(-0.03,0.05)); lines(stat_tilpas, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(max_dp_00, col= "green"); lines(obs_dp_q, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(wage_00, col= "green"); lines(wage, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

```

```{r}

### Gengivelse
plot(yk_ds_00, ylim= c(430000, 540000), col= "green"); lines(yk_ds, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(emp_00, col= "green"); lines(emp, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(pconk_ds_00,ylim= c(210000, 250000), col= "green"); lines(pconk_ds, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(ik_00, col= "green"); lines(ik, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(yd_hk_ds_00, col= "green"); lines(yd_hk_ds, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

plot(part_00, col= "green"); lines(part, col = "blue"); 
legend("top",legend=c("Baseline", "Real data"),lty=c(1,1),lwd=c(2),bty="n", col = c("green", "blue"));
grid()

```

Plots for paper

```{r, fig.width=12, fig.height=7}
par(mfrow=c(1,2))

plot(yk_ds_00, col= "red"); lines(yk_ds, col = "blue"); 
legend("top",legend=c("Baseline", "Data"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(emp_00, col= "red"); lines(emp, col = "blue"); 
legend("top",legend=c("Baseline", "Data"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

```

```{r, fig.width=12, fig.height=7}
par(mfrow=c(1,2))

plot(max_dp_00, col= "red"); lines(obs_dp_q, col = "blue"); 
legend("top",legend=c("Baseline", "Data"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(komp_r_00, col= "red"); lines(komp_r_obs, col = "blue"); 
legend("top",legend=c("Baseline", "Data"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

```

# Shocks

## Scenario 1 max_dp

```{r}
shock_max_dp_model_shock=shock_max_dp_model %>% window(start= c(2006,3))
```

```{r}
# Define add-factor list (defining shock)
constantAdjList <- list(

 shock_max_dp_model = TIMESERIES(shock_max_dp_model_shock+1 ,START=c(2006,3),  FREQ='Q')

)

```

```{r}
# Simulate the model (following shock)
Shock_1 <- SIMULATE(S_model
                    ,simType='DYNAMIC'
                    ,TSRANGE=c(2006,3,2020,1)
                    ,simConvergence=0.00001
                    ,simIterLimit=100
                    ,ConstantAdjustment=constantAdjList
                    )
```

```{r}
###Gengivelse af stød til model
max_dp_01 = Shock_1$simulation$max_dp
wage_01 = Shock_1$simulation$wage
komp_r_01 = Shock_1$simulation$komp_r
part_01 = Shock_1$simulation$part
yk_ds_01 = Shock_1$simulation$yk_ds
pc_ds_01 = Shock_1$simulation$pc_ds
emp_01 = Shock_1$simulation$emp
ur_ds_01 = Shock_1$simulation$ur_ds

yd_hk_ds_01 = Shock_1$simulation$yd_hk_ds

tilpas_pr_01 = Shock_1$simulation$tilpas_pr

part_01 = Shock_1$simulation$part
```

```{r}

plot(yk_ds_01, col= "red"); lines(yk_ds_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(max_dp_01, col= "red"); lines(max_dp_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(wage_01, col= "red"); lines(wage_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(komp_r_01, col= "red"); lines(komp_r_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(part_01, col= "red"); lines(part_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(pc_ds_01, col= "red"); lines(pc_ds_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()
```

Plots for paper:

```{r}
par(mfrow=c(1,2))

plot(yk_ds_01, col= "red"); lines(yk_ds_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(emp_01, col= "red"); lines(emp_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()
```

## Scenario 1 reg

```{r}
shock_reg_model_shock=shock_reg_model %>% window(start= c(2016,1))
```

```{r}
# Define add-factor list (defining shock)
constantAdjList1 <- list(

 shock_reg_model = TIMESERIES(shock_reg_model_shock+1 ,START=c(2016,1),  FREQ='Q')

)

```

```{r}
# Simulate the model (following shock)
Shock_1b <- SIMULATE(S_model
                    ,simType='DYNAMIC'
                    ,TSRANGE=c(2006,3,2020,1)
                    ,simConvergence=0.00001
                    ,simIterLimit=100
                    ,ConstantAdjustment=constantAdjList1
                    )
```

```{r}
###Gengivelse af stød til model
max_dp_01b = Shock_1b$simulation$max_dp
wage_01b = Shock_1b$simulation$wage
komp_r_01b = Shock_1b$simulation$komp_r
part_01b = Shock_1b$simulation$part
yk_ds_01b = Shock_1b$simulation$yk_ds
pc_ds_01b = Shock_1b$simulation$pc_ds
emp_01b = Shock_1b$simulation$emp
ur_ds_01b = Shock_1b$simulation$ur_ds

yd_hk_ds_01b = Shock_1b$simulation$yd_hk_ds

tilpas_pr_01b = Shock_1b$simulation$tilpas_pr

part_01b = Shock_1b$simulation$part
```

## Scenario 2 part

```{r}
shock_part_model_shock=shock_part_model %>% window(start= c(2006,3))
```

```{r}
# Define add-factor list (defining shock)
constantAdjList <- list(

 shock_part_model = TIMESERIES(shock_part_model_shock+1 ,START=c(2006,3),  FREQ='Q'),
  shock_max_dp_model = TIMESERIES(shock_max_dp_model_shock+1 ,START=c(2006,3),  FREQ='Q')

)

```

```{r}
# Simulate the model (following shock)
Shock_2 <- SIMULATE(S_model
                    ,simType='DYNAMIC'
                    ,TSRANGE=c(2006,3,2020,1)
                    ,simConvergence=0.00001
                    ,simIterLimit=100
                    ,ConstantAdjustment=constantAdjList
                    )
```

```{r}
###Gengivelse af stød til model
max_dp_02 = Shock_2$simulation$max_dp
wage_02 = Shock_2$simulation$wage
komp_r_02 = Shock_2$simulation$komp_r
part_02 = Shock_2$simulation$part
yk_ds_02 = Shock_2$simulation$yk_ds
pc_ds_02 = Shock_2$simulation$pc_ds
emp_02 = Shock_2$simulation$emp
yd_hk_ds_02 = Shock_2$simulation$yd_hk_ds
tilpas_pr_02 = Shock_2$simulation$tilpas_pr
ur_ds_02 = Shock_2$simulation$ur_ds
wage_ds_02 = Shock_2$simulation$wage_ds
part_02 = Shock_2$simulation$part
emp_02 = Shock_2$simulation$emp

part1_02 = Shock_2$simulation$part1
```

```{r}

plot(yk_ds_02, col= "red"); lines(yk_ds_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red"));
grid()

plot(max_dp_02, col= "red"); lines(max_dp_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(wage_02, col= "red"); lines(wage_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(komp_r_02, col= "red"); lines(komp_r_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(part_02, col= "red"); lines(part_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(pc_ds_02, col= "red"); lines(pc_ds_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(emp_02, col= "red"); lines(emp_00, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()
```

## Scenario 3 inflation

We will now look at one of the scenarios where the changes in regulation determined in "skattereformen 2012" was not conducted

```{r}
shock_1_model_shock=shock_1_model %>% window(start= c(2014,2))
```

```{r}
# Define add-factor list (defining shock)
constantAdjList3 <- list(

  shock_part_model = TIMESERIES(shock_part_model_shock+1 ,START=c(2006,3),  FREQ='Q'),
  shock_max_dp_model = TIMESERIES(shock_max_dp_model_shock+1 ,START=c(2006,3),  FREQ='Q'),
 shock_1_model = TIMESERIES(shock_1_model_shock+1 ,START=c(2014,2),  FREQ='Q')

)

```

```{r}
# Simulate the model (following shock)
Shock_3 <- SIMULATE(S_model
                    ,simType='DYNAMIC'
                    ,TSRANGE=c(2006,3,2020,1)
                    ,simConvergence=0.00001
                    ,simIterLimit=100
                    ,ConstantAdjustment=constantAdjList3
                    )
```

```{r}
###Gengivelse af stød til model

yk_ds_03 = Shock_3$simulation$yk_ds
emp_03 = Shock_3$simulation$emp
wage_ds_03 = Shock_3$simulation$wage_ds
ur_ds_03 = Shock_3$simulation$ur_ds
pc_ds_03 = Shock_3$simulation$pc_ds
yd_hk_ds_03 = Shock_3$simulation$yd_hk_ds
pconk_ds_03 = Shock_3$simulation$pconk_ds
komp_r_03 = Shock_3$simulation$komp_r
max_dp_03 = Shock_3$simulation$max_dp
nben_h_03 = Shock_3$simulation$nben_h
yd_hk_ds_03 = Shock_3$simulation$yd_hk_ds

```

```{r}

plot(komp_r_03, col= "red"); lines(komp_r_02, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(yk_ds_03, col= "red", main= "Yk_ds"); lines(yk_ds_02, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(ur_ds_03, col= "red"); lines(ur_ds_02, col = "blue"); 
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(wage_ds_03, col= "red"); lines(wage_ds_02, col = "blue");
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(pc_ds_03, col= "red"); lines(pc_ds_02, col = "blue");
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

plot(yd_hk_ds_03, col= "red"); lines(yd_hk_ds_02, col = "blue");
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

# The last one doesnt fit the working paper, but fits the model used for this replication

plot(max_dp_03, col= "red"); lines(max_dp_02, col = "blue");
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()

```

## Shock plots

### plots S1

```{r}
plot((max_dp_01 / max_dp_00), col= "red");
legend("top", legend=c("Wage growth"),
       lty=c(1),lwd=c(2),bty="n", col = c("red"));
grid()
```

```{r}
plot((max_dp_01b / max_dp_00), col= "red");
legend("top", legend=c("Wage growth"),
       lty=c(1),lwd=c(2),bty="n", col = c("red"));
grid()
```

```{r}
plot(yk_ds_01, col= "red"); lines(yk_ds_00, col = "blue");
legend("top",legend=c("Baseline", "Scenario 1"),lty=c(1,1),lwd=c(2),bty="n", col = c("blue", "red")); 
grid()
```

## Plots S2

```{r}
par(mfrow=c(1,2))
plot(part, col= "red")
plot(part1, col= "black")
```

```{r}
plot(part_01, col= "red");
lines(part_02, col="black");
legend("top", legend=c("Participation exo",
                       "Participation endo"),
       lty=c(1),lwd=c(2),bty="n", col = c("red", "black"));
grid()
```

## Other plots
```{r}
plot(max_dp_03 - max_dp_00, col= "red", ylim= c(0,4));
lines(max_dp_02 - max_dp_00, col="green");
lines(max_dp_01 - max_dp_00, col="black");
legend("top", legend=c("Inflation shock",
                        "participation endo",
                       "Wage growth"),
       lty=c(1),lwd=c(2),bty="n", col = c("red","green", "black"));
grid()
```

```{r}
plot(komp_r_03/ komp_r_00, col= "red");
lines(komp_r_01/ komp_r_00, col="black");
legend("top", legend=c("No suppression",
                       "Wage growth"),
       lty=c(1),lwd=c(2),bty="n", col = c("red", "black"));
grid()
```

```{r}
plot(yd_hk_ds_03/ yd_hk_ds_00, col= "red");
lines(yd_hk_ds_01/ yd_hk_ds_00, col="black");
legend("top", legend=c("No suppression",
                       "Wage growth"),
       lty=c(1),lwd=c(2),bty="n", col = c("red", "black"));
grid()
```

```{r}
plot(emp_03/ emp_00, col= "red");
lines(emp_01/ emp_00, col="black");
legend("top", legend=c("No suppression",
                       "Wage growth"),
       lty=c(1),lwd=c(2),bty="n", col = c("red", "black"));
grid()
```

# Sensitivitets analyse

## Sens1

### Baseline

Vi ændrer værdien i koefficienten i dp_person ligningen
```{r}
# Define add-factor list (defining shock)
constantAdjList_sens1 <- list(

 sens1 = TIMESERIES(sens1+1 ,START=c(2006,3),  FREQ='Q')

)

```

```{r}
# Simulate the model (following shock)
sens_1_base <- SIMULATE(S_model
                    ,simType='DYNAMIC'
                    ,TSRANGE=c(2006,3,2020,1)
                    ,simConvergence=0.00001
                    ,simIterLimit=100
                    ,ConstantAdjustment=constantAdjList_sens1
                    )
```

```{r}
###Gengivelse af stød til model
max_dp_sb = sens_1_base$simulation$max_dp
wage_sb = sens_1_base$simulation$wage
komp_r_sb = sens_1_base$simulation$komp_r
part_sb = sens_1_base$simulation$part
yk_ds_sb = sens_1_base$simulation$yk_ds
pc_ds_sb = sens_1_base$simulation$pc_ds
emp_sb = sens_1_base$simulation$emp
dp_person_sb = sens_1_base$simulation$dp_person

```

```{r}
plot(max_dp_sb, col= "red");
lines(max_dp_00, col="black");
legend("top", legend=c("No suppression",
                       "Wage growth"),
       lty=c(1),lwd=c(2),bty="n", col = c("red", "black"));
grid()
```

### Scenario 1

Vi ændrer værdien i koefficienten i dp_person ligningen
```{r}
# Define add-factor list (defining shock)
constantAdjList1_sens1 <- list(

 sens1 = TIMESERIES(sens1+1 ,START=c(2006,3),  FREQ='Q'),
 shock_max_dp_model = TIMESERIES(shock_max_dp_model_shock+1 ,START=c(2006,3),  FREQ='Q')
)

```

```{r}
# Simulate the model (following shock)
sens_1_S1 <- SIMULATE(S_model
                    ,simType='DYNAMIC'
                    ,TSRANGE=c(2006,3,2020,1)
                    ,simConvergence=0.00001
                    ,simIterLimit=100
                    ,ConstantAdjustment=constantAdjList1_sens1
                    )
```

```{r}
###Gengivelse af stød til model
max_dp_s1 = sens_1_S1$simulation$max_dp
wage_s1 = sens_1_S1$simulation$wage
komp_r_s1 = sens_1_S1$simulation$komp_r
part_s1 = sens_1_S1$simulation$part
yk_ds_s1 = sens_1_S1$simulation$yk_ds
pc_ds_s1 = sens_1_S1$simulation$pc_ds
emp_s1 = sens_1_S1$simulation$emp
```

```{r}
plot((komp_r_s1/komp_r_sb), col= "red");
lines((komp_r_01/komp_r_00), col="black");
legend("top", legend=c("No suppression",
                       "Wage growth"),
       lty=c(1),lwd=c(2),bty="n", col = c("red", "black"));
grid()
```

### Scenario 3

Vi ændrer værdien i koefficienten i dp_person ligningen
```{r}
# Define add-factor list (defining shock)
constantAdjList3_sens1 <- list(

 sens1 = TIMESERIES(sens1+1 ,START=c(2006,3),  FREQ='Q'),
 shock_reg_model = TIMESERIES(shock_reg_model_shock+1 ,START=c(2016,1),  FREQ='Q')
)

```

```{r}
# Simulate the model (following shock)
sens_1_S3 <- SIMULATE(S_model
                    ,simType='DYNAMIC'
                    ,TSRANGE=c(2006,3,2020,1)
                    ,simConvergence=0.00001
                    ,simIterLimit=100
                    ,ConstantAdjustment=constantAdjList3_sens1
                    )
```

```{r}
###Gengivelse af stød til model
max_dp_s3 = sens_1_S3$simulation$max_dp
wage_s3 = sens_1_S3$simulation$wage
komp_r_s3 = sens_1_S3$simulation$komp_r
part_s3 = sens_1_S3$simulation$part
yk_ds_s3 = sens_1_S3$simulation$yk_ds
pc_ds_s3 = sens_1_S3$simulation$pc_ds
emp_s3 = sens_1_S3$simulation$emp
```

```{r}
plot((komp_r_s3/komp_r_sb), col= "red");
lines((komp_r_03/komp_r_00), col="black");
legend("top", legend=c("No suppression",
                       "Wage growth"),
       lty=c(1),lwd=c(2),bty="n", col = c("red", "black"));
grid()
```