### 2022-09-23 Forsøg wage_ds_t.Rmd
commit

---
title: "calculating wage_ds_t"
author: "Simon"
date: "23/9/2022"
output: html_document
---

```{r}
wage_ds_tt= ts(c(73.62489 ,wage_ds_t), start= c(2006,1), frequency=4)
```

```{r}
S_model.txt="MODEL

COMMENT> 
IDENTITY> wage_ds_tt
EQ> wage_ds_tt= wage_ds*(1+inflation_t) 
IF> wage_gap1 > 0.4 & SD1 >=1 
IDENTITY> wage_ds_tt
EQ> wage_ds_tt= max_dp/0.6
IF> wage_gap1 <= 0.4 & SD1 >=1
IDENTITY> wage_ds_tt
EQ> wage_ds_tt= TSLAG(wage_ds_tt,1)
IF> SD2 >= 1
IDENTITY> wage_ds_tt
EQ> wage_ds_tt= TSLAG(wage_ds_tt,1)
IF> SD3 >= 1
IDENTITY> wage_ds_tt
EQ> wage_ds_tt= TSLAG(wage_ds_tt,1)
IF> SD4 >= 1

END"
```

```{r}

#Estimate model coefficients
S_model=ESTIMATE(S_model
                 ,TSRANGE=c(2006,2,2020,1)
)
```

# Simulate

```{r}
# In-sample prediction (no add factors)
S_model <- SIMULATE(S_model
                    ,simType='DYNAMIC'
                    ,TSRANGE=c(2006,2,2020,1)
                    ,simConvergence=0.00001
                    ,simIterLimit=100
                   )
```