### 2022-11-29 DAG.Rmd
Commit

---
title: "DAG test"
author: "Simon"
date: "11/11/2022"
output: html_document
---

```{r}
library(sfcr)
```

# eq all

```{r}

eq= sfcr_set(

  yk ~ pconk + ik + gk + xk - mk,

  pconk ~   TSLAG(LOG(pconk),1) + TSLAG(LOG(yd1_hk),1) + LOG(TSLAG(yd2a_hk,1) + TSLAG(yd2b_hk,1))
  + TSLAG(LOG(fnw_hk),2) + TSDELTALOG(TSLAG(pconk,2)) + TSDELTALOG(TSLAG(pconk,3)) + TSDELTALOG(yd1_hk)
  + TSDELTALOG(yd2a_hk + yd2b_hk),

  ik ~  i_equip_h_k + i_equip_nfc_k  +  i_equip_g_k + i_equip_fc_k +  i_bd_h_k + i_bd_nfc_k  +  i_bd_g_k  + i_bd_fc_k + i_adj_h_k + i_adj_nfc_k + i_adj_fc_k + i_adj_g_k,

  xk ~  TSDELTALOG(TSLAG(gdp_tp,4)) + TSDELTALOG(rer) + LOG(TSLAG(xk,1)) + LOG(TSLAG(gdp_tp,1)) + LOG(TSLAG(rer,2)),

  mk ~ TSDELTALOG(TSLAG(mk,2)) + TSDELTALOG(TSLAG(rer,1)) + TSDELTALOG(TSLAG(rer,3)) + TSDELTALOG(yk) + LOG(TSLAG(mk_ds,1)) + LOG(TSLAG(yk,1)),

  yd1_hk ~ yd1_h/pc,

  yd1_h ~(1 - tax_rate1)*(yh1) - tax_h_adj,

  yh1 ~ w_h_r +  nsben_h + oth_h,

  w_h_r ~ wage * emp,

  emp ~ p_250 * (yk/prodk),

  wage ~ TSDELTA(TSLAG(ur,4)) + TSDELTALOG(prod)  + LOG(TSLAG(wage,1)) +  LOG(TSLAG(wage_t,1)) +  LOG(TSLAG(prod,1)) ,

  prodk~prod * (yk/y),

  gk ~ g/pg,

  yd_hk ~ yd1_hk + yd2a_hk + yd2b_hk,

  i_bd_h_k ~   TSDELTALOG(TSLAG(i_bd_h_k,1)/TSLAG(bd_h_k,2))+ 
    TSDELTALOG(TSLAG(i_bd_h_k,3)/TSLAG(bd_h_k,4))+ 
    TSDELTALOG(TSLAG(p_bd,1)/TSLAG(pi_1,1)) + 
    TSDELTALOG(TSLAG(p_bd,2)/TSLAG(pi_1,2)) + 
    TSDELTALOG(TSLAG(yd_hk,2)/TSLAG(bd_h_k,3))+ 
    TSDELTALOG(TSLAG(-l_h,1)/TSLAG(bd_h,2)) + 
    LOG(TSLAG(i_bd_h_k,1)/TSLAG(bd_h_k,2)) + 
    LOG(TSLAG(yd_hk,1)/TSLAG(bd_h_k,2)) + 
    LOG(TSLAG(p_bd,1)/TSLAG(pi_1,1)) +
    LOG(TSLAG(-l_h,1)/TSLAG(bd_h,2)),

  l_h_tr ~  TSDELTA(TSLAG(l_h_tr,2))
  + TSDELTA(iloan) 
  + TSDELTALOG(TSLAG(i_bd_h_k,3)/TSLAG(yd_hk,3)) 
  + TSLAG(l_h_tr,1) 
  + LOG(TSLAG(-l_h,2)/(TSLAG(yd_hk,2)*TSLAG(pc,2))),

  i_bd_nfc_k ~  TSDELTALOG(TSLAG(i_bd_nfc_k,1)/TSLAG(bd_nfc_k,2)) + 
    TSDELTALOG(ps) + 
    TSDELTALOG(u) + 
    LOG(TSLAG(i_bd_nfc_k,1)/TSLAG(bd_nfc_k,2)) + 
    LOG(TSLAG(ps,1)) + 
    LOG(TSLAG(u,1)) + 
    TSDELTALOG(tobinq) +
    LOG(TSLAG(tobinq,1)),

  ws ~ w_nf/yf,

  w_nf ~ w_h_r +  (w_row_r - w_row_p),

  yf ~ yk*py-((p_tax+p_tax_row)-(p_sub+p_sub_row)),

  ps ~ 1 - ws,

  u~yk/(bd_nfc_k+equip_nfc_k),

   pc ~ TSDELTALOG(TSLAG(pc,1)) + TSDELTALOG(TSLAG(pc,2))+ TSDELTALOG(TSLAG(pc,3)) + TSDELTALOG(TSLAG(pc,4)) 
  + TSDELTALOG(wage)+ TSDELTALOG(pm) + TSDELTALOG(TSLAG(pm,2)) + LOG(TSLAG(pc,1)) + LOG(TSLAG(wage,1))
  + LOG(TSLAG(prod,1)) + LOG(TSLAG(pm,1)),

  rer ~ pc*xr/pf,

   i_equip_nfc_k ~  TSDELTALOG(TSLAG(i_equip_nfc_k,1)/TSLAG(equip_nfc_k,2)) +
    LOG(TSLAG(i_equip_nfc_k,1)/TSLAG(equip_nfc_k,2)) +
     LOG(TSLAG(ps,1)) +
    LOG(TSLAG(u,1)) +
    TSDELTALOG(tobinq) +
    LOG(TSLAG(tobinq,1))

)

```

# Table

```{r, fig.width= 15, fig.height=10}
tb_graph=sfcr_dag_cycles(eq)

```

```{r, fig.width= 15, fig.height=10}
library(ggraph)
library(tidygraph)
tb_graph %N>% 
  filter(!node_is_isolated()) %>%
  ggraph(layout = 'stress') + 
  geom_edge_fan(strength=3 ,arrow = arrow(angle = 30, length = unit(0.35, 'cm'),type = 'closed'), alpha = 0.4) + 
  geom_node_point(aes(col = cyclical, size = 3)) +
  geom_node_text(aes(label = name, size = 3),nudge_y=0) + 
  theme_graph() +
  theme(legend.position = "bottom") 
```

# 3d plot

```{r}
library(igraph)

coords <- layout_with_fr(tb_graph, dim=3)

rglplot(tb_graph,
        layout=coords,
        label.color = "red",
        vertex.color= NA,
        frame.color= "green",
        label.degree= pi)
```